### 2024-08-01 code/01_data_checks.qmd
init repo

---
title: "Data checks"
author: "Stefano Coretta"
format: html
---

```{r}
#| label: setup

library(tidyverse)
library(coretta2018itaegg)
```

```{r}
#| label: load-data
#| 
data("ita_egg")
```

```{r}
ita_egg |> 
  ggplot(aes(voi_clo_prop)) +
  geom_density()

ita_egg |> 
  filter(voi_clo_prop > 0, voi_clo_prop < 1) |> 
  ggplot(aes(voi_clo_prop)) +
  geom_density() +
  geom_rug()
```

```{r}
ita_egg_filt <- ita_egg |> 
  filter(voi_clo_prop > 0, voi_clo_prop < 1)
```

```{r}
ita_egg_filt |> 
  ggplot(aes(vowel, voi_clo_prop)) +
  geom_jitter(alpha = 0.1, width = 0.2) +
  stat_summary(fun.data = "mean_cl_boot")
```

```{r}
ita_egg_filt |> 
  ggplot(aes(vowel, voi_clo_prop)) +
  geom_jitter(alpha = 0.2, width = 0.2, aes(colour = vowel)) +
  stat_summary(fun.data = "mean_cl_boot") +
  facet_grid(cols = vars(c2))
```

```{r}
ita_egg_filt |> 
  ggplot(aes(vowel, voi_clo_prop)) +
  stat_summary(fun.data = "mean_cl_boot") +
  facet_grid(cols = vars(c2))
```

```{r}
ita_egg_filt |> 
  ggplot(aes(speech_rate, voi_clo_prop)) + 
  geom_point(alpha = 0.5)
```

### 2024-08-01 manuscript/manuscript.qmd
init repo

---
title: A tutorial of Bayesian beta regressions with brms in R
author:
  - name: Stefano Coretta
    email: s.coretta@ed.ac.uk
    affiliations: 
        - id: uoe
          name: University of Edinburgh
          department: Linguistics and English Language
          address: 3 George Sq
          city: Edinburgh
          state: United Kingdom
          postal-code: EH8 9AD
    attributes:
        corresponding: true
    note: This is the first author footnote.
abstract: |
  This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae. Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at tempus elit.
keywords: 
  - keyword1
  - keyword2
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: authoryear
---

### 2024-08-01 manuscript/manuscript.tex
init repo

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  authoryear,
  preprint,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\journal{Journal Name}

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A tutorial of Bayesian beta regressions with brms in R},
  pdfauthor={Stefano Coretta},
  pdfkeywords={keyword1, keyword2},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{A tutorial of Bayesian beta regressions with brms in R}
\author[1]{Stefano Coretta%
\corref{cor1}%
\fnref{fn1}}
 \ead{s.coretta@ed.ac.uk}

\affiliation[1]{organization={University of Edinburgh, Linguistics and
English Language},addressline={3 George
Sq},city={Edinburgh},postcode={EH8 9AD},postcodesep={}}

\cortext[cor1]{Corresponding author}
\fntext[fn1]{This is the first author footnote.}

\begin{abstract}
This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing
elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget
velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris
sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta
egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet
auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo
molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae.
Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at
tempus elit.
\end{abstract}

\begin{keyword}
    keyword1 \sep 
    keyword2
\end{keyword}
\end{frontmatter}

  \bibliography{bibliography.bib}

\end{document}

### 2024-08-01 manuscript/manuscript.qmd
edit manuscript

@@
 ---
+
+## Introduction
+
+Phonetic research often involves numeric continuous outcome variables, like durations, frequencies, loudness and ratios.
+Another commonly employed type of outcome variable are proportions: for example, proportion of voicing during closure, vocal folds contact quotient, gesture amplitude, nasalance.
+Moreover, virtually any measure can be MIN-MAX normalised, a procedure which transforms values so that they are in the range 0--1.
+
+Regression models are very common, but there is a tendency of using Gaussian distribution families (i.e. probability distributions for the outcome variable) for anything that is numeric.
+A possible reason is that the base R function for fitting regression models, `lm()`, and the lme4 function used to fit regression models with varying terms, `lmer()`, both fit Gaussian regressions by default and the user does not have to specify the distribution family.
+This tacit defaulting to Gaussian models is also reflected in teaching practices, where test and models using the Gaussian distribution are the first to be taught, due to their relative simplicity and the fact that other models are generalisations of Gaussian models.
+
+However, proportion are naturally not Gaussian, since they are limited between 0 and 1.
+The theoretical distribution that generates values with this characteristics is the beta distribution.
+Thus, regression models with proportions as the outcome variable should be fitted using a beta distribution as the distribution family.
+This tutorial introduces researchers to beta regression models in R using the package brms.
+Familiarity with regression modelling in R with a package like lme4 is assumed, but no prior knowledge of Bayesian statistics is necessary.
+
+## The beta distribution
+
+...
+
+## Case study 1: voicing within consonant closure
+
+```{r}
+library(tidyverse)
+theme_set(theme_light())
+
+load("data/coretta2018/ita_egg.rda")
+
+ita_egg <- ita_egg |> 
+  filter(voi_clo_prop > 0, voi_clo_prop < 1) |> 
+  mutate(
+    vowel = factor(vowel, levels = c("i", "e", "a", "o", "u")),
+    c2 = factor(c2, levels = c("k", "t", "p"))
+  )
+```
+
+```{r}
+ita_egg
+```
+
+```{r}
+#| label: fig-ita-egg
+
+ita_egg |> 
+  ggplot(aes(vowel, voi_clo_prop)) +
+  geom_jitter(alpha = 0.2, width = 0.2, aes(colour = vowel)) +
+  stat_summary(fun.data = "mean_cl_boot") +
+  facet_grid(cols = vars(c2)) +
+  scale_color_brewer(palette = "Dark2") +
+  labs(
+    x = "Vowel", y = "Voicing during closure (prop)"
+  ) +
+  theme(legend.position = "none")
+```
+
+```{r}
+#| label: brms
+#| message: false
+
+library(brms)
+```
+
+```{r}
+# Takes 3 minutes on MacBook Pro 2020, M1
+
+voi_prop_bm <- brm(
+  voi_clo_prop ~
+    0 + vowel:c2 + speech_rate_c +
+    (0 + vowel:c2 + speech_rate_c | speaker),
+  family = Beta,
+  data = ita_egg,
+  cores = 4,
+  seed = 3749,
+  file = "data/cache/voi_prop_bm"
+)
+```
+
+```{r}
+summary(voi_prop_bm, prob = 0.8)
+```
+
+```{r}
+conditional_effects(voi_prop_bm, "c2:vowel")
+```
+
+```{r}
+library(marginaleffects)
+
+comp_c2 <- avg_comparisons(voi_prop_bm, variables = list(c2 = "pairwise"), conf_level = 0.8, type = "prediction")
+```

### 2024-08-02 manuscript/manuscript.qmd
edit manuscript

@@
 
+```{r}
+#| label: setup
+#| include: false
+
+library(tidyverse)
+library(brms)
+```
+
 ## Introduction
@@
 
+For the first case study, we will model the proportion of voicing within consonant closure.
+The measurements come from a data set of audio and electroglottographic (EGG) recordings of 19 speakers of Northwestern Italian.
+The participants read frame sentences which included target words of the form /CVCo/, where /C/ was either /k, t, p/ in all permutations and /V/ was either /i, e, a, ɔ, u/ (two resulting words, /peto/ and /kako/ were excluded because they are profanities), for a total of 43 target words.
+There were 4 different frame sentence, with a total of 172 trials per participant (3,268 grand total).
+The actual observation count is 2,419, after removing speech errors, EGG measurement errors, and cases in which voicing ceased before the closure onset/after the closure offset of the second /C/.
+
+The proportion of voicing during the closure of the second /C/ was calculated as the proportion of contiguous voicing duration after closure onset to total duration of closure.
+The following code chunk attaches the tidyverse packages (for reading and wrangling data) and loads the `ita_egg` tibble (data frame).
+The tibble is filtered so as to remove voicing proportions (`voi_clo_prop`) that are smaller than 0 and greater than 1.
+The variables `vowel` and `c2` are converted to factors to specify the order of the levels.
+
 ```{r}
+#| label: read-ita-egg
+
+# attach tidyverse and set light theme for plots
 library(tidyverse)
@@
 
+# load tibble
 load("data/coretta2018/ita_egg.rda")
 
+# filter and mutate data
 ita_egg <- ita_egg |> 
@@
 
+Here is what the tibble looks like.
+
 ```{r}
+#| label: ita-egg
+
 ita_egg
@@
 
+@fig-ita-egg shows the raw voicing duration proportion values split by vowel /i, e, a, ɔ, u/ and second consonant /k, t, p/ in the /CVCo/ target words.
+The plot suggests that, on average, the voicing proportion is slightly lower with /k/ than with /t, p/.
+Moreover, there is greater variability between vowels in /t, p/ than in /k/.
+We will use a beta regression model to assess these patterns.
+["expectations" XXX]
+
 ```{r}
 #| label: fig-ita-egg
+#| fig-cap: "Proportion of voicing during the closure of the second consonant in /CVCo/ words by vowel and the second consonant."
+#| echo: false
 
@@
 
+We will use brms to fit Bayesian beta regressions. [XXX why Bayesian]. 
+The model will have voicing proportion as the outcome variable and the following terms: vowel (/i, e, a, ɔ, u/) and second consonant C2 (/k, t, p/) in interaction, centred speech rate, a varying intercept by speaker and varying slopes of vowel/consonant in interaction and centred speech rate (number of syllables per second).
+The categorical predictors vowel and C2 were coded using indexing (in R, this corresponds to suppressing the intercept with the `0 +` syntax; this type of predictor coding does not impact the model predictions, but it makes it easier to specify priors when users do so).
+For pedagogical reasons of simplicity, the model will use the default priors, but note that in real data analyses contexts, priors should be specified by the user.
+I refer the readers to XXX.
+
 ```{r}
-#| label: brms
 #| message: false
 
+# attach brms
 library(brms)
-```
 
-```{r}
+# fit the model
 # Takes 3 minutes on MacBook Pro 2020, M1
-
 voi_prop_bm <- brm(
+  # model formula
   voi_clo_prop ~
+    # constant terms
     0 + vowel:c2 + speech_rate_c +
+    # varying terms
     (0 + vowel:c2 + speech_rate_c | speaker),
+  # uses the beta family for the outcome
   family = Beta,
@@
 
+The `summary()` function prints the full model summary.
+For conciseness, we will use the `fixef()` function which prints the regression coefficients of the constant (aka fixed) terms.
+The full summary with an explanation of each part can be found in XXX.
+For each coefficient in the model, `fixef()` prints the name of the coefficient, the mean estimate, the estimate error and the lower and upper limits of a Bayesian Credible interval (CrI).
+Here, we print an 80% CrI.^[There is nothing special about 95% within Bayesian inference and instead experts recommend to check and report a variety of CrIs.
+For simplicity, I will focus on 80% CrIs in this case study.] In Bayesian inference, an 80% CrI indicates the range of values within which the true estimate falls at 80% probability or confidence.
+I round all values to the nearest 2 digits for clarity.
+
 ```{r}
-summary(voi_prop_bm, prob = 0.8)
+round(
+  fixef(voi_prop_bm, prob = c(0.1, 0.9)),
+  digits = 2
+)
 ```
+
+The coefficients of a beta regression are estimated on the log-odds scale, as in Bernoulli/binomial (aka logistic) regressions.
+From the summary, we see that speech rate (number of syllables per second) has a positive effect on voicing proportion: the 80% CrI is between 0.01 and 0.15 log-odds [$\beta$ = 0.8, SD = 0.06].
+Log-odds can be converted to odd ratios by exponentiating the value: 0.01-0.15 log odds correspond to an odd ratio of 1.01 to 1.16, or as percentages, to an increase of voicing of 1 to 16% for every increase of one syllable per second.
+Since this is an 80% CrI, we can be 80% confident that the true effect of speech rate is between 1-16% increase of voicing proportion.
+
+Turning now to the coefficients for vowel and C2, given the indexing approach of coding these variables the model summary and the output of `fixef()` reports the *predictions* in log-odds for each combination of vowel and C2, rather than differences between levels.
+The CrIs of the vowel/C2 coefficients span all negative log-odds values: these correspond to proportions that are lower than 0.5 (which is 0 in log-odds).
+This matches the general trends in the raw data, which we plotted in @fig-ita-egg.
+
+Next, we will plot the predicted proportions of each vowel/C2 combination at mean speech rate (i.e.
+centred speech rate = 0) and then calculate the average pair-wise difference in voicing proportion between /k, t, p/.
+Finally, we will assess whether there is greater between-vowel variability in /t, p/ relative to /k/.
+
+Before being able to plot the predictions, it's important to get familiar with the so called posterior draws.
+[Bayesian MCMC XXX].
+Posterior draws can be conveniently obtained with the `as_draws_df()`.
+For the moment, we can extract only the draws of the constant regression coefficients (model variables starting with `b_`).
+To check which coefficients are available in a model, use `get_variables()` from the tidybayes package.
+`as_draws_df()` returns a tibble where each column contains the drawn values of a coefficient.
+The probability distribution of the drawn values of each coefficient is the posterior probability distribution of that coefficient.
+Note that, due to the indexing coding of vowel and C2, all coefficient except `b_speech_rate_c` are *predicted log-odds* for each vowel/C2 combination (the drawn values for `b_speech_rate_c` are drawn *differences* in log-odds for each unit increase of speech rate).
+The drawn values are in log-odds, but we can convert them to proportions with `plogis()` (we will do this when plotting below).
+
+```{r}
+#| label: voi-prop-bm-draws
+
+# extract only coefficient variables starting with "b_"
+voi_prop_bm_draws <- as_draws_df(voi_prop_bm, variable = "^b_", regex = TRUE)
+voi_prop_bm_draws
+```
+
+We can now wrangle this tibble and plot the posterior distributions for each vowel/C2 combination.
+
+```{r}
+voi_prop_bm_draws_long <- voi_prop_bm_draws |> 
+  select(-b_speech_rate_c) |> 
+  pivot_longer(`b_voweli:c2k`:`b_vowelu:c2p`, names_to = "coeff") |> 
+  separate(coeff, into = c("type", "vowel", "c2"))
+voi_prop_bm_draws_long
+```
+
+For plotting, we can use ggplot2 statistics from the ggdist package.
+`stat_halfeye()` plots the density of the posterior probability (in grey), its median (point) and CrIs (lines).
+Let's use 60 and 80% CrIs and transform the log-odds values to proportions with `plogis()`.
+
+```{r}
+#| label: fig-voi-prop-bm-1
+# attach ggdist package
+library(ggdist)
+
+voi_prop_bm_draws_long |> 
+  ggplot(aes(plogis(value), vowel)) +
+  stat_halfeye(.width = c(0.6, 0.8)) +
+  facet_grid(rows = vars(c2))
+```
+
+What if we want to plot the average predicted voicing proportion for the three consonants /k, t, p/?
+One approach is to take the mean across vowels within each consonant and the resulting posterior distribution is the predicted posterior distribution of voicing proportion for each consonant, assuming an "average" vowel.
+
+```{r}
+voi_prop_bm_draws_long_c2 <- voi_prop_bm_draws_long |> 
+  group_by(.draw, c2) |> 
+  summarise(
+    value_mean = mean(value), .groups = "drop"
+  )
+
+voi_prop_bm_draws_long_c2 |> 
+  ggplot(aes(plogis(value_mean), c2)) +
+  stat_halfeye(.width = c(0.6, 0.8))
+```
+
+Based on the posterior distributions of the mean voicing proportion by consonant, /p/ has a somewhat higher voicing proportion than /k/ and /t/.
+The real question is: how higher?
+We can quantify this by taking the difference of the drawn values for /p/ and those for /t, k/.
+Since we want to compare /t, k/ to /p/, we should first average the draws of /t, k/ and then take the difference of the averaged draws and the draws of /p/.
+
+```{r}
+voi_prop_bm_diff <- voi_prop_bm_draws_long_c2 |> 
+  pivot_wider(names_from = c2, values_from = value_mean) |> 
+  mutate(
+    c2tk = mean(c(c2k, c2t)),
+    c2p_tk_diff = c2p - c2tk
+  )
+
+voi_prop_bm_diff |> 
+  ggplot(aes(c2p_tk_diff)) +
+  stat_halfeye(.width = c(0.6, 0.8, 0.9)) +
+  geom_vline(xintercept = 0)
+```
+
+Once we have the posterior difference, we can obtain CrIs of that using `quantile2()` from the posterior package.
+Beware that the values of the difference are in log-odds!
+We can convert these into odd ratios with `exp()`.
+
+```{r}
+library(posterior)
+
+voi_prop_bm_diff |> 
+  reframe(
+    # 90% CrI
+    q90 = quantile2(c2p_tk_diff, probs = c(0.05, 0.95)),
+    # 80% CrI
+    q80 = quantile2(c2p_tk_diff, probs = c(0.1, 0.9)),
+    # 60% CrI
+    q60 = quantile2(c2p_tk_diff, probs = c(0.2, 0.8)),
+  ) |> 
+  # exponentiate to get odd ratios and round
+  mutate(across(everything(), ~round(exp(.x), 2)))
+```
+
+So, based on the model and data, there is a 90% probability that the voicing proportion in /p/ is 1.07-1.38 times longer (or 7-38% increase) than in /t, k/.
+At 80% confidence, the change ratio is 1.10-1.34 (or 10-34% increase) while at 60% confidence is 1.14-1.30 (14-30% increase).
+In other words we can be quite confident that the voicing proportion in /p/ is longer than in /t, k/ and that the increase is less than 35%.
 
@@
 
-comp_c2 <- avg_comparisons(voi_prop_bm, variables = list(c2 = "pairwise"), conf_level = 0.8, type = "prediction")
+avg_comparisons(voi_prop_bm, variables = list(c2 = "pairwise"), conf_level = 0.8)
+avg_predictions(voi_prop_bm, variables = "vowel", conf_level = 0.8)
 ```
+
+## Case study 2: pre-nasalisation in Greek

### 2024-11-25 code/01_data_checks.qmd
edit manuscript

@@
 library(coretta2018itaegg)
+library(Hmisc)
 ```
@@
   ggplot(aes(speech_rate, voi_clo_prop)) + 
-  geom_point(alpha = 0.5)
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "lm") +
+  facet_wrap(vars(vowel))
 ```

### 2024-11-25 manuscript/manuscript.qmd
edit manuscript

@@
 library(brms)
+library(collapse)
 ```
 
-## Introduction
+# Introduction
 
@@
 
-## The beta distribution
+# The beta distribution
 
@@
 
-## Case study 1: voicing within consonant closure
+# Case study 1: voicing within consonant closure
 
@@
 We will use brms to fit Bayesian beta regressions. [XXX why Bayesian]. 
-The model will have voicing proportion as the outcome variable and the following terms: vowel (/i, e, a, ɔ, u/) and second consonant C2 (/k, t, p/) in interaction, centred speech rate, a varying intercept by speaker and varying slopes of vowel/consonant in interaction and centred speech rate (number of syllables per second).
-The categorical predictors vowel and C2 were coded using indexing (in R, this corresponds to suppressing the intercept with the `0 +` syntax; this type of predictor coding does not impact the model predictions, but it makes it easier to specify priors when users do so).
-For pedagogical reasons of simplicity, the model will use the default priors, but note that in real data analyses contexts, priors should be specified by the user.
+The model has voicing proportion as the outcome variable and the following terms: an interaction between vowel (/i, e, a, ɔ, u/) and second consonant C2 (/k, t, p/), centred speech rate (number of syllables per second); as varying (aka random) terms, by-speaker varying coefficients for the vowel/consonant interaction and for centred speech rate.^[Footnote about Gelman's terminology for random effects.] The categorical predictors vowel and C2 are coded using indexing rather than traditional R contrasts (in R, this corresponds to suppressing the model's intercept with the `0 +` syntax; using indexing instead of contrasts does not impact the model posterior predictions, but it makes it easier to specify priors when users do so).
+For pedagogical simplicity, the model will use the default priors, but note that in real data analyses contexts, priors should be specified by the user.
 I refer the readers to XXX.
@@
 ```{r}
+#| label: voi-prop-bm
 #| message: false
@@
 Here, we print an 80% CrI.^[There is nothing special about 95% within Bayesian inference and instead experts recommend to check and report a variety of CrIs.
-For simplicity, I will focus on 80% CrIs in this case study.] In Bayesian inference, an 80% CrI indicates the range of values within which the true estimate falls at 80% probability or confidence.
-I round all values to the nearest 2 digits for clarity.
+For simplicity, we will use 80% CrIs in this case study.] In Bayesian inference, an 80% CrI indicates the range of values within which the true estimate falls at 80% probability or confidence.
+We round all values to the nearest 2 digits for clarity.
 
 ```{r}
+#| label: voi-prop-bm-fixef
 round(
@@
 
-Before being able to plot the predictions, it's important to get familiar with the so called posterior draws.
+Before being able to plot the predictions, it's important to get familiar with the so-called posterior draws.
 [Bayesian MCMC XXX].
 Posterior draws can be conveniently obtained with the `as_draws_df()`.
-For the moment, we can extract only the draws of the constant regression coefficients (model variables starting with `b_`).
+For the moment, we will extract only the draws of the constant regression coefficients (model variables starting with `b_`).
 To check which coefficients are available in a model, use `get_variables()` from the tidybayes package.
@@
 ```{r}
+#| label: voi-prop-bm-draws-long
+
 voi_prop_bm_draws_long <- voi_prop_bm_draws |> 
+  # drop b_speech_rate_c before pivoting
   select(-b_speech_rate_c) |> 
+  # pivot vowel:c2 columns
   pivot_longer(`b_voweli:c2k`:`b_vowelu:c2p`, names_to = "coeff") |> 
+  # separate "coeff" labels into type ("b"), vowel and c2
   separate(coeff, into = c("type", "vowel", "c2"))
+
 voi_prop_bm_draws_long
@@
 
-For plotting, we can use ggplot2 statistics from the ggdist package.
+For plotting, we can use ggplot2 statistics layers from the ggdist package.
 `stat_halfeye()` plots the density of the posterior probability (in grey), its median (point) and CrIs (lines).
@@
 #| label: fig-voi-prop-bm-1
+
 # attach ggdist package
@@
 What if we want to plot the average predicted voicing proportion for the three consonants /k, t, p/?
-One approach is to take the mean across vowels within each consonant and the resulting posterior distribution is the predicted posterior distribution of voicing proportion for each consonant, assuming an "average" vowel.
+One approach is to take the mean across vowels within each consonant for each posterior draw, and the posterior distribution of the resulting list of values is the predicted posterior distribution of voicing proportion for each consonant, assuming an "average" vowel.
 
 ```{r}
+#| label: voi-prop-bm-draws-long-c2
+
 voi_prop_bm_draws_long_c2 <- voi_prop_bm_draws_long |> 
+  # grouping by .draw and c2 ensures that averaging applies only within draw and c2
   group_by(.draw, c2) |> 
+  # calculate the mean value within draw/c2
   summarise(
@@
 Based on the posterior distributions of the mean voicing proportion by consonant, /p/ has a somewhat higher voicing proportion than /k/ and /t/.
-The real question is: how higher?
+The real question is: how much higher?
 We can quantify this by taking the difference of the drawn values for /p/ and those for /t, k/.
@@
 ```{r}
+#| label: voi-prop-bm-diff
+
 voi_prop_bm_diff <- voi_prop_bm_draws_long_c2 |> 
+  # pivot data to create one column per consonant with the mean drawn values,
+  # with one draw per raw
   pivot_wider(names_from = c2, values_from = value_mean) |> 
   mutate(
+    # calculate the mean of /k/ and /t/, for each draw
     c2tk = mean(c(c2k, c2t)),
+    # calculate the difference of /p/ and /t, k/
     c2p_tk_diff = c2p - c2tk
   )
+
+voi_prop_bm_diff
+```
+
+```{r}
+#| label: fig-diff-p-tk
 
@@
 
-Once we have the posterior difference, we can obtain CrIs of that using `quantile2()` from the posterior package.
+Once we have the posterior difference, we can obtain CrIs of the difference using `quantile2()` from the posterior package.
 Beware that the values of the difference are in log-odds!
@@
 ```{r}
+#| label: voi-prop-bm-diff-cri
+
 library(posterior)
@@
   ) |> 
-  # exponentiate to get odd ratios and round
+  # exponentiate to get odd ratios and round to 2 digits
   mutate(across(everything(), ~round(exp(.x), 2)))
@@
 
+brms comes with a convenient function, `conditional_effects()`, to plot posterior means and CrIs based on predictors in the model.
+In the following example, we plot the predicted proportion of voicing by consonant and vowel.
+
 ```{r}
+#| label: voi-prop-bm-cond
 conditional_effects(voi_prop_bm, "c2:vowel")
@@
 
+Finally, the package marginaleffects has two other convenience functions that return CrIs of comparisons across predictor levels (`avg_comparisons()`) and CrIs of posterior predictions across predictor levels (`avg_predictions`).
+[XXX]
+
 ```{r}
+#| label: marginal
+
 library(marginaleffects)
 
-avg_comparisons(voi_prop_bm, variables = list(c2 = "pairwise"), conf_level = 0.8)
+avg_comparisons(voi_prop_bm, variables = list(c2 = "pairwise"), conf_level = 0.8, type = "link")
 avg_predictions(voi_prop_bm, variables = "vowel", conf_level = 0.8)
@@
 
-## Case study 2: pre-nasalisation in Greek
+# Case study 2: pre-nasalisation in Greek

### 2024-11-25 manuscript/manuscript.tex
edit manuscript

@@
 \makeatother
+
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{241,243,245}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
+\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
+\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
 
@@
 
+\section{Introduction}\label{introduction}
+
+Phonetic research often involves numeric continuous outcome variables,
+like durations, frequencies, loudness and ratios. Another commonly
+employed type of outcome variable are proportions: for example,
+proportion of voicing during closure, vocal folds contact quotient,
+gesture amplitude, nasalance. Moreover, virtually any measure can be
+MIN-MAX normalised, a procedure which transforms values so that they are
+in the range 0--1.
+
+Regression models are very common, but there is a tendency of using
+Gaussian distribution families (i.e.~probability distributions for the
+outcome variable) for anything that is numeric. A possible reason is
+that the base R function for fitting regression models, \texttt{lm()},
+and the lme4 function used to fit regression models with varying terms,
+\texttt{lmer()}, both fit Gaussian regressions by default and the user
+does not have to specify the distribution family. This tacit defaulting
+to Gaussian models is also reflected in teaching practices, where test
+and models using the Gaussian distribution are the first to be taught,
+due to their relative simplicity and the fact that other models are
+generalisations of Gaussian models.
+
+However, proportion are naturally not Gaussian, since they are limited
+between 0 and 1. The theoretical distribution that generates values with
+this characteristics is the beta distribution. Thus, regression models
+with proportions as the outcome variable should be fitted using a beta
+distribution as the distribution family. This tutorial introduces
+researchers to beta regression models in R using the package brms.
+Familiarity with regression modelling in R with a package like lme4 is
+assumed, but no prior knowledge of Bayesian statistics is necessary.
+
+\section{The beta distribution}\label{the-beta-distribution}
+
+\ldots{}
+
+\section{Case study 1: voicing within consonant
+closure}\label{case-study-1-voicing-within-consonant-closure}
+
+For the first case study, we will model the proportion of voicing within
+consonant closure. The measurements come from a data set of audio and
+electroglottographic (EGG) recordings of 19 speakers of Northwestern
+Italian. The participants read frame sentences which included target
+words of the form /CVCo/, where /C/ was either /k, t, p/ in all
+permutations and /V/ was either /i, e, a, ɔ, u/ (two resulting words,
+/peto/ and /kako/ were excluded because they are profanities), for a
+total of 43 target words. There were 4 different frame sentence, with a
+total of 172 trials per participant (3,268 grand total). The actual
+observation count is 2,419, after removing speech errors, EGG
+measurement errors, and cases in which voicing ceased before the closure
+onset/after the closure offset of the second /C/.
+
+The proportion of voicing during the closure of the second /C/ was
+calculated as the proportion of contiguous voicing duration after
+closure onset to total duration of closure. The following code chunk
+attaches the tidyverse packages (for reading and wrangling data) and
+loads the \texttt{ita\_egg} tibble (data frame). The tibble is filtered
+so as to remove voicing proportions (\texttt{voi\_clo\_prop}) that are
+smaller than 0 and greater than 1. The variables \texttt{vowel} and
+\texttt{c2} are converted to factors to specify the order of the levels.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# attach tidyverse and set light theme for plots}
+\FunctionTok{library}\NormalTok{(tidyverse)}
+\FunctionTok{theme\_set}\NormalTok{(}\FunctionTok{theme\_light}\NormalTok{())}
+
+\CommentTok{\# load tibble}
+\FunctionTok{load}\NormalTok{(}\StringTok{"data/coretta2018/ita\_egg.rda"}\NormalTok{)}
+
+\CommentTok{\# filter and mutate data}
+\NormalTok{ita\_egg }\OtherTok{\textless{}{-}}\NormalTok{ ita\_egg }\SpecialCharTok{|\textgreater{}} 
+  \FunctionTok{filter}\NormalTok{(voi\_clo\_prop }\SpecialCharTok{\textgreater{}} \DecValTok{0}\NormalTok{, voi\_clo\_prop }\SpecialCharTok{\textless{}} \DecValTok{1}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
+  \FunctionTok{mutate}\NormalTok{(}
+    \AttributeTok{vowel =} \FunctionTok{factor}\NormalTok{(vowel, }\AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\StringTok{"i"}\NormalTok{, }\StringTok{"e"}\NormalTok{, }\StringTok{"a"}\NormalTok{, }\StringTok{"o"}\NormalTok{, }\StringTok{"u"}\NormalTok{)),}
+    \AttributeTok{c2 =} \FunctionTok{factor}\NormalTok{(c2, }\AttributeTok{levels =} \FunctionTok{c}\NormalTok{(}\StringTok{"k"}\NormalTok{, }\StringTok{"t"}\NormalTok{, }\StringTok{"p"}\NormalTok{))}
+\NormalTok{  )}
+\end{Highlighting}
+\end{Shaded}
+
+Here is what the tibble looks like.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{ita\_egg}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+# A tibble: 2,419 x 53
+   speaker ipu    stimulus    sentence_ons sentence_off word_ons word_off v1_ons
+   <chr>   <chr>  <chr>              <dbl>        <dbl>    <dbl>    <dbl>  <dbl>
+ 1 it01    ipu_1  Ripete 'po~         13.2         14.9     13.7     14.1   13.9
+ 2 it01    ipu_2  Ripete 'to~         16.9         18.6     17.4     17.9   17.5
+ 3 it01    ipu_3  Ripete 'pa~         20.2         21.9     20.7     21.1   20.8
+ 4 it01    ipu_4  Sentivo 't~         23.5         25.1     24.0     24.4   24.1
+ 5 it01    ipu_5  Sentivo 't~         26.3         27.8     26.8     27.2   26.9
+ 6 it01    ipu_6  Scrivete '~         29.2         30.9     29.7     30.1   29.8
+ 7 it01    ipu_7  Sentivo 'c~         32.1         33.6     32.6     33.1   32.8
+ 8 it01    ipu_8  Scrivete '~         35.0         36.6     35.5     35.9   35.6
+ 9 it01    ipu_9  Ha detto '~         41.9         43.5     42.3     42.7   42.4
+10 it01    ipu_10 Sentivo 'p~         47.4         48.9     47.9     48.4   48.1
+# i 2,409 more rows
+# i 45 more variables: c2_ons <dbl>, v2_ons <dbl>, voice_ons <dbl>,
+#   voice_off <dbl>, c1_rel <dbl>, c2_rel <dbl>, stimulus_id <dbl>,
+#   sentence <chr>, word <chr>, c1 <chr>, vowel <fct>, c2 <fct>,
+#   backness <chr>, height <fct>, c1_place <fct>, c2_place <fct>,
+#   v1_duration <dbl>, c2_clos_duration <dbl>, rel_voff <dbl>,
+#   sent_duration <dbl>, speech_rate <dbl>, speech_rate_c <dbl>, ...
+\end{verbatim}
+
+Figure~\ref{fig-ita-egg} shows the raw voicing duration proportion
+values split by vowel /i, e, a, ɔ, u/ and second consonant /k, t, p/ in
+the /CVCo/ target words. The plot suggests that, on average, the voicing
+proportion is slightly lower with /k/ than with /t, p/. Moreover, there
+is greater variability between vowels in /t, p/ than in /k/. We will use
+a beta regression model to assess these patterns. {[}``expectations''
+XXX{]}
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{manuscript_files/figure-pdf/fig-ita-egg-1.pdf}
+
+}
+
+\caption{\label{fig-ita-egg}Proportion of voicing during the closure of
+the second consonant in /CVCo/ words by vowel and the second consonant.}
+
+\end{figure}%
+
+We will use brms to fit Bayesian beta regressions. {[}XXX why
+Bayesian{]}. The model has voicing proportion as the outcome variable
+and the following terms: an interaction between vowel (/i, e, a, ɔ, u/)
+and second consonant C2 (/k, t, p/), centred speech rate (number of
+syllables per second); as varying (aka random) terms, by-speaker varying
+coefficients for the vowel/consonant interaction and for centred speech
+rate.\footnote{Footnote about Gelman's terminology for random effects.}
+The categorical predictors vowel and C2 are coded using indexing rather
+than traditional R contrasts (in R, this corresponds to suppressing the
+model's intercept with the \texttt{0\ +} syntax; using indexing instead
+of contrasts does not impact the model posterior predictions, but it
+makes it easier to specify priors when users do so). For pedagogical
+simplicity, the model will use the default priors, but note that in real
+data analyses contexts, priors should be specified by the user. I refer
+the readers to XXX.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# attach brms}
+\FunctionTok{library}\NormalTok{(brms)}
+
+\CommentTok{\# fit the model}
+\CommentTok{\# Takes 3 minutes on MacBook Pro 2020, M1}
+\NormalTok{voi\_prop\_bm }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
+  \CommentTok{\# model formula}
+\NormalTok{  voi\_clo\_prop }\SpecialCharTok{\textasciitilde{}}
+    \CommentTok{\# constant terms}
+    \DecValTok{0} \SpecialCharTok{+}\NormalTok{ vowel}\SpecialCharTok{:}\NormalTok{c2 }\SpecialCharTok{+}\NormalTok{ speech\_rate\_c }\SpecialCharTok{+}
+    \CommentTok{\# varying terms}
+\NormalTok{    (}\DecValTok{0} \SpecialCharTok{+}\NormalTok{ vowel}\SpecialCharTok{:}\NormalTok{c2 }\SpecialCharTok{+}\NormalTok{ speech\_rate\_c }\SpecialCharTok{|}\NormalTok{ speaker),}
+  \CommentTok{\# uses the beta family for the outcome}
+  \AttributeTok{family =}\NormalTok{ Beta,}
+  \AttributeTok{data =}\NormalTok{ ita\_egg,}
+  \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
+  \AttributeTok{seed =} \DecValTok{3749}\NormalTok{,}
+  \AttributeTok{file =} \StringTok{"data/cache/voi\_prop\_bm"}
+\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+The \texttt{summary()} function prints the full model summary. For
+conciseness, we will use the \texttt{fixef()} function which prints the
+regression coefficients of the constant (aka fixed) terms. The full
+summary with an explanation of each part can be found in XXX. For each
+coefficient in the model, \texttt{fixef()} prints the name of the
+coefficient, the mean estimate, the estimate error and the lower and
+upper limits of a Bayesian Credible interval (CrI). Here, we print an
+80\% CrI.\footnote{There is nothing special about 95\% within Bayesian
+  inference and instead experts recommend to check and report a variety
+  of CrIs. For simplicity, we will use 80\% CrIs in this case study.} In
+Bayesian inference, an 80\% CrI indicates the range of values within
+which the true estimate falls at 80\% probability or confidence. We
+round all values to the nearest 2 digits for clarity.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{round}\NormalTok{(}
+  \FunctionTok{fixef}\NormalTok{(voi\_prop\_bm, }\AttributeTok{prob =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{)),}
+  \AttributeTok{digits =} \DecValTok{2}
+\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+              Estimate Est.Error   Q10   Q90
+speech_rate_c     0.08      0.06  0.01  0.15
+voweli:c2k       -0.91      0.14 -1.08 -0.74
+vowele:c2k       -1.08      0.11 -1.22 -0.94
+vowela:c2k       -0.99      0.12 -1.14 -0.84
+vowelo:c2k       -0.79      0.14 -0.96 -0.62
+vowelu:c2k       -1.00      0.16 -1.20 -0.80
+voweli:c2t       -0.66      0.11 -0.79 -0.53
+vowele:c2t       -0.84      0.14 -1.02 -0.66
+vowela:c2t       -1.43      0.13 -1.60 -1.26
+vowelo:c2t       -1.15      0.13 -1.31 -0.99
+vowelu:c2t       -0.68      0.12 -0.83 -0.54
+voweli:c2p       -0.68      0.11 -0.81 -0.54
+vowele:c2p       -0.88      0.15 -1.07 -0.68
+vowela:c2p       -1.14      0.13 -1.31 -0.98
+vowelo:c2p       -0.66      0.11 -0.80 -0.53
+vowelu:c2p       -0.44      0.12 -0.59 -0.28
+\end{verbatim}
+
+The coefficients of a beta regression are estimated on the log-odds
+scale, as in Bernoulli/binomial (aka logistic) regressions. From the
+summary, we see that speech rate (number of syllables per second) has a
+positive effect on voicing proportion: the 80\% CrI is between 0.01 and
+0.15 log-odds {[}\(\beta\) = 0.8, SD = 0.06{]}. Log-odds can be
+converted to odd ratios by exponentiating the value: 0.01-0.15 log odds
+correspond to an odd ratio of 1.01 to 1.16, or as percentages, to an
+increase of voicing of 1 to 16\% for every increase of one syllable per
+second. Since this is an 80\% CrI, we can be 80\% confident that the
+true effect of speech rate is between 1-16\% increase of voicing
+proportion.
+
+Turning now to the coefficients for vowel and C2, given the indexing
+approach of coding these variables the model summary and the output of
+\texttt{fixef()} reports the \emph{predictions} in log-odds for each
+combination of vowel and C2, rather than differences between levels. The
+CrIs of the vowel/C2 coefficients span all negative log-odds values:
+these correspond to proportions that are lower than 0.5 (which is 0 in
+log-odds). This matches the general trends in the raw data, which we
+plotted in Figure~\ref{fig-ita-egg}.
+
+Next, we will plot the predicted proportions of each vowel/C2
+combination at mean speech rate (i.e.~centred speech rate = 0) and then
+calculate the average pair-wise difference in voicing proportion between
+/k, t, p/. Finally, we will assess whether there is greater
+between-vowel variability in /t, p/ relative to /k/.
+
+Before being able to plot the predictions, it's important to get
+familiar with the so-called posterior draws. {[}Bayesian MCMC XXX{]}.
+Posterior draws can be conveniently obtained with the
+\texttt{as\_draws\_df()}. For the moment, we will extract only the draws
+of the constant regression coefficients (model variables starting with
+\texttt{b\_}). To check which coefficients are available in a model, use
+\texttt{get\_variables()} from the tidybayes package.
+\texttt{as\_draws\_df()} returns a tibble where each column contains the
+drawn values of a coefficient. The probability distribution of the drawn
+values of each coefficient is the posterior probability distribution of
+that coefficient. Note that, due to the indexing coding of vowel and C2,
+all coefficient except \texttt{b\_speech\_rate\_c} are \emph{predicted
+log-odds} for each vowel/C2 combination (the drawn values for
+\texttt{b\_speech\_rate\_c} are drawn \emph{differences} in log-odds for
+each unit increase of speech rate). The drawn values are in log-odds,
+but we can convert them to proportions with \texttt{plogis()} (we will
+do this when plotting below).
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# extract only coefficient variables starting with "b\_"}
+\NormalTok{voi\_prop\_bm\_draws }\OtherTok{\textless{}{-}} \FunctionTok{as\_draws\_df}\NormalTok{(voi\_prop\_bm, }\AttributeTok{variable =} \StringTok{"\^{}b\_"}\NormalTok{, }\AttributeTok{regex =} \ConstantTok{TRUE}\NormalTok{)}
+\NormalTok{voi\_prop\_bm\_draws}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+# A draws_df: 1000 iterations, 4 chains, and 16 variables
+   b_speech_rate_c b_voweli:c2k b_vowele:c2k b_vowela:c2k b_vowelo:c2k
+1            0.159        -0.72        -1.05        -1.12        -1.02
+2            0.035        -0.75        -1.39        -1.06        -0.89
+3            0.126        -0.85        -0.98        -1.10        -0.64
+4            0.064        -0.70        -1.07        -1.05        -0.83
+5            0.032        -0.77        -1.18        -0.94        -0.70
+6            0.073        -0.85        -1.23        -1.06        -0.77
+7            0.101        -0.95        -1.07        -1.14        -0.86
+8            0.074        -1.02        -1.07        -1.17        -0.83
+9            0.087        -1.04        -1.07        -1.26        -0.79
+10           0.092        -1.20        -1.04        -0.92        -0.64
+   b_vowelu:c2k b_voweli:c2t b_vowele:c2t
+1         -1.05        -0.62        -0.80
+2         -1.04        -0.67        -0.99
+3         -0.91        -0.66        -0.80
+4         -1.07        -0.53        -0.87
+5         -1.06        -0.63        -0.87
+6         -1.00        -0.69        -0.73
+7         -1.22        -0.74        -1.08
+8         -1.22        -0.73        -1.04
+9         -1.14        -0.66        -0.91
+10        -0.84        -0.63        -0.61
+# ... with 3990 more draws, and 8 more variables
+# ... hidden reserved variables {'.chain', '.iteration', '.draw'}
+\end{verbatim}
+
+We can now wrangle this tibble and plot the posterior distributions for
+each vowel/C2 combination.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{voi\_prop\_bm\_draws\_long }\OtherTok{\textless{}{-}}\NormalTok{ voi\_prop\_bm\_draws }\SpecialCharTok{|\textgreater{}} 
+  \CommentTok{\# drop b\_speech\_rate\_c before pivoting}
+  \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{b\_speech\_rate\_c) }\SpecialCharTok{|\textgreater{}} 
+  \CommentTok{\# pivot vowel:c2 columns}
+  \FunctionTok{pivot\_longer}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{b\_voweli:c2k}\StringTok{\textasciigrave{}}\SpecialCharTok{:}\StringTok{\textasciigrave{}}\AttributeTok{b\_vowelu:c2p}\StringTok{\textasciigrave{}}\NormalTok{, }\AttributeTok{names\_to =} \StringTok{"coeff"}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
+  \CommentTok{\# separate "coeff" labels into type ("b"), vowel and c2}
+  \FunctionTok{separate}\NormalTok{(coeff, }\AttributeTok{into =} \FunctionTok{c}\NormalTok{(}\StringTok{"type"}\NormalTok{, }\StringTok{"vowel"}\NormalTok{, }\StringTok{"c2"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+Warning: Dropping 'draws_df' class as required metadata was removed.
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{voi\_prop\_bm\_draws\_long}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+# A tibble: 60,000 x 7
+   .chain .iteration .draw type  vowel  c2     value
+    <int>      <int> <int> <chr> <chr>  <chr>  <dbl>
+ 1      1          1     1 b     voweli c2k   -0.721
+ 2      1          1     1 b     vowele c2k   -1.05 
+ 3      1          1     1 b     vowela c2k   -1.12 
+ 4      1          1     1 b     vowelo c2k   -1.02 
+ 5      1          1     1 b     vowelu c2k   -1.05 
+ 6      1          1     1 b     voweli c2t   -0.625
+ 7      1          1     1 b     vowele c2t   -0.795
+ 8      1          1     1 b     vowela c2t   -1.40 
+ 9      1          1     1 b     vowelo c2t   -1.10 
+10      1          1     1 b     vowelu c2t   -0.668
+# i 59,990 more rows
+\end{verbatim}
+
+For plotting, we can use ggplot2 statistics layers from the ggdist
+package. \texttt{stat\_halfeye()} plots the density of the posterior
+probability (in grey), its median (point) and CrIs (lines). Let's use 60
+and 80\% CrIs and transform the log-odds values to proportions with
+\texttt{plogis()}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# attach ggdist package}
+\FunctionTok{library}\NormalTok{(ggdist)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+
+Attaching package: 'ggdist'
+\end{verbatim}
+
+\begin{verbatim}
+The following objects are masked from 'package:brms':
+
+    dstudent_t, pstudent_t, qstudent_t, rstudent_t
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{voi\_prop\_bm\_draws\_long }\SpecialCharTok{|\textgreater{}} 
+  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\FunctionTok{plogis}\NormalTok{(value), vowel)) }\SpecialCharTok{+}
+  \FunctionTok{stat\_halfeye}\NormalTok{(}\AttributeTok{.width =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.6}\NormalTok{, }\FloatTok{0.8}\NormalTok{)) }\SpecialCharTok{+}
+  \FunctionTok{facet\_grid}\NormalTok{(}\AttributeTok{rows =} \FunctionTok{vars}\NormalTok{(c2))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics{manuscript_files/figure-pdf/fig-voi-prop-bm-1-1.pdf}
+
+}
+
+\caption{\label{fig-voi-prop-bm-1}}
+
+\end{figure}%
+
+What if we want to plot the average predicted voicing proportion for the
+three consonants /k, t, p/? One approach is to take the mean across
+vowels within each consonant for each posterior draw, and the posterior
+distribution of the resulting list of values is the predicted posterior
+distribution of voicing proportion for each consonant, assuming an
+``average'' vowel.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{voi\_prop\_bm\_draws\_long\_c2 }\OtherTok{\textless{}{-}}\NormalTok{ voi\_prop\_bm\_draws\_long }\SpecialCharTok{|\textgreater{}} 
+  \CommentTok{\# grouping by .draw and c2 ensures that averaging applies only within draw and c2}
+  \FunctionTok{group\_by}\NormalTok{(.draw, c2) }\SpecialCharTok{|\textgreater{}} 
+  \CommentTok{\# calculate the mean value within draw/c2}
+  \FunctionTok{summarise}\NormalTok{(}
+    \AttributeTok{value\_mean =} \FunctionTok{mean}\NormalTok{(value), }\AttributeTok{.groups =} \StringTok{"drop"}
+\NormalTok{  )}
+
+\NormalTok{voi\_prop\_bm\_draws\_long\_c2 }\SpecialCharTok{|\textgreater{}} 
+  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\FunctionTok{plogis}\NormalTok{(value\_mean), c2)) }\SpecialCharTok{+}
+  \FunctionTok{stat\_halfeye}\NormalTok{(}\AttributeTok{.width =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.6}\NormalTok{, }\FloatTok{0.8}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{manuscript_files/figure-pdf/voi-prop-bm-draws-long-c2-1.pdf}
+
+Based on the posterior distributions of the mean voicing proportion by
+consonant, /p/ has a somewhat higher voicing proportion than /k/ and
+/t/. The real question is: how much higher? We can quantify this by
+taking the difference of the drawn values for /p/ and those for /t, k/.
+Since we want to compare /t, k/ to /p/, we should first average the
+draws of /t, k/ and then take the difference of the averaged draws and
+the draws of /p/.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{voi\_prop\_bm\_diff }\OtherTok{\textless{}{-}}\NormalTok{ voi\_prop\_bm\_draws\_long\_c2 }\SpecialCharTok{|\textgreater{}} 
+  \CommentTok{\# pivot data to create one column per consonant with the mean drawn values,}
+  \CommentTok{\# with one draw per raw}
+  \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ c2, }\AttributeTok{values\_from =}\NormalTok{ value\_mean) }\SpecialCharTok{|\textgreater{}} 
+  \FunctionTok{mutate}\NormalTok{(}
+    \CommentTok{\# calculate the mean of /k/ and /t/, for each draw}
+    \AttributeTok{c2tk =} \FunctionTok{mean}\NormalTok{(}\FunctionTok{c}\NormalTok{(c2k, c2t)),}
+    \CommentTok{\# calculate the difference of /p/ and /t, k/}
+    \AttributeTok{c2p\_tk\_diff =}\NormalTok{ c2p }\SpecialCharTok{{-}}\NormalTok{ c2tk}
+\NormalTok{  )}
+
+\NormalTok{voi\_prop\_bm\_diff}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+# A tibble: 4,000 x 6
+   .draw    c2k    c2p    c2t   c2tk c2p_tk_diff
+   <int>  <dbl>  <dbl>  <dbl>  <dbl>       <dbl>
+ 1     1 -0.992 -0.745 -0.917 -0.953      0.208 
+ 2     2 -1.02  -0.790 -1.03  -0.953      0.163 
+ 3     3 -0.895 -0.687 -0.890 -0.953      0.266 
+ 4     4 -0.944 -0.814 -0.910 -0.953      0.140 
+ 5     5 -0.928 -0.791 -0.920 -0.953      0.162 
+ 6     6 -0.983 -0.728 -0.926 -0.953      0.225 
+ 7     7 -1.05  -0.873 -1.08  -0.953      0.0802
+ 8     8 -1.06  -0.875 -1.07  -0.953      0.0782
+ 9     9 -1.06  -0.974 -1.05  -0.953     -0.0206
+10    10 -0.929 -0.829 -0.953 -0.953      0.125 
+# i 3,990 more rows
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{voi\_prop\_bm\_diff }\SpecialCharTok{|\textgreater{}} 
+  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(c2p\_tk\_diff)) }\SpecialCharTok{+}
+  \FunctionTok{stat\_halfeye}\NormalTok{(}\AttributeTok{.width =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.6}\NormalTok{, }\FloatTok{0.8}\NormalTok{, }\FloatTok{0.9}\NormalTok{)) }\SpecialCharTok{+}
+  \FunctionTok{geom\_vline}\NormalTok{(}\AttributeTok{xintercept =} \DecValTok{0}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics{manuscript_files/figure-pdf/fig-diff-p-tk-1.pdf}
+
+}
+
+\caption{\label{fig-diff-p-tk}}
+
+\end{figure}%
+
+Once we have the posterior difference, we can obtain CrIs of the
+difference using \texttt{quantile2()} from the posterior package. Beware
+that the values of the difference are in log-odds! We can convert these
+into odd ratios with \texttt{exp()}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{library}\NormalTok{(posterior)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+This is posterior version 1.6.0
+\end{verbatim}
+
+\begin{verbatim}
+
+Attaching package: 'posterior'
+\end{verbatim}
+
+\begin{verbatim}
+The following objects are masked from 'package:stats':
+
+    mad, sd, var
+\end{verbatim}
+
+\begin{verbatim}
+The following objects are masked from 'package:base':
+
+    %in%, match
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{voi\_prop\_bm\_diff }\SpecialCharTok{|\textgreater{}} 
+  \FunctionTok{reframe}\NormalTok{(}
+    \CommentTok{\# 90\% CrI}
+    \AttributeTok{q90 =} \FunctionTok{quantile2}\NormalTok{(c2p\_tk\_diff, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.05}\NormalTok{, }\FloatTok{0.95}\NormalTok{)),}
+    \CommentTok{\# 80\% CrI}
+    \AttributeTok{q80 =} \FunctionTok{quantile2}\NormalTok{(c2p\_tk\_diff, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{)),}
+    \CommentTok{\# 60\% CrI}
+    \AttributeTok{q60 =} \FunctionTok{quantile2}\NormalTok{(c2p\_tk\_diff, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.2}\NormalTok{, }\FloatTok{0.8}\NormalTok{)),}
+\NormalTok{  ) }\SpecialCharTok{|\textgreater{}} 
+  \CommentTok{\# exponentiate to get odd ratios and round to 2 digits}
+  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{everything}\NormalTok{(), }\SpecialCharTok{\textasciitilde{}}\FunctionTok{round}\NormalTok{(}\FunctionTok{exp}\NormalTok{(.x), }\DecValTok{2}\NormalTok{)))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+# A tibble: 2 x 3
+    q90   q80   q60
+  <dbl> <dbl> <dbl>
+1  1.07  1.1   1.14
+2  1.38  1.34  1.3 
+\end{verbatim}
+
+So, based on the model and data, there is a 90\% probability that the
+voicing proportion in /p/ is 1.07-1.38 times longer (or 7-38\% increase)
+than in /t, k/. At 80\% confidence, the change ratio is 1.10-1.34 (or
+10-34\% increase) while at 60\% confidence is 1.14-1.30 (14-30\%
+increase). In other words we can be quite confident that the voicing
+proportion in /p/ is longer than in /t, k/ and that the increase is less
+than 35\%.
+
+brms comes with a convenient function, \texttt{conditional\_effects()},
+to plot posterior means and CrIs based on predictors in the model. In
+the following example, we plot the predicted proportion of voicing by
+consonant and vowel.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{conditional\_effects}\NormalTok{(voi\_prop\_bm, }\StringTok{"c2:vowel"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\includegraphics{manuscript_files/figure-pdf/voi-prop-bm-cond-1.pdf}
+
+Finally, the package marginaleffects has two other convenience functions
+that return CrIs of comparisons across predictor levels
+(\texttt{avg\_comparisons()}) and CrIs of posterior predictions across
+predictor levels (\texttt{avg\_predictions}). {[}XXX{]}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{library}\NormalTok{(marginaleffects)}
+
+\FunctionTok{avg\_comparisons}\NormalTok{(voi\_prop\_bm, }\AttributeTok{variables =} \FunctionTok{list}\NormalTok{(}\AttributeTok{c2 =} \StringTok{"pairwise"}\NormalTok{), }\AttributeTok{conf\_level =} \FloatTok{0.8}\NormalTok{, }\AttributeTok{type =} \StringTok{"link"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+
+          Contrast Estimate   10.0 % 90.0 %
+ mean(t) - mean(k)   0.0348 -0.00418 0.0738
+ mean(p) - mean(k)   0.2176  0.17837 0.2558
+ mean(p) - mean(t)   0.1829  0.14474 0.2192
+
+Term: c2
+Type:  link 
+Columns: term, contrast, estimate, conf.low, conf.high, predicted_lo, predicted_hi, predicted, tmp_idx 
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{avg\_predictions}\NormalTok{(voi\_prop\_bm, }\AttributeTok{variables =} \StringTok{"vowel"}\NormalTok{, }\AttributeTok{conf\_level =} \FloatTok{0.8}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+
+ vowel Estimate 10.0 % 90.0 %
+     i    0.331  0.325  0.338
+     e    0.300  0.293  0.307
+     a    0.245  0.238  0.252
+     o    0.305  0.298  0.312
+     u    0.348  0.341  0.354
+
+Type:  response 
+Columns: vowel, estimate, conf.low, conf.high 
+\end{verbatim}
+
+\renewcommand\refname{Case study 2: pre-nasalisation in Greek}
   \bibliography{bibliography.bib}

### 2024-12-06 manuscript/manuscript.qmd
edits based on Paul

@@
 We will use brms to fit Bayesian beta regressions. [XXX why Bayesian]. 
-The model has voicing proportion as the outcome variable and the following terms: an interaction between vowel (/i, e, a, ɔ, u/) and second consonant C2 (/k, t, p/), centred speech rate (number of syllables per second); as varying (aka random) terms, by-speaker varying coefficients for the vowel/consonant interaction and for centred speech rate.^[Footnote about Gelman's terminology for random effects.] The categorical predictors vowel and C2 are coded using indexing rather than traditional R contrasts (in R, this corresponds to suppressing the model's intercept with the `0 +` syntax; using indexing instead of contrasts does not impact the model posterior predictions, but it makes it easier to specify priors when users do so).
+The model has voicing proportion as the outcome variable and the following terms: an interaction between vowel (/i, e, a, ɔ, u/) and second consonant C2 (/k, t, p/), centred speech rate (number of syllables per second); as varying (aka random) terms, by-speaker varying coefficients for the vowel/consonant interaction and for centred speech rate.^[Footnote about Gelman's terminology for random effects.] The categorical predictors vowel and C2 are coded using indexing rather than traditional R contrasts: in R, this corresponds to suppressing the model's intercept with the `0 +` syntax; using indexing instead of contrasts makes it easier to specify priors.
 For pedagogical simplicity, the model will use the default priors, but note that in real data analyses contexts, priors should be specified by the user.
@@
 The `summary()` function prints the full model summary.
-For conciseness, we will use the `fixef()` function which prints the regression coefficients of the constant (aka fixed) terms.
+For conciseness, we will use the `fixef()` function which prints the regression coefficients.
 The full summary with an explanation of each part can be found in XXX.
 For each coefficient in the model, `fixef()` prints the name of the coefficient, the mean estimate, the estimate error and the lower and upper limits of a Bayesian Credible interval (CrI).
-Here, we print an 80% CrI.^[There is nothing special about 95% within Bayesian inference and instead experts recommend to check and report a variety of CrIs.
-For simplicity, we will use 80% CrIs in this case study.] In Bayesian inference, an 80% CrI indicates the range of values within which the true estimate falls at 80% probability or confidence.
+Here, we print an 80% CrI.
+There is nothing special about 95% CrI within Bayesian inference and instead experts recommend to check and report a variety of CrIs.
+Bayesian CrIs indicate that at a certain probability levels the "true" estimate lies within that interval: so, for example, a 90% CrI [A, B] indicates that there is a 90% probability that the "true" estimate is between A and B.
+Different probability levels correspond to different levels of confidence: the higher the probability the higher the confidence (always conditional on data and model).
+Obtaining CrIs at different probability levels allows researchers to make more fine-grained inferential statements than the frequentist significance dichotomy affords.
+For simplicity of exposition, we will use 80% CrIs in this case study but we strongly recommend researchers to always obtain CrIs at different levels of probability and base their inferences on all and not one in particular.
+To reiterate, in Bayesian inference, an 80% CrI indicates the range of values within which the true estimate falls at 80% probability or confidence.
 We round all values to the nearest 2 digits for clarity.
@@
 Log-odds can be converted to odd ratios by exponentiating the value: 0.01-0.15 log odds correspond to an odd ratio of 1.01 to 1.16, or as percentages, to an increase of voicing of 1 to 16% for every increase of one syllable per second.
-Since this is an 80% CrI, we can be 80% confident that the true effect of speech rate is between 1-16% increase of voicing proportion.
+Since this is an 80% CrI, we can be 80% confident that the true effect of speech rate is between 1-16% increase of voicing proportion, conditional on the data and model.
+Note that transforming measures this way is appropriate *only* with quantile-based measures (like CrIs) but not with moments like the mean and standard deviation: to correctly get mean and SDs in the transformed scale, you must first extract the posterior draws (see below), convert them and then take moments such as mean and SD (for a more detailed explanation, see XXX).
 

### 2024-12-06 manuscript/manuscript.tex
edits based on Paul

@@
 The categorical predictors vowel and C2 are coded using indexing rather
-than traditional R contrasts (in R, this corresponds to suppressing the
+than traditional R contrasts: in R, this corresponds to suppressing the
 model's intercept with the \texttt{0\ +} syntax; using indexing instead
-of contrasts does not impact the model posterior predictions, but it
-makes it easier to specify priors when users do so). For pedagogical
+of contrasts makes it easier to specify priors. For pedagogical
 simplicity, the model will use the default priors, but note that in real
@@
 conciseness, we will use the \texttt{fixef()} function which prints the
-regression coefficients of the constant (aka fixed) terms. The full
-summary with an explanation of each part can be found in XXX. For each
-coefficient in the model, \texttt{fixef()} prints the name of the
-coefficient, the mean estimate, the estimate error and the lower and
-upper limits of a Bayesian Credible interval (CrI). Here, we print an
-80\% CrI.\footnote{There is nothing special about 95\% within Bayesian
-  inference and instead experts recommend to check and report a variety
-  of CrIs. For simplicity, we will use 80\% CrIs in this case study.} In
-Bayesian inference, an 80\% CrI indicates the range of values within
-which the true estimate falls at 80\% probability or confidence. We
-round all values to the nearest 2 digits for clarity.
+regression coefficients. The full summary with an explanation of each
+part can be found in XXX. For each coefficient in the model,
+\texttt{fixef()} prints the name of the coefficient, the mean estimate,
+the estimate error and the lower and upper limits of a Bayesian Credible
+interval (CrI). Here, we print an 80\% CrI. There is nothing special
+about 95\% CrI within Bayesian inference and instead experts recommend
+to check and report a variety of CrIs. Bayesian CrIs indicate that at a
+certain probability levels the ``true'' estimate lies within that
+interval: so, for example, a 90\% CrI {[}A, B{]} indicates that there is
+a 90\% probability that the ``true'' estimate is between A and B.
+Different probability levels correspond to different levels of
+confidence: the higher the probability the higher the confidence (always
+conditional on data and model). Obtaining CrIs at different probability
+levels allows researchers to make more fine-grained inferential
+statements than the frequentist significance dichotomy affords. For
+simplicity of exposition, we will use 80\% CrIs in this case study but
+we strongly recommend researchers to always obtain CrIs at different
+levels of probability and base their inferences on all and not one in
+particular. To reiterate, in Bayesian inference, an 80\% CrI indicates
+the range of values within which the true estimate falls at 80\%
+probability or confidence. We round all values to the nearest 2 digits
+for clarity.
 
@@
 true effect of speech rate is between 1-16\% increase of voicing
-proportion.
+proportion, conditional on the data and model. Note that transforming
+measures this way is appropriate \emph{only} with quantile-based
+measures (like CrIs) but not with moments like the mean and standard
+deviation: to correctly get mean and SDs in the transformed scale, you
+must first extract the posterior draws (see below), convert them and
+then take moments such as mean and SD (for a more detailed explanation,
+see XXX).
 

### 2024-12-08 code/01_data_checks.qmd
add second case study

@@
 ---
+
+## Voicing within closure
 
@@
 ```{r}
-#| label: load-data
+#| label: load-ita-egg
 #| 
@@
 ```
+
+## Vowel nasalisation
+
+```{r}
+#| label: read-nasal
+nasal <- read_csv("https://raw.githubusercontent.com/ChristopherCarignan/journal-articles/refs/heads/master/rtMRI-velum/velum_data.csv") |> 
+  filter(
+    stress == "N",
+    vowel %in% c("a_", "E_", "I_", "O_", "U")
+  ) |> 
+  mutate(
+    voicing = case_when(
+      post %in% c("nt__", "nt_@", "nt_6", "nt_a") ~ "voiceless",
+      post %in% c("nd_@", "nd_6", "nd_a") ~ "voiced"
+    )
+  ) |> 
+  # drop codas not included in the analysis
+  drop_na(voicing) |> 
+  mutate(
+    nas_dur = (Vokal_off - velumopening_maxvel_on) * 1000,
+    nas_prop = nas_dur / (Vokal_dur * 1000)
+  ) |> 
+  # drop observations with wrong fMRI tracking
+  filter(nas_prop > 0, nas_prop < 1) |> 
+  select(
+    speaker, label, vowel, post, voicing, nas_prop
+  )
+
+write_csv(nasal, "data/carignan2021/nasal.csv")
+```
+
+```{r}
+nasal |> 
+  ggplot(aes(nas_prop, fill = voicing)) +
+  geom_density(alpha = 0.5)
+```
+
+```{r}
+nasal |> 
+  ggplot(aes(nas_prop, fill = voicing)) +
+  geom_density(alpha = 0.5) +
+  facet_wrap(~vowel)
+```

### 2024-12-08 code/01_data_checks.qmd
edit second case study

--- 2024-11-25 code/01_data_checks.qmd
@@
 ```
+
+## Vowel nasalisation
+
+```{r}
+#| label: read-nasal
+nasal <- read_csv("https://raw.githubusercontent.com/ChristopherCarignan/journal-articles/refs/heads/master/rtMRI-velum/velum_data.csv") |> 
+  filter(
+    stress == "N",
+    vowel %in% c("a_", "E_", "I_", "O_", "U_")
+  ) |> 
+  mutate(
+    voicing = case_when(
+      post %in% c("nt__", "nt_@", "nt_6", "nt_a") ~ "voiceless",
+      post %in% c("nd_@", "nd_6", "nd_a") ~ "voiced"
+    ),
+    vowel = str_to_lower(vowel) |> str_remove("_")
+  ) |> 
+  # drop codas not included in the analysis
+  drop_na(voicing) |> 
+  mutate(
+    nas_dur = (Vokal_off - velumopening_maxvel_on) * 1000,
+    nas_prop = nas_dur / (Vokal_dur * 1000),
+    NC = ifelse(voicing == "voiceless", "nt", "nd")
+  ) |> 
+  # drop observations with wrong fMRI tracking
+  filter(nas_prop > 0, nas_prop < 1) |> 
+  select(
+    speaker, label, vowel, NC, voicing, nas_prop
+  )
+
+write_csv(nasal, "data/carignan2021/nasal.csv")
+```
+
+```{r}
+nasal |> 
+  ggplot(aes(nas_prop, fill = voicing)) +
+  geom_density(alpha = 0.5)
+```
+
+```{r}
+nasal |> 
+  ggplot(aes(nas_prop, fill = voicing)) +
+  geom_density(alpha = 0.5) +
+  facet_wrap(~vowel)
+```

### 2024-12-08 manuscript/manuscript.qmd
add second case study
edit second case study

@@
 date: last-modified
-bibliography: bibliography.bib
+bibliography: linguistics.bib
 format:
@@
 
-Regression models are very common, but there is a tendency of using Gaussian distribution families (i.e. probability distributions for the outcome variable) for anything that is numeric.
+Regression models are very common, but there is a tendency of using Gaussian distribution families (i.e.
+probability distributions for the outcome variable) for anything that is numeric.
 A possible reason is that the base R function for fitting regression models, `lm()`, and the lme4 function used to fit regression models with varying terms, `lmer()`, both fit Gaussian regressions by default and the user does not have to specify the distribution family.
@@
 We will use a beta regression model to assess these patterns.
-["expectations" XXX]
+\["expectations" XXX\]
 
@@
 
-We will use brms to fit Bayesian beta regressions. [XXX why Bayesian]. 
-The model has voicing proportion as the outcome variable and the following terms: an interaction between vowel (/i, e, a, ɔ, u/) and second consonant C2 (/k, t, p/), centred speech rate (number of syllables per second); as varying (aka random) terms, by-speaker varying coefficients for the vowel/consonant interaction and for centred speech rate.^[Footnote about Gelman's terminology for random effects.] The categorical predictors vowel and C2 are coded using indexing rather than traditional R contrasts: in R, this corresponds to suppressing the model's intercept with the `0 +` syntax; using indexing instead of contrasts makes it easier to specify priors.
+We will use brms to fit Bayesian beta regressions.
+\[XXX why Bayesian\].
+The model has voicing proportion as the outcome variable and the following terms: an interaction between vowel (/i, e, a, ɔ, u/) and second consonant C2 (/k, t, p/), centred speech rate (number of syllables per second); as varying (aka random) terms, by-speaker varying coefficients for the vowel/consonant interaction and for centred speech rate.[^1] The categorical predictors vowel and C2 are coded using indexing rather than traditional R contrasts: in R, this corresponds to suppressing the model's intercept with the `0 +` syntax; using indexing instead of contrasts makes it easier to specify priors.
 For pedagogical simplicity, the model will use the default priors, but note that in real data analyses contexts, priors should be specified by the user.
 I refer the readers to XXX.
+
+[^1]: Footnote about Gelman's terminology for random effects.
 
@@
 There is nothing special about 95% CrI within Bayesian inference and instead experts recommend to check and report a variety of CrIs.
-Bayesian CrIs indicate that at a certain probability levels the "true" estimate lies within that interval: so, for example, a 90% CrI [A, B] indicates that there is a 90% probability that the "true" estimate is between A and B.
+Bayesian CrIs indicate that at a certain probability levels the "true" estimate lies within that interval: so, for example, a 90% CrI \[A, B\] indicates that there is a 90% probability that the "true" estimate is between A and B.
 Different probability levels correspond to different levels of confidence: the higher the probability the higher the confidence (always conditional on data and model).
@@
 The coefficients of a beta regression are estimated on the log-odds scale, as in Bernoulli/binomial (aka logistic) regressions.
-From the summary, we see that speech rate (number of syllables per second) has a positive effect on voicing proportion: the 80% CrI is between 0.01 and 0.15 log-odds [$\beta$ = 0.8, SD = 0.06].
+From the summary, we see that speech rate (number of syllables per second) has a positive effect on voicing proportion: the 80% CrI is between 0.01 and 0.15 log-odds \[$\beta$ = 0.8, SD = 0.06\].
 Log-odds can be converted to odd ratios by exponentiating the value: 0.01-0.15 log odds correspond to an odd ratio of 1.01 to 1.16, or as percentages, to an increase of voicing of 1 to 16% for every increase of one syllable per second.
@@
 Before being able to plot the predictions, it's important to get familiar with the so-called posterior draws.
-[Bayesian MCMC XXX].
+\[Bayesian MCMC XXX\].
 Posterior draws can be conveniently obtained with the `as_draws_df()`.
@@
 Finally, the package marginaleffects has two other convenience functions that return CrIs of comparisons across predictor levels (`avg_comparisons()`) and CrIs of posterior predictions across predictor levels (`avg_predictions`).
-[XXX]
+\[XXX\]
 
@@
 
-# Case study 2: pre-nasalisation in Greek
+# Case study 2: coarticulatory vowel nasalisation
+
+For the second case study we will use data from @carignan2021.
+The study looked at properties of nasality in German VNC sequences.
+Here, we will focus on the effect of C voicing (voiceless /t/ vs voiced /d/) on the proportion of nasalisation within the vowel in the VNC sequence.
+Previous work on coarticulatory nasalisation in English has suggested that vowels followed by an NC sequence where C is voiceless (NT) should show earlier coarticulatory nasalisation than vowels followed by an NC sequence where C is voiced [ND, see review in @carignan2021].
+This pattern has been suggested to be driven by the articulatory and acoustic incompatibility of voicelessness and nasalisation, by which the velum opening gesture of the nasal consonant is pushed away (i.e.
+earlier) when the consonant following the nasal is voiceless.
+Everything else being equal, a greater proportion of vowel nasalisation (from the perspective of time) should be found in vowels followed by NT than in vowels followed by ND.
+
+We will model the proportion of coarticulatory nasalisation in the German short vowels /i, e, a, o, u/ when followed by /nt/ or /nd/, using a Bayesian beta regression model.
+The proportion was calculated as the proportion of the nasal interval to the duration of the vowel.
+The nasal interval was defined thus: the interval between the time of peak velocity of velum opening to the offset of the vowel.
+
+First, let's read and plot the data.
+
+```{r}
+#| label: nasal
+nasal <- read_csv("data/carignan2021/nasal.csv")
+nasal
+```
+
+-   `speaker` indicates the speaker ID.
+
+-   `label` is the word label as given in the original data.
+
+-   `vowel` is the target vowel in the VNC sequence.
+
+-   `NC` is the NC sequence.
+
+-   `voicing` indicates the voicing of C.
+
+-   `nas_prop` is the proportion of coarticulatory nasalisation of the vowel.
+
+```{r}
+#| label: fig-nasal
+#| fig-cap: "Proportion of coarticulatory nasalisation during the vowel in VNC sequences in German, depending on C voicing."
+#| echo: false
+
+nasal |> 
+  ggplot(aes(vowel, nas_prop, colour = voicing)) +
+  geom_jitter(position = position_jitterdodge(jitter.width = 0.2), alpha = 0.8) +
+  scale_color_brewer(type = "qual") +
+  labs(x = "Vowel", y = "Nasalisation proportion", colour = "C voicing")
+
+```
+
+@fig-nasal shows the proportion of coarticulatory nasalisation in vowels followed by /nd/ (voiced) vs /nt/ (voiceless) sequences, for the short vowels /i, e, a, o, u/.
+We can see a pattern of higher nasalisation proportion in vowels followed by /nt/, at least in the vowels /a, i, o/.
+For /e, u/, the distribution of nasalisation proportion seems to be similar between the voiced and voiceless contexts.
+
+Now onto modelling with a beta regression.
+Note that a full appropriate model would include further predictors (both constant and varying), but for simplicity here we include only the following predictors: voicing (voiced /nd/ vs voiceless /nt/), vowel (/i, e, a, o, u/), including an interaction between them.
+As varying terms, we include a varying intercept by speaker and a by-speaker varying slope for voicing and vowel in interaction.
+As with the model from the first case study, voicing and vowel are coded using indexing, by suppressing the intercept with `0 +`.
+Here's the code of the model.
+
+```{r}
+#| label: nas-prop-bm
+
+nas_prop_bm <- brm(
+  nas_prop ~ 0 + voicing:vowel + (0 + voicing:vowel | speaker),
+  data = nasal,
+  family = Beta,
+  cores = 4,
+  seed = 3749,
+  file = "data/cache/nas_prop_bm"
+)
+```

### 2024-12-08 manuscript/manuscript.tex
edit second case study

@@
 \makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\newsavebox\pandoc@box
+\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
+  \sbox\pandoc@box{#1}%
+  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
+  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
+  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
+  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
+  \else\usebox{\pandoc@box}%
+  \fi%
+}
 % Set default figure placement to htbp
-\makeatletter
 \def\fps@figure{htbp}
@@
 
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
 \usepackage[]{natbib}
@@
 
-\includegraphics{manuscript_files/figure-pdf/fig-ita-egg-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-ita-egg-1.pdf}}
 
@@
 
-\includegraphics{manuscript_files/figure-pdf/fig-voi-prop-bm-1-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-voi-prop-bm-1-1.pdf}}
 
@@
 
-\includegraphics{manuscript_files/figure-pdf/voi-prop-bm-draws-long-c2-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/voi-prop-bm-draws-long-c2-1.pdf}}
 
@@
 
-\includegraphics{manuscript_files/figure-pdf/fig-diff-p-tk-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-diff-p-tk-1.pdf}}
 
@@
 
-\includegraphics{manuscript_files/figure-pdf/voi-prop-bm-cond-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/voi-prop-bm-cond-1.pdf}}
 
@@
 
-\renewcommand\refname{Case study 2: pre-nasalisation in Greek}
-  \bibliography{bibliography.bib}
+\section{Case study 2: coarticulatory vowel
+nasalisation}\label{case-study-2-coarticulatory-vowel-nasalisation}
+
+For the second case study we will use data from \citet{carignan2021}.
+The study looked at properties of nasality in German VNC sequences.
+Here, we will focus on the effect of C voicing (voiceless /t/ vs voiced
+/d/) on the proportion of nasalisation within the vowel in the VNC
+sequence. Previous work on coarticulatory nasalisation in English has
+suggested that vowels followed by an NC sequence where C is voiceless
+should show earlier coarticulatory nasalisation than vowels followed by
+an NC sequence where C is voiced \citep[see review in][]{carignan2021}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{nasal }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"data/carignan2021/nasal.csv"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+Rows: 643 Columns: 6
+-- Column specification --------------------------------------------------------
+Delimiter: ","
+chr (5): speaker, label, vowel, post, voicing
+dbl (1): nas_prop
+
+i Use `spec()` to retrieve the full column specification for this data.
+i Specify the column types or set `show_col_types = FALSE` to quiet this message.
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{nas\_prop\_bm }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
+\NormalTok{  nas\_prop }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ voicing}\SpecialCharTok{:}\NormalTok{vowel }\SpecialCharTok{+}\NormalTok{ (}\DecValTok{0} \SpecialCharTok{+}\NormalTok{ voicing}\SpecialCharTok{:}\NormalTok{vowel }\SpecialCharTok{|}\NormalTok{ speaker),}
+  \AttributeTok{data =}\NormalTok{ nasal,}
+  \AttributeTok{family =}\NormalTok{ Beta,}
+  \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
+  \AttributeTok{seed =} \DecValTok{3749}\NormalTok{,}
+  \AttributeTok{file =} \StringTok{"data/cache/nas\_prop\_bm"}
+\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{summary}\NormalTok{(nas\_prop\_bm, }\AttributeTok{prob =} \FloatTok{0.8}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+ Family: beta 
+  Links: mu = logit; phi = identity 
+Formula: nas_prop ~ 0 + voicing:vowel + (0 + voicing:vowel | speaker) 
+   Data: nasal (Number of observations: 643) 
+  Draws: 4 chains, each with iter = 2000; warmup = 1000; thin = 1;
+         total post-warmup draws = 4000
+
+Multilevel Hyperparameters:
+~speaker (Number of levels: 35) 
+                                                       Estimate Est.Error
+sd(voicingvoiced:vowela_)                                  0.45      0.13
+sd(voicingvoiceless:vowela_)                               0.37      0.10
+sd(voicingvoiced:vowelE_)                                  0.56      0.18
+sd(voicingvoiceless:vowelE_)                               0.34      0.13
+sd(voicingvoiced:vowelI_)                                  0.71      0.18
+sd(voicingvoiceless:vowelI_)                               0.83      0.16
+sd(voicingvoiced:vowelO_)                                  0.36      0.22
+sd(voicingvoiceless:vowelO_)                               0.24      0.18
+cor(voicingvoiced:vowela,_voicingvoiceless:vowela_)        0.49      0.22
+cor(voicingvoiced:vowela,_voicingvoiced:vowelE_)           0.29      0.25
+cor(voicingvoiceless:vowela,_voicingvoiced:vowelE_)        0.16      0.25
+cor(voicingvoiced:vowela,_voicingvoiceless:vowelE_)        0.25      0.27
+cor(voicingvoiceless:vowela,_voicingvoiceless:vowelE_)     0.42      0.25
+cor(voicingvoiced:vowelE,_voicingvoiceless:vowelE_)        0.17      0.28
+cor(voicingvoiced:vowela,_voicingvoiced:vowelI_)           0.15      0.26
+cor(voicingvoiceless:vowela,_voicingvoiced:vowelI_)        0.17      0.26
+cor(voicingvoiced:vowelE,_voicingvoiced:vowelI_)           0.23      0.26
+cor(voicingvoiceless:vowelE,_voicingvoiced:vowelI_)        0.29      0.27
+cor(voicingvoiced:vowela,_voicingvoiceless:vowelI_)       -0.11      0.24
+cor(voicingvoiceless:vowela,_voicingvoiceless:vowelI_)     0.16      0.23
+cor(voicingvoiced:vowelE,_voicingvoiceless:vowelI_)        0.14      0.25
+cor(voicingvoiceless:vowelE,_voicingvoiceless:vowelI_)     0.33      0.25
+cor(voicingvoiced:vowelI,_voicingvoiceless:vowelI_)        0.34      0.23
+cor(voicingvoiced:vowela,_voicingvoiced:vowelO_)           0.20      0.31
+cor(voicingvoiceless:vowela,_voicingvoiced:vowelO_)        0.13      0.31
+cor(voicingvoiced:vowelE,_voicingvoiced:vowelO_)           0.10      0.31
+cor(voicingvoiceless:vowelE,_voicingvoiced:vowelO_)        0.07      0.31
+cor(voicingvoiced:vowelI,_voicingvoiced:vowelO_)           0.14      0.31
+cor(voicingvoiceless:vowelI,_voicingvoiced:vowelO_)       -0.06      0.29
+cor(voicingvoiced:vowela,_voicingvoiceless:vowelO_)        0.03      0.32
+cor(voicingvoiceless:vowela,_voicingvoiceless:vowelO_)    -0.05      0.32
+cor(voicingvoiced:vowelE,_voicingvoiceless:vowelO_)        0.09      0.32
+cor(voicingvoiceless:vowelE,_voicingvoiceless:vowelO_)     0.03      0.33
+cor(voicingvoiced:vowelI,_voicingvoiceless:vowelO_)        0.12      0.32
+cor(voicingvoiceless:vowelI,_voicingvoiceless:vowelO_)     0.03      0.31
+cor(voicingvoiced:vowelO,_voicingvoiceless:vowelO_)        0.08      0.33
+                                                       l-80% CI u-80% CI Rhat
+sd(voicingvoiced:vowela_)                                  0.29     0.62 1.00
+sd(voicingvoiceless:vowela_)                               0.25     0.50 1.00
+sd(voicingvoiced:vowelE_)                                  0.33     0.79 1.00
+sd(voicingvoiceless:vowelE_)                               0.18     0.50 1.00
+sd(voicingvoiced:vowelI_)                                  0.50     0.96 1.00
+sd(voicingvoiceless:vowelI_)                               0.64     1.04 1.00
+sd(voicingvoiced:vowelO_)                                  0.08     0.66 1.00
+sd(voicingvoiceless:vowelO_)                               0.04     0.49 1.00
+cor(voicingvoiced:vowela,_voicingvoiceless:vowela_)        0.20     0.76 1.00
+cor(voicingvoiced:vowela,_voicingvoiced:vowelE_)          -0.05     0.60 1.00
+cor(voicingvoiceless:vowela,_voicingvoiced:vowelE_)       -0.17     0.50 1.00
+cor(voicingvoiced:vowela,_voicingvoiceless:vowelE_)       -0.10     0.59 1.00
+cor(voicingvoiceless:vowela,_voicingvoiceless:vowelE_)     0.09     0.72 1.00
+cor(voicingvoiced:vowelE,_voicingvoiceless:vowelE_)       -0.22     0.53 1.00
+cor(voicingvoiced:vowela,_voicingvoiced:vowelI_)          -0.20     0.49 1.00
+cor(voicingvoiceless:vowela,_voicingvoiced:vowelI_)       -0.18     0.50 1.00
+cor(voicingvoiced:vowelE,_voicingvoiced:vowelI_)          -0.11     0.56 1.00
+cor(voicingvoiceless:vowelE,_voicingvoiced:vowelI_)       -0.07     0.62 1.00
+cor(voicingvoiced:vowela,_voicingvoiceless:vowelI_)       -0.42     0.20 1.00
+cor(voicingvoiceless:vowela,_voicingvoiceless:vowelI_)    -0.13     0.46 1.00
+cor(voicingvoiced:vowelE,_voicingvoiceless:vowelI_)       -0.19     0.44 1.00
+cor(voicingvoiceless:vowelE,_voicingvoiceless:vowelI_)     0.00     0.63 1.00
+cor(voicingvoiced:vowelI,_voicingvoiceless:vowelI_)        0.03     0.62 1.00
+cor(voicingvoiced:vowela,_voicingvoiced:vowelO_)          -0.22     0.59 1.00
+cor(voicingvoiceless:vowela,_voicingvoiced:vowelO_)       -0.27     0.53 1.00
+cor(voicingvoiced:vowelE,_voicingvoiced:vowelO_)          -0.33     0.50 1.00
+cor(voicingvoiceless:vowelE,_voicingvoiced:vowelO_)       -0.34     0.47 1.00
+cor(voicingvoiced:vowelI,_voicingvoiced:vowelO_)          -0.27     0.52 1.00
+cor(voicingvoiceless:vowelI,_voicingvoiced:vowelO_)       -0.44     0.33 1.00
+cor(voicingvoiced:vowela,_voicingvoiceless:vowelO_)       -0.40     0.45 1.00
+cor(voicingvoiceless:vowela,_voicingvoiceless:vowelO_)    -0.46     0.37 1.00
+cor(voicingvoiced:vowelE,_voicingvoiceless:vowelO_)       -0.34     0.51 1.00
+cor(voicingvoiceless:vowelE,_voicingvoiceless:vowelO_)    -0.41     0.46 1.00
+cor(voicingvoiced:vowelI,_voicingvoiceless:vowelO_)       -0.32     0.54 1.00
+cor(voicingvoiceless:vowelI,_voicingvoiceless:vowelO_)    -0.39     0.43 1.00
+cor(voicingvoiced:vowelO,_voicingvoiceless:vowelO_)       -0.37     0.50 1.00
+                                                       Bulk_ESS Tail_ESS
+sd(voicingvoiced:vowela_)                                  1686     1641
+sd(voicingvoiceless:vowela_)                               2099     2574
+sd(voicingvoiced:vowelE_)                                  1536     1551
+sd(voicingvoiceless:vowelE_)                               2319     1734
+sd(voicingvoiced:vowelI_)                                  2498     3094
+sd(voicingvoiceless:vowelI_)                               3333     3115
+sd(voicingvoiced:vowelO_)                                  1620     1698
+sd(voicingvoiceless:vowelO_)                               2295     2196
+cor(voicingvoiced:vowela,_voicingvoiceless:vowela_)        2144     2507
+cor(voicingvoiced:vowela,_voicingvoiced:vowelE_)           2466     2751
+cor(voicingvoiceless:vowela,_voicingvoiced:vowelE_)        2821     3574
+cor(voicingvoiced:vowela,_voicingvoiceless:vowelE_)        3289     3353
+cor(voicingvoiceless:vowela,_voicingvoiceless:vowelE_)     3534     3625
+cor(voicingvoiced:vowelE,_voicingvoiceless:vowelE_)        4027     3622
+cor(voicingvoiced:vowela,_voicingvoiced:vowelI_)           2057     2599
+cor(voicingvoiceless:vowela,_voicingvoiced:vowelI_)        2350     2928
+cor(voicingvoiced:vowelE,_voicingvoiced:vowelI_)           2143     2832
+cor(voicingvoiceless:vowelE,_voicingvoiced:vowelI_)        1964     3115
+cor(voicingvoiced:vowela,_voicingvoiceless:vowelI_)        1792     2459
+cor(voicingvoiceless:vowela,_voicingvoiceless:vowelI_)     2219     2880
+cor(voicingvoiced:vowelE,_voicingvoiceless:vowelI_)        2232     2731
+cor(voicingvoiceless:vowelE,_voicingvoiceless:vowelI_)     1602     2249
+cor(voicingvoiced:vowelI,_voicingvoiceless:vowelI_)        2271     3182
+cor(voicingvoiced:vowela,_voicingvoiced:vowelO_)           3973     3016
+cor(voicingvoiceless:vowela,_voicingvoiced:vowelO_)        5614     3512
+cor(voicingvoiced:vowelE,_voicingvoiced:vowelO_)           4179     3070
+cor(voicingvoiceless:vowelE,_voicingvoiced:vowelO_)        4760     3338
+cor(voicingvoiced:vowelI,_voicingvoiced:vowelO_)           4135     3265
+cor(voicingvoiceless:vowelI,_voicingvoiced:vowelO_)        4320     3583
+cor(voicingvoiced:vowela,_voicingvoiceless:vowelO_)        7265     3088
+cor(voicingvoiceless:vowela,_voicingvoiceless:vowelO_)     5934     3240
+cor(voicingvoiced:vowelE,_voicingvoiceless:vowelO_)        6140     3465
+cor(voicingvoiceless:vowelE,_voicingvoiceless:vowelO_)     5229     3304
+cor(voicingvoiced:vowelI,_voicingvoiceless:vowelO_)        4597     3612
+cor(voicingvoiceless:vowelI,_voicingvoiceless:vowelO_)     6014     3479
+cor(voicingvoiced:vowelO,_voicingvoiceless:vowelO_)        3033     3351
+
+Regression Coefficients:
+                         Estimate Est.Error l-80% CI u-80% CI Rhat Bulk_ESS
+voicingvoiced:vowela_       -0.61      0.12    -0.76    -0.45 1.00     3695
+voicingvoiceless:vowela_    -0.07      0.09    -0.19     0.04 1.00     3582
+voicingvoiced:vowelE_       -0.19      0.15    -0.38     0.01 1.00     3940
+voicingvoiceless:vowelE_    -0.25      0.10    -0.38    -0.12 1.00     5156
+voicingvoiced:vowelI_       -0.74      0.19    -0.98    -0.51 1.00     4090
+voicingvoiceless:vowelI_    -0.12      0.17    -0.33     0.10 1.00     4077
+voicingvoiced:vowelO_       -0.33      0.16    -0.54    -0.13 1.00     6095
+voicingvoiceless:vowelO_     0.25      0.16     0.05     0.45 1.00     6834
+                         Tail_ESS
+voicingvoiced:vowela_        3160
+voicingvoiceless:vowela_     3300
+voicingvoiced:vowelE_        2601
+voicingvoiceless:vowelE_     3268
+voicingvoiced:vowelI_        3357
+voicingvoiceless:vowelI_     3305
+voicingvoiced:vowelO_        2719
+voicingvoiceless:vowelO_     2927
+
+Further Distributional Parameters:
+    Estimate Est.Error l-80% CI u-80% CI Rhat Bulk_ESS Tail_ESS
+phi     4.73      0.30     4.35     5.12 1.00     2477     3192
+
+Draws were sampled using sampling(NUTS). For each parameter, Bulk_ESS
+and Tail_ESS are effective sample size measures, and Rhat is the potential
+scale reduction factor on split chains (at convergence, Rhat = 1).
+\end{verbatim}
+
+  \bibliography{linguistics.bib}
 

### 2024-12-11 manuscript/manuscript.qmd
edit manuscript

@@
 For conciseness, we will use the `fixef()` function which prints the regression coefficients.
+\[EXPLAIN EXPECTED PREDICTIONS XXX\].
 The full summary with an explanation of each part can be found in XXX.
@@
 ```
+
+Let's inspect the output of `fixef()`.
+
+```{r}
+#| label: nas-prop-bm-fixef
+round(
+  fixef(nas_prop_bm, prob = c(0.1, 0.9)),
+  digits = 2
+)
+```
+
+Negative log-odds indicate a proportion that is smaller than 50%, while positive log-odds a proportion that is greater than 50%.
+Generally, the expected log-odd predictions are negative, indicating an overall tendency for the nasalisation to take less than 50% of the duration of the vowel.
+Moreover, the predictions are higher for voiceless NC sequences than for voiced NC sequences, indicating a greater proportion of nasalisation in the former.
+However there is vowel-specific variation, and there doesn't seem to be much of a difference in nasalisation proportion in /e/ and /u/.
+Plotting the expected predictions with `conditional_effects()` should make this clearer.
+
+```{r}
+#| label: nas-prop
+
+conditional_effects(nas_prop_bm, "vowel:voicing")
+
+```
+
+Now that we fitted the model we can use the draws to answer two questions:
+
+1.  Is the nasalisation proportion, on average across vowels, greater in voiceless NC sequences?
+2.  Is there individual speaker variation?
+
+To answer question 1, we can calculate the average difference in nasalisation proportion by first calculating the average nasalisation across all vowels for voiced and voiceless sequences and then take the difference of those, similarly to what we have done in the Case Study 1 above.
+
+```{r}
+#| label: nas-prop-bm-draws
+
+# extract only coefficient variables starting with "b_"
+nas_prop_bm_draws <- as_draws_df(nas_prop_bm, variable = "^b_", regex = TRUE)
+
+nas_prop_bm_draws_long <- nas_prop_bm_draws |> 
+  # pivot vowel:c2 columns
+  pivot_longer(`b_voicingvoiced:vowela`:`b_voicingvoiceless:vowelu`, names_to = "coeff") |>
+  # separate "coeff" labels into type ("b"), vowel and c2
+  separate(coeff, into = c("type", "voicing", "vowel"))
+
+nas_prop_bm_draws_long
+```
+
+Now let's calculate the mean nasalisation proportion within each draw by voicing, and plot the resulting posterior distributions.
+
+```{r}
+#| label: nas-prop-bm-draws-long-voicing
+
+nas_prop_bm_draws_long_voicing <- nas_prop_bm_draws_long |> 
+  # grouping by .draw and voicing ensures that averaging applies only within draw and voicing
+  group_by(.draw, voicing) |> 
+  # calculate the mean value within draw/voicing
+  summarise(
+    value_mean = mean(value), .groups = "drop"
+  )
+
+nas_prop_bm_draws_long_voicing |> 
+  ggplot(aes(plogis(value_mean), voicing)) +
+  stat_halfeye(.width = c(0.6, 0.8))
+```
+
+The plot suggests an overall greater nasalisation proportion in voiceless NC sequences. Let's quantify how greater as we did in Case Study 1.
+
+```{r}
+#| label: nas-prop-bm-diff
+
+nas_prop_bm_diff <- nas_prop_bm_draws_long_voicing |> 
+  # pivot data to create one column per voicing with the mean drawn values,
+  # with one draw per raw
+  pivot_wider(names_from = voicing, values_from = value_mean) |> 
+  mutate(
+    # calculate the difference of voiceless and voiced
+    voicing_diff = voicingvoiceless - voicingvoiced
+  )
+
+nas_prop_bm_diff |> 
+  reframe(
+    # 90% CrI
+    q90 = quantile2(exp(voicing_diff), probs = c(0.05, 0.95)),
+    # 80% CrI
+    q80 = quantile2(exp(voicing_diff), probs = c(0.1, 0.9)),
+    # 60% CrI
+    q60 = quantile2(exp(voicing_diff), probs = c(0.2, 0.8)),
+  ) |> 
+  mutate(across(everything(), ~round(.x, 2)))
+```
+
+The CrIs of the ratio difference in nasalisation proportion in voiceless vs voiced NC sequences suggest a robust increase of nasalisation in the voiceless NC sequences, with a 90% probability that the increase is between 23% and 69% of the proportion in voiced NC sequences.
+
+```{r}
+#| label: nas-prop-r
+library(tidybayes)
+nas_prop_r <- nas_prop_bm |> 
+  spread_draws(r_speaker[speaker,voicingvowel]) |> 
+  separate(voicingvowel, c("voicing", "vowel")) |> 
+  left_join(y = nas_prop_bm_draws_long) |> 
+  mutate(r_speaker_value = value + r_speaker) |> 
+  group_by(.draw, speaker, voicing) |> 
+  summarise(r_speaker_value_mean = mean(r_speaker_value)) |> 
+  pivot_wider(names_from = voicing, values_from = r_speaker_value_mean) |> 
+  mutate(voicing_diff = voicingvoiceless - voicingvoiced)
+```
+
+```{r}
+#| label: fig-nas-prop-r
+#| warning: false
+
+nas_prop_r |> 
+  ggplot(aes(voicing_diff, reorder(speaker, voicing_diff))) +
+  stat_halfeye() +
+  geom_vline(xintercept = mean(nas_prop_bm_diff$voicing_diff), colour = "red") +
+  geom_vline(xintercept = 0, linetype = "dashed") +
+  lims(x = c(-1, 1.5))
+```

### 2024-12-13 manuscript/manuscript.qmd
add Paul to authors
edit manuscript

@@
     note: This is the first author footnote.
+  - name: Paul Bürkner
+    email: paul.buerkner@gmail.com
 abstract: |
@@
 
+Moving onto question 2: is there individual speaker variation?
+[TAMMINGA XXX] For this, we will use the `spread_draws()` function from tidybayes [XXX] to extract the draws of the varying terms (in brms these are the coefficients that start with `r_`).
+There is quite a few steps of processing to get from the raw draws to the estimand we need: while we have commented the following code, we encourage readers to test each line sequentialy and inspect the intermediate output to fully understand the process.
+
 ```{r}
@@
 library(tidybayes)
+
 nas_prop_r <- nas_prop_bm |> 
+  # extract draws from model, only `r_speaker` varying terms
   spread_draws(r_speaker[speaker,voicingvowel]) |> 
+  # separate the column voicingvowel to two columns
   separate(voicingvowel, c("voicing", "vowel")) |> 
+  # join the draws with the `b_` terms
   left_join(y = nas_prop_bm_draws_long) |> 
+  # get the expected log-odd value of each speaker, in each draw
+  # this is the sum of the `value` from the b_ terms and the value from the
+  # r_speaker term.
   mutate(r_speaker_value = value + r_speaker) |> 
+  # group the data for summarise
   group_by(.draw, speaker, voicing) |> 
+  # get mean expected log-odds by draw, speaker and voicing (averaging across vowel)
   summarise(r_speaker_value_mean = mean(r_speaker_value)) |> 
+  # make the data wider: two columns, one for voiced and one for voiceless
   pivot_wider(names_from = voicing, values_from = r_speaker_value_mean) |> 
+  # finally, calculate the difference in expected log-odds of voiceless and voiced
   mutate(voicing_diff = voicingvoiceless - voicingvoiced)
+
+nas_prop_r
 ```
+
+@fig-nas-prop-r plots the posterior distributions of the expected log-odd difference of coarticulatory nasalisation in voiceless vs voiced NC sequences (*x*-axis), for each speaker in the data (*y*-axis), as predicted by the model.
+The red solid vertical line indicates the constant (overall) expected log-odd difference based on the (constant) `b_` terms.
+The black dashed vertical line marks log-difference 0 (i.e., no difference in proportion of nasalisation between voiceless and voiced NC).
+There is a lot of uncertainty within and between speakers: while the distributions of most speakers are located in the positive range, some expected distributions (see bottom of figure) do span both negative and positive values.
+In other words, while most speakers are expected to have a larger nasalisation proportion in voiceless NC sequences, a few in fact might have the opposite pattern.
+Even among those speakers that do have a positive difference, there is a lot of uncertainty as to the magnitude of the difference.
 

### 2024-12-14 manuscript/manuscript.qmd
edit manuscript

@@
 
-Finally, the package marginaleffects has two other convenience functions that return CrIs of comparisons across predictor levels (`avg_comparisons()`) and CrIs of posterior predictions across predictor levels (`avg_predictions`).
+Finally, the package marginaleffects [XXX] has two other convenience functions that return CrIs of comparisons across predictor levels (`avg_comparisons()`) and CrIs of posterior predictions across predictor levels (`avg_predictions`).
 \[XXX\]
@@
 The black dashed vertical line marks log-difference 0 (i.e., no difference in proportion of nasalisation between voiceless and voiced NC).
-There is a lot of uncertainty within and between speakers: while the distributions of most speakers are located in the positive range, some expected distributions (see bottom of figure) do span both negative and positive values.
-In other words, while most speakers are expected to have a larger nasalisation proportion in voiceless NC sequences, a few in fact might have the opposite pattern.
-Even among those speakers that do have a positive difference, there is a lot of uncertainty as to the magnitude of the difference.
 
@@
 ```
+
+There is a lot of uncertainty within and between speakers: while the distributions of most speakers are located in the positive range, some expected distributions (see bottom of figure) do span both negative and positive values.
+In other words, while most speakers are expected to have a larger nasalisation proportion in voiceless NC sequences, a few in fact might have the opposite pattern.
+Even among those speakers that do have a positive difference, there is a lot of uncertainty as to the magnitude of the difference.

### 2024-12-14 manuscript/manuscript.tex
edit manuscript

--- 2024-12-06 manuscript/manuscript.tex
@@
   pdftitle={A tutorial of Bayesian beta regressions with brms in R},
-  pdfauthor={Stefano Coretta},
+  pdfauthor={Stefano Coretta; Paul Bürkner},
   pdfkeywords={keyword1, keyword2},
@@
  \ead{s.coretta@ed.ac.uk} 
+\author[]{Paul Bürkner%
+%
+}
+ \ead{paul.buerkner@gmail.com}
 
@@
 conciseness, we will use the \texttt{fixef()} function which prints the
-regression coefficients. The full summary with an explanation of each
-part can be found in XXX. For each coefficient in the model,
-\texttt{fixef()} prints the name of the coefficient, the mean estimate,
-the estimate error and the lower and upper limits of a Bayesian Credible
-interval (CrI). Here, we print an 80\% CrI. There is nothing special
-about 95\% CrI within Bayesian inference and instead experts recommend
-to check and report a variety of CrIs. Bayesian CrIs indicate that at a
-certain probability levels the ``true'' estimate lies within that
-interval: so, for example, a 90\% CrI {[}A, B{]} indicates that there is
-a 90\% probability that the ``true'' estimate is between A and B.
-Different probability levels correspond to different levels of
-confidence: the higher the probability the higher the confidence (always
-conditional on data and model). Obtaining CrIs at different probability
-levels allows researchers to make more fine-grained inferential
-statements than the frequentist significance dichotomy affords. For
-simplicity of exposition, we will use 80\% CrIs in this case study but
-we strongly recommend researchers to always obtain CrIs at different
-levels of probability and base their inferences on all and not one in
-particular. To reiterate, in Bayesian inference, an 80\% CrI indicates
-the range of values within which the true estimate falls at 80\%
-probability or confidence. We round all values to the nearest 2 digits
-for clarity.
+regression coefficients. {[}EXPLAIN EXPECTED PREDICTIONS XXX{]}. The
+full summary with an explanation of each part can be found in XXX. For
+each coefficient in the model, \texttt{fixef()} prints the name of the
+coefficient, the mean estimate, the estimate error and the lower and
+upper limits of a Bayesian Credible interval (CrI). Here, we print an
+80\% CrI. There is nothing special about 95\% CrI within Bayesian
+inference and instead experts recommend to check and report a variety of
+CrIs. Bayesian CrIs indicate that at a certain probability levels the
+``true'' estimate lies within that interval: so, for example, a 90\% CrI
+{[}A, B{]} indicates that there is a 90\% probability that the ``true''
+estimate is between A and B. Different probability levels correspond to
+different levels of confidence: the higher the probability the higher
+the confidence (always conditional on data and model). Obtaining CrIs at
+different probability levels allows researchers to make more
+fine-grained inferential statements than the frequentist significance
+dichotomy affords. For simplicity of exposition, we will use 80\% CrIs
+in this case study but we strongly recommend researchers to always
+obtain CrIs at different levels of probability and base their inferences
+on all and not one in particular. To reiterate, in Bayesian inference,
+an 80\% CrI indicates the range of values within which the true estimate
+falls at 80\% probability or confidence. We round all values to the
+nearest 2 digits for clarity.
 
@@
 
-\includegraphics{manuscript_files/figure-pdf/voi-prop-bm-cond-1.pdf}
-
-Finally, the package marginaleffects has two other convenience functions
-that return CrIs of comparisons across predictor levels
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/voi-prop-bm-cond-1.pdf}}
+
+Finally, the package marginaleffects {[}XXX{]} has two other convenience
+functions that return CrIs of comparisons across predictor levels
 (\texttt{avg\_comparisons()}) and CrIs of posterior predictions across
@@
 
-\renewcommand\refname{Case study 2: pre-nasalisation in Greek}
-  \bibliography{bibliography.bib}
+\section{Case study 2: coarticulatory vowel
+nasalisation}\label{case-study-2-coarticulatory-vowel-nasalisation}
+
+For the second case study we will use data from \citet{carignan2021}.
+The study looked at properties of nasality in German VNC sequences.
+Here, we will focus on the effect of C voicing (voiceless /t/ vs voiced
+/d/) on the proportion of nasalisation within the vowel in the VNC
+sequence. Previous work on coarticulatory nasalisation in English has
+suggested that vowels followed by an NC sequence where C is voiceless
+(NT) should show earlier coarticulatory nasalisation than vowels
+followed by an NC sequence where C is voiced \citep[ND, see review
+in][]{carignan2021}. This pattern has been suggested to be driven by the
+articulatory and acoustic incompatibility of voicelessness and
+nasalisation, by which the velum opening gesture of the nasal consonant
+is pushed away (i.e.~earlier) when the consonant following the nasal is
+voiceless. Everything else being equal, a greater proportion of vowel
+nasalisation (from the perspective of time) should be found in vowels
+followed by NT than in vowels followed by ND.
+
+We will model the proportion of coarticulatory nasalisation in the
+German short vowels /i, e, a, o, u/ when followed by /nt/ or /nd/, using
+a Bayesian beta regression model. The proportion was calculated as the
+proportion of the nasal interval to the duration of the vowel. The nasal
+interval was defined thus: the interval between the time of peak
+velocity of velum opening to the offset of the vowel.
+
+First, let's read and plot the data.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{nasal }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"data/carignan2021/nasal.csv"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+Rows: 705 Columns: 6
+-- Column specification --------------------------------------------------------
+Delimiter: ","
+chr (5): speaker, label, vowel, NC, voicing
+dbl (1): nas_prop
+
+i Use `spec()` to retrieve the full column specification for this data.
+i Specify the column types or set `show_col_types = FALSE` to quiet this message.
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{nasal}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+# A tibble: 705 x 6
+   speaker label               vowel NC    voicing   nas_prop
+   <chr>   <chr>               <chr> <chr> <chr>        <dbl>
+ 1 S03     b__U_nt_@___N_B17/s u     nt    voiceless  0.367  
+ 2 S03     b__a_nd_@___N_B19/s a     nd    voiced     0.195  
+ 3 S03     b__a_nt_@___N_B15/s a     nt    voiceless  0.279  
+ 4 S03     f__I_nt_@___N_B05/s i     nt    voiceless  0.764  
+ 5 S03     l__I_nd_@___N_B06/s i     nd    voiced     0.00529
+ 6 S03     p__E_nt_____N_B09/s e     nt    voiceless  0.335  
+ 7 S03     r__a_nt_@___N_B06/s a     nt    voiceless  0.243  
+ 8 S03     v__I_nd_@___N_B07/s i     nd    voiced     0.0248 
+ 9 S03     v__I_nt_6___N_B15/s i     nt    voiceless  0.135  
+10 S03     z__E_nd_@___N_B17/s e     nd    voiced     0.538  
+# i 695 more rows
+\end{verbatim}
+
+\begin{itemize}
+\item
+  \texttt{speaker} indicates the speaker ID.
+\item
+  \texttt{label} is the word label as given in the original data.
+\item
+  \texttt{vowel} is the target vowel in the VNC sequence.
+\item
+  \texttt{NC} is the NC sequence.
+\item
+  \texttt{voicing} indicates the voicing of C.
+\item
+  \texttt{nas\_prop} is the proportion of coarticulatory nasalisation of
+  the vowel.
+\end{itemize}
+
+\begin{figure}
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-nasal-1.pdf}}
+
+}
+
+\caption{\label{fig-nasal}Proportion of coarticulatory nasalisation
+during the vowel in VNC sequences in German, depending on C voicing.}
+
+\end{figure}%
+
+Figure~\ref{fig-nasal} shows the proportion of coarticulatory
+nasalisation in vowels followed by /nd/ (voiced) vs /nt/ (voiceless)
+sequences, for the short vowels /i, e, a, o, u/. We can see a pattern of
+higher nasalisation proportion in vowels followed by /nt/, at least in
+the vowels /a, i, o/. For /e, u/, the distribution of nasalisation
+proportion seems to be similar between the voiced and voiceless
+contexts.
+
+Now onto modelling with a beta regression. Note that a full appropriate
+model would include further predictors (both constant and varying), but
+for simplicity here we include only the following predictors: voicing
+(voiced /nd/ vs voiceless /nt/), vowel (/i, e, a, o, u/), including an
+interaction between them. As varying terms, we include a varying
+intercept by speaker and a by-speaker varying slope for voicing and
+vowel in interaction. As with the model from the first case study,
+voicing and vowel are coded using indexing, by suppressing the intercept
+with \texttt{0\ +}. Here's the code of the model.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{nas\_prop\_bm }\OtherTok{\textless{}{-}} \FunctionTok{brm}\NormalTok{(}
+\NormalTok{  nas\_prop }\SpecialCharTok{\textasciitilde{}} \DecValTok{0} \SpecialCharTok{+}\NormalTok{ voicing}\SpecialCharTok{:}\NormalTok{vowel }\SpecialCharTok{+}\NormalTok{ (}\DecValTok{0} \SpecialCharTok{+}\NormalTok{ voicing}\SpecialCharTok{:}\NormalTok{vowel }\SpecialCharTok{|}\NormalTok{ speaker),}
+  \AttributeTok{data =}\NormalTok{ nasal,}
+  \AttributeTok{family =}\NormalTok{ Beta,}
+  \AttributeTok{cores =} \DecValTok{4}\NormalTok{,}
+  \AttributeTok{seed =} \DecValTok{3749}\NormalTok{,}
+  \AttributeTok{file =} \StringTok{"data/cache/nas\_prop\_bm"}
+\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+Let's inspect the output of \texttt{fixef()}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{round}\NormalTok{(}
+  \FunctionTok{fixef}\NormalTok{(nas\_prop\_bm, }\AttributeTok{prob =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{)),}
+  \AttributeTok{digits =} \DecValTok{2}
+\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+                        Estimate Est.Error   Q10   Q90
+voicingvoiced:vowela       -0.61      0.12 -0.77 -0.46
+voicingvoiceless:vowela    -0.08      0.09 -0.19  0.04
+voicingvoiced:vowele       -0.19      0.16 -0.38  0.01
+voicingvoiceless:vowele    -0.25      0.10 -0.38 -0.12
+voicingvoiced:voweli       -0.74      0.19 -0.98 -0.51
+voicingvoiceless:voweli    -0.12      0.18 -0.35  0.10
+voicingvoiced:vowelo       -0.34      0.16 -0.54 -0.14
+voicingvoiceless:vowelo     0.25      0.15  0.05  0.44
+voicingvoiced:vowelu       -0.71      0.18 -0.94 -0.49
+voicingvoiceless:vowelu    -0.58      0.16 -0.78 -0.38
+\end{verbatim}
+
+Negative log-odds indicate a proportion that is smaller than 50\%, while
+positive log-odds a proportion that is greater than 50\%. Generally, the
+expected log-odd predictions are negative, indicating an overall
+tendency for the nasalisation to take less than 50\% of the duration of
+the vowel. Moreover, the predictions are higher for voiceless NC
+sequences than for voiced NC sequences, indicating a greater proportion
+of nasalisation in the former. However there is vowel-specific
+variation, and there doesn't seem to be much of a difference in
+nasalisation proportion in /e/ and /u/. Plotting the expected
+predictions with \texttt{conditional\_effects()} should make this
+clearer.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{conditional\_effects}\NormalTok{(nas\_prop\_bm, }\StringTok{"vowel:voicing"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/nas-prop-1.pdf}}
+
+Now that we fitted the model we can use the draws to answer two
+questions:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Is the nasalisation proportion, on average across vowels, greater in
+  voiceless NC sequences?
+\item
+  Is there individual speaker variation?
+\end{enumerate}
+
+To answer question 1, we can calculate the average difference in
+nasalisation proportion by first calculating the average nasalisation
+across all vowels for voiced and voiceless sequences and then take the
+difference of those, similarly to what we have done in the Case Study 1
+above.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# extract only coefficient variables starting with "b\_"}
+\NormalTok{nas\_prop\_bm\_draws }\OtherTok{\textless{}{-}} \FunctionTok{as\_draws\_df}\NormalTok{(nas\_prop\_bm, }\AttributeTok{variable =} \StringTok{"\^{}b\_"}\NormalTok{, }\AttributeTok{regex =} \ConstantTok{TRUE}\NormalTok{)}
+
+\NormalTok{nas\_prop\_bm\_draws\_long }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm\_draws }\SpecialCharTok{|\textgreater{}} 
+  \CommentTok{\# pivot vowel:c2 columns}
+  \FunctionTok{pivot\_longer}\NormalTok{(}\StringTok{\textasciigrave{}}\AttributeTok{b\_voicingvoiced:vowela}\StringTok{\textasciigrave{}}\SpecialCharTok{:}\StringTok{\textasciigrave{}}\AttributeTok{b\_voicingvoiceless:vowelu}\StringTok{\textasciigrave{}}\NormalTok{, }\AttributeTok{names\_to =} \StringTok{"coeff"}\NormalTok{) }\SpecialCharTok{|\textgreater{}}
+  \CommentTok{\# separate "coeff" labels into type ("b"), vowel and c2}
+  \FunctionTok{separate}\NormalTok{(coeff, }\AttributeTok{into =} \FunctionTok{c}\NormalTok{(}\StringTok{"type"}\NormalTok{, }\StringTok{"voicing"}\NormalTok{, }\StringTok{"vowel"}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+Warning: Dropping 'draws_df' class as required metadata was removed.
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{nas\_prop\_bm\_draws\_long}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+# A tibble: 40,000 x 7
+   .chain .iteration .draw type  voicing          vowel    value
+    <int>      <int> <int> <chr> <chr>            <chr>    <dbl>
+ 1      1          1     1 b     voicingvoiced    vowela -0.463 
+ 2      1          1     1 b     voicingvoiceless vowela -0.0944
+ 3      1          1     1 b     voicingvoiced    vowele -0.155 
+ 4      1          1     1 b     voicingvoiceless vowele -0.208 
+ 5      1          1     1 b     voicingvoiced    voweli -0.481 
+ 6      1          1     1 b     voicingvoiceless voweli -0.0529
+ 7      1          1     1 b     voicingvoiced    vowelo -0.310 
+ 8      1          1     1 b     voicingvoiceless vowelo  0.321 
+ 9      1          1     1 b     voicingvoiced    vowelu -0.620 
+10      1          1     1 b     voicingvoiceless vowelu -0.368 
+# i 39,990 more rows
+\end{verbatim}
+
+Now let's calculate the mean nasalisation proportion within each draw by
+voicing, and plot the resulting posterior distributions.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{nas\_prop\_bm\_draws\_long\_voicing }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm\_draws\_long }\SpecialCharTok{|\textgreater{}} 
+  \CommentTok{\# grouping by .draw and voicing ensures that averaging applies only within draw and voicing}
+  \FunctionTok{group\_by}\NormalTok{(.draw, voicing) }\SpecialCharTok{|\textgreater{}} 
+  \CommentTok{\# calculate the mean value within draw/voicing}
+  \FunctionTok{summarise}\NormalTok{(}
+    \AttributeTok{value\_mean =} \FunctionTok{mean}\NormalTok{(value), }\AttributeTok{.groups =} \StringTok{"drop"}
+\NormalTok{  )}
+
+\NormalTok{nas\_prop\_bm\_draws\_long\_voicing }\SpecialCharTok{|\textgreater{}} 
+  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\FunctionTok{plogis}\NormalTok{(value\_mean), voicing)) }\SpecialCharTok{+}
+  \FunctionTok{stat\_halfeye}\NormalTok{(}\AttributeTok{.width =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.6}\NormalTok{, }\FloatTok{0.8}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/nas-prop-bm-draws-long-voicing-1.pdf}}
+
+The plot suggests an overall greater nasalisation proportion in
+voiceless NC sequences. Let's quantify how greater as we did in Case
+Study 1.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{nas\_prop\_bm\_diff }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm\_draws\_long\_voicing }\SpecialCharTok{|\textgreater{}} 
+  \CommentTok{\# pivot data to create one column per voicing with the mean drawn values,}
+  \CommentTok{\# with one draw per raw}
+  \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ voicing, }\AttributeTok{values\_from =}\NormalTok{ value\_mean) }\SpecialCharTok{|\textgreater{}} 
+  \FunctionTok{mutate}\NormalTok{(}
+    \CommentTok{\# calculate the difference of voiceless and voiced}
+    \AttributeTok{voicing\_diff =}\NormalTok{ voicingvoiceless }\SpecialCharTok{{-}}\NormalTok{ voicingvoiced}
+\NormalTok{  )}
+
+\NormalTok{nas\_prop\_bm\_diff }\SpecialCharTok{|\textgreater{}} 
+  \FunctionTok{reframe}\NormalTok{(}
+    \CommentTok{\# 90\% CrI}
+    \AttributeTok{q90 =} \FunctionTok{quantile2}\NormalTok{(}\FunctionTok{exp}\NormalTok{(voicing\_diff), }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.05}\NormalTok{, }\FloatTok{0.95}\NormalTok{)),}
+    \CommentTok{\# 80\% CrI}
+    \AttributeTok{q80 =} \FunctionTok{quantile2}\NormalTok{(}\FunctionTok{exp}\NormalTok{(voicing\_diff), }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{)),}
+    \CommentTok{\# 60\% CrI}
+    \AttributeTok{q60 =} \FunctionTok{quantile2}\NormalTok{(}\FunctionTok{exp}\NormalTok{(voicing\_diff), }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.2}\NormalTok{, }\FloatTok{0.8}\NormalTok{)),}
+\NormalTok{  ) }\SpecialCharTok{|\textgreater{}} 
+  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{everything}\NormalTok{(), }\SpecialCharTok{\textasciitilde{}}\FunctionTok{round}\NormalTok{(.x, }\DecValTok{2}\NormalTok{)))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+# A tibble: 2 x 3
+    q90   q80   q60
+  <dbl> <dbl> <dbl>
+1  1.23  1.27  1.33
+2  1.69  1.63  1.56
+\end{verbatim}
+
+The CrIs of the ratio difference in nasalisation proportion in voiceless
+vs voiced NC sequences suggest a robust increase of nasalisation in the
+voiceless NC sequences, with a 90\% probability that the increase is
+between 23\% and 69\% of the proportion in voiced NC sequences.
+
+Moving onto question 2: is there individual speaker variation?
+{[}TAMMINGA XXX{]} For this, we will use the \texttt{spread\_draws()}
+function from tidybayes {[}XXX{]} to extract the draws of the varying
+terms (in brms these are the coefficients that start with \texttt{r\_}).
+There is quite a few steps of processing to get from the raw draws to
+the estimand we need: while we have commented the following code, we
+encourage readers to test each line sequentialy and inspect the
+intermediate output to fully understand the process.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{library}\NormalTok{(tidybayes)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+
+Attaching package: 'tidybayes'
+\end{verbatim}
+
+\begin{verbatim}
+The following objects are masked from 'package:brms':
+
+    dstudent_t, pstudent_t, qstudent_t, rstudent_t
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{nas\_prop\_r }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm }\SpecialCharTok{|\textgreater{}} 
+  \CommentTok{\# extract draws from model, only \textasciigrave{}r\_speaker\textasciigrave{} varying terms}
+  \FunctionTok{spread\_draws}\NormalTok{(r\_speaker[speaker,voicingvowel]) }\SpecialCharTok{|\textgreater{}} 
+  \CommentTok{\# separate the column voicingvowel to two columns}
+  \FunctionTok{separate}\NormalTok{(voicingvowel, }\FunctionTok{c}\NormalTok{(}\StringTok{"voicing"}\NormalTok{, }\StringTok{"vowel"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}} 
+  \CommentTok{\# join the draws with the \textasciigrave{}b\_\textasciigrave{} terms}
+  \FunctionTok{left\_join}\NormalTok{(}\AttributeTok{y =}\NormalTok{ nas\_prop\_bm\_draws\_long) }\SpecialCharTok{|\textgreater{}} 
+  \CommentTok{\# get the expected log{-}odd value of each speaker, in each draw}
+  \CommentTok{\# this is the sum of the \textasciigrave{}value\textasciigrave{} from the b\_ terms and the value from the}
+  \CommentTok{\# r\_speaker term.}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{r\_speaker\_value =}\NormalTok{ value }\SpecialCharTok{+}\NormalTok{ r\_speaker) }\SpecialCharTok{|\textgreater{}} 
+  \CommentTok{\# group the data for summarise}
+  \FunctionTok{group\_by}\NormalTok{(.draw, speaker, voicing) }\SpecialCharTok{|\textgreater{}} 
+  \CommentTok{\# get mean expected log{-}odds by draw, speaker and voicing (averaging across vowel)}
+  \FunctionTok{summarise}\NormalTok{(}\AttributeTok{r\_speaker\_value\_mean =} \FunctionTok{mean}\NormalTok{(r\_speaker\_value)) }\SpecialCharTok{|\textgreater{}} 
+  \CommentTok{\# make the data wider: two columns, one for voiced and one for voiceless}
+  \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ voicing, }\AttributeTok{values\_from =}\NormalTok{ r\_speaker\_value\_mean) }\SpecialCharTok{|\textgreater{}} 
+  \CommentTok{\# finally, calculate the difference in expected log{-}odds of voiceless and voiced}
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{voicing\_diff =}\NormalTok{ voicingvoiceless }\SpecialCharTok{{-}}\NormalTok{ voicingvoiced)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+Joining with `by = join_by(voicing, vowel, .chain, .iteration, .draw)`
+\end{verbatim}
+
+\begin{verbatim}
+`summarise()` has grouped output by '.draw', 'speaker'. You can override using
+the `.groups` argument.
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{nas\_prop\_r}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+# A tibble: 140,000 x 5
+# Groups:   .draw, speaker [140,000]
+   .draw speaker voicingvoiced voicingvoiceless voicing_diff
+   <int> <chr>           <dbl>            <dbl>        <dbl>
+ 1     1 S03          -1.42              0.0676      1.49   
+ 2     1 S04           0.209             0.185      -0.0238 
+ 3     1 S05          -0.284            -0.0546      0.229  
+ 4     1 S06           0.00669          -0.324      -0.331  
+ 5     1 S07          -0.446            -0.147       0.299  
+ 6     1 S08          -0.610            -0.127       0.483  
+ 7     1 S09          -0.473             0.0497      0.523  
+ 8     1 S10          -0.556            -0.311       0.245  
+ 9     1 S11           0.201            -0.426      -0.628  
+10     1 S12          -0.0812           -0.0856     -0.00439
+# i 139,990 more rows
+\end{verbatim}
+
+Figure~\ref{fig-nas-prop-r} plots the posterior distributions of the
+expected log-odd difference of coarticulatory nasalisation in voiceless
+vs voiced NC sequences (\emph{x}-axis), for each speaker in the data
+(\emph{y}-axis), as predicted by the model. The red solid vertical line
+indicates the constant (overall) expected log-odd difference based on
+the (constant) \texttt{b\_} terms. The black dashed vertical line marks
+log-difference 0 (i.e., no difference in proportion of nasalisation
+between voiceless and voiced NC).
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{nas\_prop\_r }\SpecialCharTok{|\textgreater{}} 
+  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(voicing\_diff, }\FunctionTok{reorder}\NormalTok{(speaker, voicing\_diff))) }\SpecialCharTok{+}
+  \FunctionTok{stat\_halfeye}\NormalTok{() }\SpecialCharTok{+}
+  \FunctionTok{geom\_vline}\NormalTok{(}\AttributeTok{xintercept =} \FunctionTok{mean}\NormalTok{(nas\_prop\_bm\_diff}\SpecialCharTok{$}\NormalTok{voicing\_diff), }\AttributeTok{colour =} \StringTok{"red"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{geom\_vline}\NormalTok{(}\AttributeTok{xintercept =} \DecValTok{0}\NormalTok{, }\AttributeTok{linetype =} \StringTok{"dashed"}\NormalTok{) }\SpecialCharTok{+}
+  \FunctionTok{lims}\NormalTok{(}\AttributeTok{x =} \FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\DecValTok{1}\NormalTok{, }\FloatTok{1.5}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{figure}[H]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-nas-prop-r-1.pdf}}
+
+}
+
+\caption{\label{fig-nas-prop-r}}
+
+\end{figure}%
+
+There is a lot of uncertainty within and between speakers: while the
+distributions of most speakers are located in the positive range, some
+expected distributions (see bottom of figure) do span both negative and
+positive values. In other words, while most speakers are expected to
+have a larger nasalisation proportion in voiceless NC sequences, a few
+in fact might have the opposite pattern. Even among those speakers that
+do have a positive difference, there is a lot of uncertainty as to the
+magnitude of the difference.
+
+  \bibliography{linguistics.bib}
 

### 2024-12-19 manuscript/manuscript.qmd
edit according to Paul's comments

@@
 #| label: read-ita-egg
+#| message: false
+#| warning: false
 
@@
 From the summary, we see that speech rate (number of syllables per second) has a positive effect on voicing proportion: the 80% CrI is between 0.01 and 0.15 log-odds \[$\beta$ = 0.8, SD = 0.06\].
-Log-odds can be converted to odd ratios by exponentiating the value: 0.01-0.15 log odds correspond to an odd ratio of 1.01 to 1.16, or as percentages, to an increase of voicing of 1 to 16% for every increase of one syllable per second.
+Log-odds can be converted to odd-ratios by exponentiating the value: 0.01-0.15 log odds correspond to an odd-ratio of 1.01 to 1.16, or as percentages, to an increase of voicing of 1 to 16% for every increase of one syllable per second.
 Since this is an 80% CrI, we can be 80% confident that the true effect of speech rate is between 1-16% increase of voicing proportion, conditional on the data and model.
 Note that transforming measures this way is appropriate *only* with quantile-based measures (like CrIs) but not with moments like the mean and standard deviation: to correctly get mean and SDs in the transformed scale, you must first extract the posterior draws (see below), convert them and then take moments such as mean and SD (for a more detailed explanation, see XXX).
+In the avoidance of doubt, we will always transform the drawn values first and then take summary measures.
 
@@
 Beware that the values of the difference are in log-odds!
-We can convert these into odd ratios with `exp()`.
+We can convert these into odd-ratios with `exp()`.
+odd-ratios indicate the ratio of the difference between A and B, so that 1 means no difference, values greater than 1 indicate an increase in A relative to B and values lower than 1 indicate a decrease in A relative to B.
+odd-ratios are useful when looking at differences that are in log-odds because while the relative magnitude of the difference in proportion between two groups is the same independent of the baseline proportion, the *absolute* magnitude of the difference depends on the baseline value.
+For example, an odd-ratio difference of 1.25 would correspond to a proportion increase of 13 percentage points if the baseline proportion is 0.62 but it would correspond to a proportion increase of 17 percentage points if the baseline proportion is 0.73.
+Of course, in real research contexts it is still usefull to think about absolute magnitudes and their relevance from a conceptual and methodological perspective.
+In this tutorial we just focus on odd-ratios for simplicity.
 
@@
 voi_prop_bm_diff |> 
+  mutate(c2p_tk_diff_ratio = exp(c2p_tk_diff)) |> 
   reframe(
     # 90% CrI
-    q90 = quantile2(c2p_tk_diff, probs = c(0.05, 0.95)),
+    q90 = quantile2(c2p_tk_diff_ratio, probs = c(0.05, 0.95)),
     # 80% CrI
-    q80 = quantile2(c2p_tk_diff, probs = c(0.1, 0.9)),
+    q80 = quantile2(c2p_tk_diff_ratio, probs = c(0.1, 0.9)),
     # 60% CrI
-    q60 = quantile2(c2p_tk_diff, probs = c(0.2, 0.8)),
+    q60 = quantile2(c2p_tk_diff_ratio, probs = c(0.2, 0.8)),
   ) |> 
-  # exponentiate to get odd ratios and round to 2 digits
-  mutate(across(everything(), ~round(exp(.x), 2)))
+  # round to 2 digits
+  mutate(across(everything(), ~round(.x, 2)))
 ```
@@
 The nasal interval was defined thus: the interval between the time of peak velocity of velum opening to the offset of the vowel.
+We will use the results of the regression model to answer the following questions:
+
+1.  Is the nasalisation proportion, on average across vowels, greater in voiceless NC sequences?
+2.  Is there individual speaker variation?
 
@@
 #| label: nasal
+#| message: false
+#| warning: false
+
 nasal <- read_csv("data/carignan2021/nasal.csv")
@@
 
-Now that we fitted the model we can use the draws to answer two questions:
+Now that we fitted the model we can use the draws to answer the two research questions (repeated from above):
 
@@
 Now let's calculate the mean nasalisation proportion within each draw by voicing, and plot the resulting posterior distributions.
+Note that, as discussed for Case study 1, when working wit log-odds it is important to first do all necessary calculations in log-odds, here calculate the mean log-odds across vowels, and *then* transform the calculated estimands to proportions/probabilities.
 
@@
   group_by(.draw, voicing) |> 
-  # calculate the mean value within draw/voicing
   summarise(
-    value_mean = mean(value), .groups = "drop"
+    # calculate the mean value within draw/voicing in log-odds
+    value_mean = mean(value),
+    # we can now transform log-odds to proportion with plogis()
+    value_mean_prop = plogis(value_mean),
+    .groups = "drop"
   )
 
+nas_prop_bm_draws_long_voicing
+```
+
+```{r}
+#| label: nas-prop-bm-draws-long-voicing-plot
+
 nas_prop_bm_draws_long_voicing |> 
-  ggplot(aes(plogis(value_mean), voicing)) +
+  ggplot(aes(value_mean_prop, voicing)) +
   stat_halfeye(.width = c(0.6, 0.8))
@@
 
-The plot suggests an overall greater nasalisation proportion in voiceless NC sequences. Let's quantify how greater as we did in Case Study 1.
+The plot suggests an overall greater nasalisation proportion in voiceless NC sequences.
+Let's quantify how greater as we did in Case Study 1.
+We will use odd-ratios in this context as well, i.e.
+we will convert log-odds to odd-ratios using the `exp()` (exponential) function (and as before we first calculate the difference and then exponentiate the resulting values, after which we can take summary measures, like means and quantile-based measures such as CrIs).
 
@@
   # pivot data to create one column per voicing with the mean drawn values,
-  # with one draw per raw
+  # with one draw per raw. we need to drop the value_mean_prop col
+  select(-value_mean_prop) |> 
   pivot_wider(names_from = voicing, values_from = value_mean) |> 
   mutate(
-    # calculate the difference of voiceless and voiced
-    voicing_diff = voicingvoiceless - voicingvoiced
+    # calculate the difference of voiceless and voiced in log-odds
+    voicing_diff = voicingvoiceless - voicingvoiced,
+    # now transform with exp() to get the ratio difference
+    voicing_diff_ratio = exp(voicing_diff)
   )
+nas_prop_bm_diff
+```
+
+```{r}
+#| label: nas-prop-bm-diff-cris
 
@@
     # 90% CrI
-    q90 = quantile2(exp(voicing_diff), probs = c(0.05, 0.95)),
+    q90 = quantile2(voicing_diff_ratio, probs = c(0.05, 0.95)),
     # 80% CrI
-    q80 = quantile2(exp(voicing_diff), probs = c(0.1, 0.9)),
+    q80 = quantile2(voicing_diff_ratio, probs = c(0.1, 0.9)),
     # 60% CrI
-    q60 = quantile2(exp(voicing_diff), probs = c(0.2, 0.8)),
+    q60 = quantile2(voicing_diff_ratio, probs = c(0.2, 0.8)),
   ) |> 
@@
 There is quite a few steps of processing to get from the raw draws to the estimand we need: while we have commented the following code, we encourage readers to test each line sequentialy and inspect the intermediate output to fully understand the process.
+We assume that readers are familiar enough with models with varying terms (aka random effects, mixed-effects models).
+What readers should note is that to obtain the expected predictions of nasalisation proportion for each speaker, the constant terms and the varying terms should be added (since the varying terms indicate the deviation of each speaker from the overall estimate).
 
@@
 #| label: nas-prop-r
+#| message: false
+#| warning: false
 library(tidybayes)
@@
 #| warning: false
+#| fig-asp: 1.5
 
@@
 
-There is a lot of uncertainty within and between speakers: while the distributions of most speakers are located in the positive range, some expected distributions (see bottom of figure) do span both negative and positive values.
-In other words, while most speakers are expected to have a larger nasalisation proportion in voiceless NC sequences, a few in fact might have the opposite pattern.
-Even among those speakers that do have a positive difference, there is a lot of uncertainty as to the magnitude of the difference.
+There is a lot of uncertainty within and between speakers: while the distributions of most speakers are located in the positive range, some expected distributions (see last 5 speakers at the bottom of figure) do substantially span both negative and positive values.
+In other words, while most speakers are more likely to have a larger nasalisation proportion in voiceless NC sequences, a few might in fact have the opposite pattern.
+Even among those speakers that do have a more robust positive difference, there is a lot of uncertainty as to the magnitude of the difference.

### 2024-12-19 manuscript/manuscript.tex
edit according to Paul's comments

@@
 0.15 log-odds {[}\(\beta\) = 0.8, SD = 0.06{]}. Log-odds can be
-converted to odd ratios by exponentiating the value: 0.01-0.15 log odds
-correspond to an odd ratio of 1.01 to 1.16, or as percentages, to an
+converted to odd-ratios by exponentiating the value: 0.01-0.15 log odds
+correspond to an odd-ratio of 1.01 to 1.16, or as percentages, to an
 increase of voicing of 1 to 16\% for every increase of one syllable per
@@
 then take moments such as mean and SD (for a more detailed explanation,
-see XXX).
+see XXX). In the avoidance of doubt, we will always transform the drawn
+values first and then take summary measures.
 
@@
 that the values of the difference are in log-odds! We can convert these
-into odd ratios with \texttt{exp()}.
+into odd-ratios with \texttt{exp()}. odd-ratios indicate the ratio of
+the difference between A and B, so that 1 means no difference, values
+greater than 1 indicate an increase in A relative to B and values lower
+than 1 indicate a decrease in A relative to B. odd-ratios are useful
+when looking at differences that are in log-odds because while the
+relative magnitude of the difference in proportion between two groups is
+the same independent of the baseline proportion, the \emph{absolute}
+magnitude of the difference depends on the baseline value. For example,
+an odd-ratio difference of 1.25 would correspond to a proportion
+increase of 13 percentage points if the baseline proportion is 0.62 but
+it would correspond to a proportion increase of 17 percentage points if
+the baseline proportion is 0.73. Of course, in real research contexts it
+is still usefull to think about absolute magnitudes and their relevance
+from a conceptual and methodological perspective. In this tutorial we
+just focus on odd-ratios for simplicity.
 
@@
 \NormalTok{voi\_prop\_bm\_diff }\SpecialCharTok{|\textgreater{}} 
+  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{c2p\_tk\_diff\_ratio =} \FunctionTok{exp}\NormalTok{(c2p\_tk\_diff)) }\SpecialCharTok{|\textgreater{}} 
   \FunctionTok{reframe}\NormalTok{(}
     \CommentTok{\# 90\% CrI}
-    \AttributeTok{q90 =} \FunctionTok{quantile2}\NormalTok{(c2p\_tk\_diff, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.05}\NormalTok{, }\FloatTok{0.95}\NormalTok{)),}
+    \AttributeTok{q90 =} \FunctionTok{quantile2}\NormalTok{(c2p\_tk\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.05}\NormalTok{, }\FloatTok{0.95}\NormalTok{)),}
     \CommentTok{\# 80\% CrI}
-    \AttributeTok{q80 =} \FunctionTok{quantile2}\NormalTok{(c2p\_tk\_diff, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{)),}
+    \AttributeTok{q80 =} \FunctionTok{quantile2}\NormalTok{(c2p\_tk\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{)),}
     \CommentTok{\# 60\% CrI}
-    \AttributeTok{q60 =} \FunctionTok{quantile2}\NormalTok{(c2p\_tk\_diff, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.2}\NormalTok{, }\FloatTok{0.8}\NormalTok{)),}
+    \AttributeTok{q60 =} \FunctionTok{quantile2}\NormalTok{(c2p\_tk\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.2}\NormalTok{, }\FloatTok{0.8}\NormalTok{)),}
 \NormalTok{  ) }\SpecialCharTok{|\textgreater{}} 
-  \CommentTok{\# exponentiate to get odd ratios and round to 2 digits}
-  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{everything}\NormalTok{(), }\SpecialCharTok{\textasciitilde{}}\FunctionTok{round}\NormalTok{(}\FunctionTok{exp}\NormalTok{(.x), }\DecValTok{2}\NormalTok{)))}
+  \CommentTok{\# round to 2 digits}
+  \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{everything}\NormalTok{(), }\SpecialCharTok{\textasciitilde{}}\FunctionTok{round}\NormalTok{(.x, }\DecValTok{2}\NormalTok{)))}
 \end{Highlighting}
@@
 interval was defined thus: the interval between the time of peak
-velocity of velum opening to the offset of the vowel.
+velocity of velum opening to the offset of the vowel. We will use the
+results of the regression model to answer the following questions:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  Is the nasalisation proportion, on average across vowels, greater in
+  voiceless NC sequences?
+\item
+  Is there individual speaker variation?
+\end{enumerate}
 
@@
 \NormalTok{nasal }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"data/carignan2021/nasal.csv"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-Rows: 705 Columns: 6
--- Column specification --------------------------------------------------------
-Delimiter: ","
-chr (5): speaker, label, vowel, NC, voicing
-dbl (1): nas_prop
-
-i Use `spec()` to retrieve the full column specification for this data.
-i Specify the column types or set `show_col_types = FALSE` to quiet this message.
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
 \NormalTok{nasal}
@@
 
-Now that we fitted the model we can use the draws to answer two
-questions:
+Now that we fitted the model we can use the draws to answer the two
+research questions (repeated from above):
 
@@
 Now let's calculate the mean nasalisation proportion within each draw by
-voicing, and plot the resulting posterior distributions.
+voicing, and plot the resulting posterior distributions. Note that, as
+discussed for Case study 1, when working wit log-odds it is important to
+first do all necessary calculations in log-odds, here calculate the mean
+log-odds across vowels, and \emph{then} transform the calculated
+estimands to proportions/probabilities.
 
@@
   \FunctionTok{group\_by}\NormalTok{(.draw, voicing) }\SpecialCharTok{|\textgreater{}} 
-  \CommentTok{\# calculate the mean value within draw/voicing}
   \FunctionTok{summarise}\NormalTok{(}
-    \AttributeTok{value\_mean =} \FunctionTok{mean}\NormalTok{(value), }\AttributeTok{.groups =} \StringTok{"drop"}
+    \CommentTok{\# calculate the mean value within draw/voicing in log{-}odds}
+    \AttributeTok{value\_mean =} \FunctionTok{mean}\NormalTok{(value),}
+    \CommentTok{\# we can now transform log{-}odds to proportion with plogis()}
+    \AttributeTok{value\_mean\_prop =} \FunctionTok{plogis}\NormalTok{(value\_mean),}
+    \AttributeTok{.groups =} \StringTok{"drop"}
 \NormalTok{  )}
 
+\NormalTok{nas\_prop\_bm\_draws\_long\_voicing}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+# A tibble: 8,000 x 4
+   .draw voicing          value_mean value_mean_prop
+   <int> <chr>                 <dbl>           <dbl>
+ 1     1 voicingvoiced       -0.406            0.400
+ 2     1 voicingvoiceless    -0.0804           0.480
+ 3     2 voicingvoiced       -0.391            0.404
+ 4     2 voicingvoiceless    -0.123            0.469
+ 5     3 voicingvoiced       -0.444            0.391
+ 6     3 voicingvoiceless    -0.132            0.467
+ 7     4 voicingvoiced       -0.439            0.392
+ 8     4 voicingvoiceless    -0.0982           0.475
+ 9     5 voicingvoiced       -0.383            0.405
+10     5 voicingvoiceless    -0.266            0.434
+# i 7,990 more rows
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
 \NormalTok{nas\_prop\_bm\_draws\_long\_voicing }\SpecialCharTok{|\textgreater{}} 
-  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\FunctionTok{plogis}\NormalTok{(value\_mean), voicing)) }\SpecialCharTok{+}
+  \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(value\_mean\_prop, voicing)) }\SpecialCharTok{+}
   \FunctionTok{stat\_halfeye}\NormalTok{(}\AttributeTok{.width =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.6}\NormalTok{, }\FloatTok{0.8}\NormalTok{))}
@@
 
-\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/nas-prop-bm-draws-long-voicing-1.pdf}}
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/nas-prop-bm-draws-long-voicing-plot-1.pdf}}
 
@@
 voiceless NC sequences. Let's quantify how greater as we did in Case
-Study 1.
+Study 1. We will use odd-ratios in this context as well, i.e.~we will
+convert log-odds to odd-ratios using the \texttt{exp()} (exponential)
+function (and as before we first calculate the difference and then
+exponentiate the resulting values, after which we can take summary
+measures, like means and quantile-based measures such as CrIs).
 
@@
   \CommentTok{\# pivot data to create one column per voicing with the mean drawn values,}
-  \CommentTok{\# with one draw per raw}
+  \CommentTok{\# with one draw per raw. we need to drop the value\_mean\_prop col}
+  \FunctionTok{select}\NormalTok{(}\SpecialCharTok{{-}}\NormalTok{value\_mean\_prop) }\SpecialCharTok{|\textgreater{}} 
   \FunctionTok{pivot\_wider}\NormalTok{(}\AttributeTok{names\_from =}\NormalTok{ voicing, }\AttributeTok{values\_from =}\NormalTok{ value\_mean) }\SpecialCharTok{|\textgreater{}} 
   \FunctionTok{mutate}\NormalTok{(}
-    \CommentTok{\# calculate the difference of voiceless and voiced}
-    \AttributeTok{voicing\_diff =}\NormalTok{ voicingvoiceless }\SpecialCharTok{{-}}\NormalTok{ voicingvoiced}
+    \CommentTok{\# calculate the difference of voiceless and voiced in log{-}odds}
+    \AttributeTok{voicing\_diff =}\NormalTok{ voicingvoiceless }\SpecialCharTok{{-}}\NormalTok{ voicingvoiced,}
+    \CommentTok{\# now transform with exp() to get the ratio difference}
+    \AttributeTok{voicing\_diff\_ratio =} \FunctionTok{exp}\NormalTok{(voicing\_diff)}
 \NormalTok{  )}
-
+\NormalTok{nas\_prop\_bm\_diff}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+# A tibble: 4,000 x 5
+   .draw voicingvoiced voicingvoiceless voicing_diff voicing_diff_ratio
+   <int>         <dbl>            <dbl>        <dbl>              <dbl>
+ 1     1        -0.406          -0.0804        0.325               1.38
+ 2     2        -0.391          -0.123         0.268               1.31
+ 3     3        -0.444          -0.132         0.312               1.37
+ 4     4        -0.439          -0.0982        0.340               1.41
+ 5     5        -0.383          -0.266         0.117               1.12
+ 6     6        -0.541          -0.0903        0.450               1.57
+ 7     7        -0.512          -0.111         0.401               1.49
+ 8     8        -0.493          -0.152         0.341               1.41
+ 9     9        -0.527          -0.175         0.352               1.42
+10    10        -0.491          -0.257         0.234               1.26
+# i 3,990 more rows
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
 \NormalTok{nas\_prop\_bm\_diff }\SpecialCharTok{|\textgreater{}} 
@@
     \CommentTok{\# 90\% CrI}
-    \AttributeTok{q90 =} \FunctionTok{quantile2}\NormalTok{(}\FunctionTok{exp}\NormalTok{(voicing\_diff), }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.05}\NormalTok{, }\FloatTok{0.95}\NormalTok{)),}
+    \AttributeTok{q90 =} \FunctionTok{quantile2}\NormalTok{(voicing\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.05}\NormalTok{, }\FloatTok{0.95}\NormalTok{)),}
     \CommentTok{\# 80\% CrI}
-    \AttributeTok{q80 =} \FunctionTok{quantile2}\NormalTok{(}\FunctionTok{exp}\NormalTok{(voicing\_diff), }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{)),}
+    \AttributeTok{q80 =} \FunctionTok{quantile2}\NormalTok{(voicing\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{)),}
     \CommentTok{\# 60\% CrI}
-    \AttributeTok{q60 =} \FunctionTok{quantile2}\NormalTok{(}\FunctionTok{exp}\NormalTok{(voicing\_diff), }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.2}\NormalTok{, }\FloatTok{0.8}\NormalTok{)),}
+    \AttributeTok{q60 =} \FunctionTok{quantile2}\NormalTok{(voicing\_diff\_ratio, }\AttributeTok{probs =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.2}\NormalTok{, }\FloatTok{0.8}\NormalTok{)),}
 \NormalTok{  ) }\SpecialCharTok{|\textgreater{}} 
@@
 encourage readers to test each line sequentialy and inspect the
-intermediate output to fully understand the process.
+intermediate output to fully understand the process. We assume that
+readers are familiar enough with models with varying terms (aka random
+effects, mixed-effects models). What readers should note is that to
+obtain the expected predictions of nasalisation proportion for each
+speaker, the constant terms and the varying terms should be added (since
+the varying terms indicate the deviation of each speaker from the
+overall estimate).
 
@@
 \FunctionTok{library}\NormalTok{(tidybayes)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-
-Attaching package: 'tidybayes'
-\end{verbatim}
-
-\begin{verbatim}
-The following objects are masked from 'package:brms':
-
-    dstudent_t, pstudent_t, qstudent_t, rstudent_t
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
+
 \NormalTok{nas\_prop\_r }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm }\SpecialCharTok{|\textgreater{}} 
@@
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{voicing\_diff =}\NormalTok{ voicingvoiceless }\SpecialCharTok{{-}}\NormalTok{ voicingvoiced)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-Joining with `by = join_by(voicing, vowel, .chain, .iteration, .draw)`
-\end{verbatim}
-
-\begin{verbatim}
-`summarise()` has grouped output by '.draw', 'speaker'. You can override using
-the `.groups` argument.
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
+
 \NormalTok{nas\_prop\_r}
@@
 distributions of most speakers are located in the positive range, some
-expected distributions (see bottom of figure) do span both negative and
-positive values. In other words, while most speakers are expected to
-have a larger nasalisation proportion in voiceless NC sequences, a few
-in fact might have the opposite pattern. Even among those speakers that
-do have a positive difference, there is a lot of uncertainty as to the
-magnitude of the difference.
+expected distributions (see last 5 speakers at the bottom of figure) do
+substantially span both negative and positive values. In other words,
+while most speakers are more likely to have a larger nasalisation
+proportion in voiceless NC sequences, a few might in fact have the
+opposite pattern. Even among those speakers that do have a more robust
+positive difference, there is a lot of uncertainty as to the magnitude
+of the difference.
 

### 2025-01-06 manuscript/manuscript.qmd
fix typo

@@
 The coefficients of a beta regression are estimated on the log-odds scale, as in Bernoulli/binomial (aka logistic) regressions.
-From the summary, we see that speech rate (number of syllables per second) has a positive effect on voicing proportion: the 80% CrI is between 0.01 and 0.15 log-odds \[$\beta$ = 0.8, SD = 0.06\].
+From the summary, we see that speech rate (number of syllables per second) has a positive effect on voicing proportion: the 80% CrI is between 0.01 and 0.15 log-odds \[$\beta$ = 0.08, SD = 0.06\].
 Log-odds can be converted to odd-ratios by exponentiating the value: 0.01-0.15 log odds correspond to an odd-ratio of 1.01 to 1.16, or as percentages, to an increase of voicing of 1 to 16% for every increase of one syllable per second.

### 2025-01-06 manuscript/manuscript.tex
fix typo

@@
 positive effect on voicing proportion: the 80\% CrI is between 0.01 and
-0.15 log-odds {[}\(\beta\) = 0.8, SD = 0.06{]}. Log-odds can be
+0.15 log-odds {[}\(\beta\) = 0.08, SD = 0.06{]}. Log-odds can be
 converted to odd-ratios by exponentiating the value: 0.01-0.15 log odds

### 2025-02-03 manuscript/manuscript.qmd
add some basic information on the beta distribution

@@
 date: last-modified
-bibliography: linguistics.bib
+bibliography: 
+  - linguistics.bib
+  - statistics.bib
 format:
@@
 
-However, proportion are naturally not Gaussian, since they are limited between 0 and 1.
-The theoretical distribution that generates values with this characteristics is the beta distribution.
-Thus, regression models with proportions as the outcome variable should be fitted using a beta distribution as the distribution family.
+However, proportion are usually not Gaussian, since they are limited between 0 and 1.
+Thus, regression models with proportions as outcome variable should be fitted using a likelihood family that assumes such unit interval data.
+A common and flexible distribution for this purpose is the beta distribution.
 This tutorial introduces researchers to beta regression models in R using the package brms.
@@
 
-...
+The beta distribution is often considered as *the* distribution for the modeling of unit interval data [@ferrari_beta_2004; @cribari-neto_beta_2010].
+It has been extensively studied theoretically [@krysicki_new_1999; @gupta_handbook_2014; @espinheira_beta_2008] and is used a baseline to compare other unit interval distributions against [@kieschnick_regression_2003; @bonat_regression_2013; @lopez_bayesian_2013].
+Specifically in Bayesian regression models, it was shown to perform well across a wide range of scenarios [@scholz_posterior_2023; @scholz_prediction_2025], which is why we focus on the beta distribution in this tutorial.
+
+We use a common mean parameterization for the beta distribution with mean parameter $\mu$, bounded between 0 and 1, and positive precision parameter $\phi$ that is roughly proportional to the inverse of the variance $\text{Var}(y) = \frac{\mu \, (1 - \mu)}{\phi + 1}$.
+That is, the larger $\phi$ the smaller the variance of the corresponding beta distribution.
+While the mathematical details are not needed to understand the content of this tutorial, we still write down the density below for reasons of completeness:
+
+$$
+p(y \mid \mu, \phi) = \frac{1}{B(\mu \, \phi, \; (1-\mu) \, \phi)} \; y^{\mu \, \phi -1} \; (1-y)^{(1-\mu) \, \phi -1},
+$$
+
+where $B$ is the Beta function, a complex integral for whose numerical approximation efficient algorithms
+exist in every programming language.
 

### 2025-02-04 manuscript/manuscript.qmd
edit case study 1

@@
       cite-style: authoryear
+execute:
+  cache: true
 ---
@@
 
-Here is what the tibble looks like.
+@tbl-ita-egg shows the first ten rows of the tibble (only relevant columns are included).
 
 ```{r}
-#| label: ita-egg
+#| label: tbl-ita-egg
+#| tbl-cap: "First 10 rows of the data of voicing proportion within stop closure."
+#| echo: false
 
-ita_egg
+ita_egg |> select(speaker, word, vowel, c2, voi_clo_prop) |> head(n = 10) |> knitr::kable()
 ```
@@
 The full summary with an explanation of each part can be found in XXX.
+@tbl-voi-prop-fixef reports the output of `fixef()` as a table (we round all values to the nearest 2 digits for clarity.).
 For each coefficient in the model, `fixef()` prints the name of the coefficient, the mean estimate, the estimate error and the lower and upper limits of a Bayesian Credible interval (CrI).
@@
 To reiterate, in Bayesian inference, an 80% CrI indicates the range of values within which the true estimate falls at 80% probability or confidence.
-We round all values to the nearest 2 digits for clarity.
 
 ```{r}
-#| label: voi-prop-bm-fixef
+#| label: voi-prop-fixef
+#| tbl-cap: "Regression coefficients of a beta regression of voicing proportion (`voi_prop_bm`)."
+#| eval: false
+
+fixef(voi_prop_bm, prob = c(0.1, 0.9))
+```
+
+```{r}
+#| label: tbl-voi-prop-fixef
+#| tbl-cap: "Regression coefficients of a beta regression of voicing proportion (`voi_prop_bm`)."
+#| echo: false
+
 round(
@@
   digits = 2
-)
+) |> knitr::kable()
 ```
@@
 `as_draws_df()` returns a tibble where each column contains the drawn values of a coefficient.
+@tbl-voi-prop-draws shows the first ten rows and first five columns of the output of `as_draws_df()`.
 The probability distribution of the drawn values of each coefficient is the posterior probability distribution of that coefficient.
@@
 voi_prop_bm_draws <- as_draws_df(voi_prop_bm, variable = "^b_", regex = TRUE)
-voi_prop_bm_draws
 ```
 
-We can now wrangle this tibble and plot the posterior distributions for each vowel/C2 combination.
+```{r}
+#| label: tbl-voi-prop-draws
+#| tbl-cap: "First ten rows and 5 columns of the posterior draws for the model `voi_prop_bm`."
+#| echo: false
+
+voi_prop_bm_draws[,1:5] |> head(n = 10) |> knitr::kable()
+```
+
+We can now wrangle this tibble and plot the posterior distributions for each vowel/C2 combination. @tbl-voi-long shows the first ten rows.
 
@@
 #| label: voi-prop-bm-draws-long
+#| warning: false
 
@@
   separate(coeff, into = c("type", "vowel", "c2"))
+```
 
-voi_prop_bm_draws_long
+```{r}
+#| label: tbl-voi-long
+#| tbl-cap: "First ten rows of posterior draws from `voi_prop_bm` in long format."
+#| echo: false
+
+voi_prop_bm_draws_long |> head(n = 10) |> knitr::kable()
 ```
@@
 Let's use 60 and 80% CrIs and transform the log-odds values to proportions with `plogis()`.
+See @fig-voi-prop-bm-1.
 
@@
 #| label: fig-voi-prop-bm-1
+#| fig-cap: "Expected predictions of voicing proportion by vowel and consonant from `voi_prop_bm`."
+#| message: false
 
@@
   )
+```
 
+```{r}
+#| label: fig-voi-pbdl-c2
+#| fig-cap: "Expected predictions of voicing proportion by consonant, averaged across vowels."
 voi_prop_bm_draws_long_c2 |> 
@@
 
-Based on the posterior distributions of the mean voicing proportion by consonant, /p/ has a somewhat higher voicing proportion than /k/ and /t/.
+Based on the expected posterior distributions of the mean voicing proportion by consonant, /p/ has a somewhat higher voicing proportion than /k/ and /t/.
 The real question is: how much higher?
-We can quantify this by taking the difference of the drawn values for /p/ and those for /t, k/.
-Since we want to compare /t, k/ to /p/, we should first average the draws of /t, k/ and then take the difference of the averaged draws and the draws of /p/.
+We can quantify this by taking the difference of the drawn values for /p/ and those for /t, k/ (all averaged across vowels).
+Since we want to compare /t, k/ with /p/, we should first average the average draws of /t, k/ and then take the difference of the averaged /t, k/ draws and the draws of /p/.
+@tbl-voi-pbd shows the first ten rows of the resulting data frame.
+The posterior distribution of the expected difference is shown in @fig-diff-p-tk.
 
@@
   )
+```
 
-voi_prop_bm_diff
+```{r}
+#| label: tbl-voi-pbd
+#| tbl-cap: "First 10 rows of expected difference of voicing proportion between /t, k/ and /p/."
+#| echo: false
+
+voi_prop_bm_diff |> head(n = 10) |> knitr::kable()
 ```
@@
 #| label: fig-diff-p-tk
+#| fig-cap: "Posterior distribution of the difference between the expected voicing proportion of /t, k/ vs /p/, averaged across vowels."
 

### 2025-02-04 manuscript/manuscript.tex
edit case study 1

@@
 
-However, proportion are naturally not Gaussian, since they are limited
-between 0 and 1. The theoretical distribution that generates values with
-this characteristics is the beta distribution. Thus, regression models
-with proportions as the outcome variable should be fitted using a beta
-distribution as the distribution family. This tutorial introduces
-researchers to beta regression models in R using the package brms.
-Familiarity with regression modelling in R with a package like lme4 is
-assumed, but no prior knowledge of Bayesian statistics is necessary.
+However, proportion are usually not Gaussian, since they are limited
+between 0 and 1. Thus, regression models with proportions as outcome
+variable should be fitted using a likelihood family that assumes such
+unit interval data. A common and flexible distribution for this purpose
+is the beta distribution. This tutorial introduces researchers to beta
+regression models in R using the package brms. Familiarity with
+regression modelling in R with a package like lme4 is assumed, but no
+prior knowledge of Bayesian statistics is necessary.
 
@@
 
-\ldots{}
+The beta distribution is often considered as \emph{the} distribution for
+the modeling of unit interval data
+\citep{ferrari_beta_2004, cribari-neto_beta_2010}. It has been
+extensively studied theoretically
+\citep{krysicki_new_1999, gupta_handbook_2014, espinheira_beta_2008} and
+is used a baseline to compare other unit interval distributions against
+\citep{kieschnick_regression_2003, bonat_regression_2013, lopez_bayesian_2013}.
+Specifically in Bayesian regression models, it was shown to perform well
+across a wide range of scenarios
+\citep{scholz_posterior_2023, scholz_prediction_2025}, which is why we
+focus on the beta distribution in this tutorial.
+
+We use a common mean parameterization for the beta distribution with
+mean parameter \(\mu\), bounded between 0 and 1, and positive precision
+parameter \(\phi\) that is roughly proportional to the inverse of the
+variance \(\text{Var}(y) = \frac{\mu \, (1 - \mu)}{\phi + 1}\). That is,
+the larger \(\phi\) the smaller the variance of the corresponding beta
+distribution. While the mathematical details are not needed to
+understand the content of this tutorial, we still write down the density
+below for reasons of completeness:
+
+\[
+p(y \mid \mu, \phi) = \frac{1}{B(\mu \, \phi, \; (1-\mu) \, \phi)} \; y^{\mu \, \phi -1} \; (1-y)^{(1-\mu) \, \phi -1},
+\]
+
+where \(B\) is the Beta function, a complex integral for whose numerical
+approximation efficient algorithms exist in every programming language.
 
@@
 
-Here is what the tibble looks like.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{ita\_egg}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-# A tibble: 2,419 x 53
-   speaker ipu    stimulus    sentence_ons sentence_off word_ons word_off v1_ons
-   <chr>   <chr>  <chr>              <dbl>        <dbl>    <dbl>    <dbl>  <dbl>
- 1 it01    ipu_1  Ripete 'po~         13.2         14.9     13.7     14.1   13.9
- 2 it01    ipu_2  Ripete 'to~         16.9         18.6     17.4     17.9   17.5
- 3 it01    ipu_3  Ripete 'pa~         20.2         21.9     20.7     21.1   20.8
- 4 it01    ipu_4  Sentivo 't~         23.5         25.1     24.0     24.4   24.1
- 5 it01    ipu_5  Sentivo 't~         26.3         27.8     26.8     27.2   26.9
- 6 it01    ipu_6  Scrivete '~         29.2         30.9     29.7     30.1   29.8
- 7 it01    ipu_7  Sentivo 'c~         32.1         33.6     32.6     33.1   32.8
- 8 it01    ipu_8  Scrivete '~         35.0         36.6     35.5     35.9   35.6
- 9 it01    ipu_9  Ha detto '~         41.9         43.5     42.3     42.7   42.4
-10 it01    ipu_10 Sentivo 'p~         47.4         48.9     47.9     48.4   48.1
-# i 2,409 more rows
-# i 45 more variables: c2_ons <dbl>, v2_ons <dbl>, voice_ons <dbl>,
-#   voice_off <dbl>, c1_rel <dbl>, c2_rel <dbl>, stimulus_id <dbl>,
-#   sentence <chr>, word <chr>, c1 <chr>, vowel <fct>, c2 <fct>,
-#   backness <chr>, height <fct>, c1_place <fct>, c2_place <fct>,
-#   v1_duration <dbl>, c2_clos_duration <dbl>, rel_voff <dbl>,
-#   sent_duration <dbl>, speech_rate <dbl>, speech_rate_c <dbl>, ...
-\end{verbatim}
+Table~\ref{tbl-ita-egg} shows the first ten rows of the tibble (only
+relevant columns are included).
+
+\begin{longtable}[]{@{}llllr@{}}
+
+\caption{\label{tbl-ita-egg}First 10 rows of the data of voicing
+proportion within stop closure.}
+
+\tabularnewline
+
+\toprule\noalign{}
+speaker & word & vowel & c2 & voi\_clo\_prop \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+it01 & poto & o & t & 0.0457950 \\
+it01 & topo & o & p & 0.3522513 \\
+it01 & pato & a & t & 0.1440749 \\
+it01 & teto & e & t & 0.3610916 \\
+it01 & toto & o & t & 0.2743519 \\
+it01 & puco & u & k & 0.1943984 \\
+it01 & chipo & i & p & 0.2227896 \\
+it01 & peco & e & k & 0.1835596 \\
+it01 & poco & o & k & 0.1771007 \\
+it01 & poto & o & t & 0.1525163 \\
+
+\end{longtable}
 
@@
 regression coefficients. {[}EXPLAIN EXPECTED PREDICTIONS XXX{]}. The
-full summary with an explanation of each part can be found in XXX. For
+full summary with an explanation of each part can be found in XXX.
+Table~\ref{tbl-voi-prop-fixef} reports the output of \texttt{fixef()} as
+a table (we round all values to the nearest 2 digits for clarity.). For
 each coefficient in the model, \texttt{fixef()} prints the name of the
@@
 an 80\% CrI indicates the range of values within which the true estimate
-falls at 80\% probability or confidence. We round all values to the
-nearest 2 digits for clarity.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{round}\NormalTok{(}
-  \FunctionTok{fixef}\NormalTok{(voi\_prop\_bm, }\AttributeTok{prob =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{)),}
-  \AttributeTok{digits =} \DecValTok{2}
-\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-              Estimate Est.Error   Q10   Q90
-speech_rate_c     0.08      0.06  0.01  0.15
-voweli:c2k       -0.91      0.14 -1.08 -0.74
-vowele:c2k       -1.08      0.11 -1.22 -0.94
-vowela:c2k       -0.99      0.12 -1.14 -0.84
-vowelo:c2k       -0.79      0.14 -0.96 -0.62
-vowelu:c2k       -1.00      0.16 -1.20 -0.80
-voweli:c2t       -0.66      0.11 -0.79 -0.53
-vowele:c2t       -0.84      0.14 -1.02 -0.66
-vowela:c2t       -1.43      0.13 -1.60 -1.26
-vowelo:c2t       -1.15      0.13 -1.31 -0.99
-vowelu:c2t       -0.68      0.12 -0.83 -0.54
-voweli:c2p       -0.68      0.11 -0.81 -0.54
-vowele:c2p       -0.88      0.15 -1.07 -0.68
-vowela:c2p       -1.14      0.13 -1.31 -0.98
-vowelo:c2p       -0.66      0.11 -0.80 -0.53
-vowelu:c2p       -0.44      0.12 -0.59 -0.28
-\end{verbatim}
+falls at 80\% probability or confidence.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{fixef}\NormalTok{(voi\_prop\_bm, }\AttributeTok{prob =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{longtable}[]{@{}lrrrr@{}}
+
+\caption{\label{tbl-voi-prop-fixef}Regression coefficients of a beta
+regression of voicing proportion (\texttt{voi\_prop\_bm}).}
+
+\tabularnewline
+
+\toprule\noalign{}
+& Estimate & Est.Error & Q10 & Q90 \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+speech\_rate\_c & 0.08 & 0.06 & 0.01 & 0.15 \\
+voweli:c2k & -0.91 & 0.14 & -1.08 & -0.74 \\
+vowele:c2k & -1.08 & 0.11 & -1.22 & -0.94 \\
+vowela:c2k & -0.99 & 0.12 & -1.14 & -0.84 \\
+vowelo:c2k & -0.79 & 0.14 & -0.96 & -0.62 \\
+vowelu:c2k & -1.00 & 0.16 & -1.20 & -0.80 \\
+voweli:c2t & -0.66 & 0.11 & -0.79 & -0.53 \\
+vowele:c2t & -0.84 & 0.14 & -1.02 & -0.66 \\
+vowela:c2t & -1.43 & 0.13 & -1.60 & -1.26 \\
+vowelo:c2t & -1.15 & 0.13 & -1.31 & -0.99 \\
+vowelu:c2t & -0.68 & 0.12 & -0.83 & -0.54 \\
+voweli:c2p & -0.68 & 0.11 & -0.81 & -0.54 \\
+vowele:c2p & -0.88 & 0.15 & -1.07 & -0.68 \\
+vowela:c2p & -1.14 & 0.13 & -1.31 & -0.98 \\
+vowelo:c2p & -0.66 & 0.11 & -0.80 & -0.53 \\
+vowelu:c2p & -0.44 & 0.12 & -0.59 & -0.28 \\
+
+\end{longtable}
 
@@
 \texttt{as\_draws\_df()} returns a tibble where each column contains the
-drawn values of a coefficient. The probability distribution of the drawn
+drawn values of a coefficient. Table~\ref{tbl-voi-prop-draws} shows the
+first ten rows and first five columns of the output of
+\texttt{as\_draws\_df()}. The probability distribution of the drawn
 values of each coefficient is the posterior probability distribution of
@@
 \NormalTok{voi\_prop\_bm\_draws }\OtherTok{\textless{}{-}} \FunctionTok{as\_draws\_df}\NormalTok{(voi\_prop\_bm, }\AttributeTok{variable =} \StringTok{"\^{}b\_"}\NormalTok{, }\AttributeTok{regex =} \ConstantTok{TRUE}\NormalTok{)}
-\NormalTok{voi\_prop\_bm\_draws}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-# A draws_df: 1000 iterations, 4 chains, and 16 variables
-   b_speech_rate_c b_voweli:c2k b_vowele:c2k b_vowela:c2k b_vowelo:c2k
-1            0.159        -0.72        -1.05        -1.12        -1.02
-2            0.035        -0.75        -1.39        -1.06        -0.89
-3            0.126        -0.85        -0.98        -1.10        -0.64
-4            0.064        -0.70        -1.07        -1.05        -0.83
-5            0.032        -0.77        -1.18        -0.94        -0.70
-6            0.073        -0.85        -1.23        -1.06        -0.77
-7            0.101        -0.95        -1.07        -1.14        -0.86
-8            0.074        -1.02        -1.07        -1.17        -0.83
-9            0.087        -1.04        -1.07        -1.26        -0.79
-10           0.092        -1.20        -1.04        -0.92        -0.64
-   b_vowelu:c2k b_voweli:c2t b_vowele:c2t
-1         -1.05        -0.62        -0.80
-2         -1.04        -0.67        -0.99
-3         -0.91        -0.66        -0.80
-4         -1.07        -0.53        -0.87
-5         -1.06        -0.63        -0.87
-6         -1.00        -0.69        -0.73
-7         -1.22        -0.74        -1.08
-8         -1.22        -0.73        -1.04
-9         -1.14        -0.66        -0.91
-10        -0.84        -0.63        -0.61
-# ... with 3990 more draws, and 8 more variables
-# ... hidden reserved variables {'.chain', '.iteration', '.draw'}
-\end{verbatim}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+Warning: Dropping 'draws_df' class as required metadata was removed.
+\end{verbatim}
+
+\begin{longtable}[]{@{}
+  >{\raggedleft\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2353}}
+  >{\raggedleft\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1912}}
+  >{\raggedleft\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1912}}
+  >{\raggedleft\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1912}}
+  >{\raggedleft\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1912}}@{}}
+
+\caption{\label{tbl-voi-prop-draws}First ten rows and 5 columns of the
+posterior draws for the model \texttt{voi\_prop\_bm}.}
+
+\tabularnewline
+
+\toprule\noalign{}
+\begin{minipage}[b]{\linewidth}\raggedleft
+b\_speech\_rate\_c
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+b\_voweli:c2k
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+b\_vowele:c2k
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+b\_vowela:c2k
+\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
+b\_vowelo:c2k
+\end{minipage} \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+0.1593890 & -0.7213516 & -1.046577 & -1.1197185 & -1.0191525 \\
+0.0351033 & -0.7492306 & -1.385450 & -1.0633164 & -0.8854320 \\
+0.1259352 & -0.8472495 & -0.975776 & -1.0972735 & -0.6440836 \\
+0.0640492 & -0.7033469 & -1.068626 & -1.0475190 & -0.8328252 \\
+0.0319081 & -0.7659281 & -1.177656 & -0.9357499 & -0.6955130 \\
+0.0729965 & -0.8494392 & -1.226692 & -1.0635364 & -0.7692243 \\
+0.1009609 & -0.9546148 & -1.072955 & -1.1408255 & -0.8566750 \\
+0.0735392 & -1.0160675 & -1.065025 & -1.1736263 & -0.8334884 \\
+0.0870061 & -1.0436475 & -1.069337 & -1.2632061 & -0.7941857 \\
+0.0923815 & -1.2042168 & -1.041238 & -0.9158689 & -0.6444492 \\
+
+\end{longtable}
 
 We can now wrangle this tibble and plot the posterior distributions for
-each vowel/C2 combination.
+each vowel/C2 combination. Table~\ref{tbl-voi-long} shows the first ten
+rows.
 
@@
 
-\begin{verbatim}
-Warning: Dropping 'draws_df' class as required metadata was removed.
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{voi\_prop\_bm\_draws\_long}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-# A tibble: 60,000 x 7
-   .chain .iteration .draw type  vowel  c2     value
-    <int>      <int> <int> <chr> <chr>  <chr>  <dbl>
- 1      1          1     1 b     voweli c2k   -0.721
- 2      1          1     1 b     vowele c2k   -1.05 
- 3      1          1     1 b     vowela c2k   -1.12 
- 4      1          1     1 b     vowelo c2k   -1.02 
- 5      1          1     1 b     vowelu c2k   -1.05 
- 6      1          1     1 b     voweli c2t   -0.625
- 7      1          1     1 b     vowele c2t   -0.795
- 8      1          1     1 b     vowela c2t   -1.40 
- 9      1          1     1 b     vowelo c2t   -1.10 
-10      1          1     1 b     vowelu c2t   -0.668
-# i 59,990 more rows
-\end{verbatim}
+\begin{longtable}[]{@{}rrrlllr@{}}
+
+\caption{\label{tbl-voi-long}First ten rows of posterior draws from
+\texttt{voi\_prop\_bm} in long format.}
+
+\tabularnewline
+
+\toprule\noalign{}
+.chain & .iteration & .draw & type & vowel & c2 & value \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+1 & 1 & 1 & b & voweli & c2k & -0.7213516 \\
+1 & 1 & 1 & b & vowele & c2k & -1.0465766 \\
+1 & 1 & 1 & b & vowela & c2k & -1.1197185 \\
+1 & 1 & 1 & b & vowelo & c2k & -1.0191525 \\
+1 & 1 & 1 & b & vowelu & c2k & -1.0523007 \\
+1 & 1 & 1 & b & voweli & c2t & -0.6248875 \\
+1 & 1 & 1 & b & vowele & c2t & -0.7951552 \\
+1 & 1 & 1 & b & vowela & c2t & -1.3999773 \\
+1 & 1 & 1 & b & vowelo & c2t & -1.0950180 \\
+1 & 1 & 1 & b & vowelu & c2t & -0.6682030 \\
+
+\end{longtable}
 
@@
 and 80\% CrIs and transform the log-odds values to proportions with
-\texttt{plogis()}.
+\texttt{plogis()}. See Figure~\ref{fig-voi-prop-bm-1}.
 
@@
 \FunctionTok{library}\NormalTok{(ggdist)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-
-Attaching package: 'ggdist'
-\end{verbatim}
-
-\begin{verbatim}
-The following objects are masked from 'package:brms':
-
-    dstudent_t, pstudent_t, qstudent_t, rstudent_t
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
+
 \NormalTok{voi\_prop\_bm\_draws\_long }\SpecialCharTok{|\textgreater{}} 
@@
 
-\caption{\label{fig-voi-prop-bm-1}}
+\caption{\label{fig-voi-prop-bm-1}Expected predictions of voicing
+proportion by vowel and consonant from \texttt{voi\_prop\_bm}.}
 
@@
 \NormalTok{  )}
-
+\end{Highlighting}
+\end{Shaded}
+
+\begin{Shaded}
+\begin{Highlighting}[]
 \NormalTok{voi\_prop\_bm\_draws\_long\_c2 }\SpecialCharTok{|\textgreater{}} 
@@
 
-\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/voi-prop-bm-draws-long-c2-1.pdf}}
-
-Based on the posterior distributions of the mean voicing proportion by
-consonant, /p/ has a somewhat higher voicing proportion than /k/ and
-/t/. The real question is: how much higher? We can quantify this by
-taking the difference of the drawn values for /p/ and those for /t, k/.
-Since we want to compare /t, k/ to /p/, we should first average the
-draws of /t, k/ and then take the difference of the averaged draws and
-the draws of /p/.
+\begin{figure}[H]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-voi-pbdl-c2-1.pdf}}
+
+}
+
+\caption{\label{fig-voi-pbdl-c2}Expected predictions of voicing
+proportion by consonant, averaged across vowels.}
+
+\end{figure}%
+
+Based on the expected posterior distributions of the mean voicing
+proportion by consonant, /p/ has a somewhat higher voicing proportion
+than /k/ and /t/. The real question is: how much higher? We can quantify
+this by taking the difference of the drawn values for /p/ and those for
+/t, k/ (all averaged across vowels). Since we want to compare /t, k/
+with /p/, we should first average the average draws of /t, k/ and then
+take the difference of the averaged /t, k/ draws and the draws of /p/.
+Table~\ref{tbl-voi-pbd} shows the first ten rows of the resulting data
+frame. The posterior distribution of the expected difference is shown in
+Figure~\ref{fig-diff-p-tk}.
 
@@
 \NormalTok{  )}
-
-\NormalTok{voi\_prop\_bm\_diff}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-# A tibble: 4,000 x 6
-   .draw    c2k    c2p    c2t   c2tk c2p_tk_diff
-   <int>  <dbl>  <dbl>  <dbl>  <dbl>       <dbl>
- 1     1 -0.992 -0.745 -0.917 -0.953      0.208 
- 2     2 -1.02  -0.790 -1.03  -0.953      0.163 
- 3     3 -0.895 -0.687 -0.890 -0.953      0.266 
- 4     4 -0.944 -0.814 -0.910 -0.953      0.140 
- 5     5 -0.928 -0.791 -0.920 -0.953      0.162 
- 6     6 -0.983 -0.728 -0.926 -0.953      0.225 
- 7     7 -1.05  -0.873 -1.08  -0.953      0.0802
- 8     8 -1.06  -0.875 -1.07  -0.953      0.0782
- 9     9 -1.06  -0.974 -1.05  -0.953     -0.0206
-10    10 -0.929 -0.829 -0.953 -0.953      0.125 
-# i 3,990 more rows
-\end{verbatim}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{longtable}[]{@{}rrrrrr@{}}
+
+\caption{\label{tbl-voi-pbd}First 10 rows of expected difference of
+voicing proportion between /t, k/ and /p/.}
+
+\tabularnewline
+
+\toprule\noalign{}
+.draw & c2k & c2p & c2t & c2tk & c2p\_tk\_diff \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+1 & -0.9918200 & -0.7450832 & -0.9166482 & -0.9533426 & 0.2082595 \\
+2 & -1.0240673 & -0.7904348 & -1.0322325 & -0.9533426 & 0.1629079 \\
+3 & -0.8954946 & -0.6873228 & -0.8904412 & -0.9533426 & 0.2660199 \\
+4 & -0.9442361 & -0.8135191 & -0.9095122 & -0.9533426 & 0.1398235 \\
+5 & -0.9278262 & -0.7908457 & -0.9200903 & -0.9533426 & 0.1624970 \\
+6 & -0.9825895 & -0.7283054 & -0.9262196 & -0.9533426 & 0.2250372 \\
+7 & -1.0482359 & -0.8731592 & -1.0841208 & -0.9533426 & 0.0801835 \\
+8 & -1.0613713 & -0.8751219 & -1.0741555 & -0.9533426 & 0.0782208 \\
+9 & -1.0622522 & -0.9739388 & -1.0481834 & -0.9533426 & -0.0205962 \\
+10 & -0.9293998 & -0.8287910 & -0.9526105 & -0.9533426 & 0.1245517 \\
+
+\end{longtable}
 
@@
 
-\caption{\label{fig-diff-p-tk}}
+\caption{\label{fig-diff-p-tk}Posterior distribution of the difference
+between the expected voicing proportion of /t, k/ vs /p/, averaged
+across vowels.}
 
@@
 
-  \bibliography{linguistics.bib}
+  \bibliography{linguistics.bib,statistics.bib}
 

### 2025-02-05 manuscript/manuscript.qmd
use tables in manuscript

@@
 library(collapse)
+library(marginaleffects)
 ```
@@
 #| label: voi-prop-fixef
-#| tbl-cap: "Regression coefficients of a beta regression of voicing proportion (`voi_prop_bm`)."
 #| eval: false
@@
 We can convert these into odd-ratios with `exp()`.
-odd-ratios indicate the ratio of the difference between A and B, so that 1 means no difference, values greater than 1 indicate an increase in A relative to B and values lower than 1 indicate a decrease in A relative to B.
+The output of the code below is shown in @tbl-voi-bm-quant.
+Odd-ratios indicate the ratio of the difference between A and B, so that 1 means no difference, values greater than 1 indicate an increase in A relative to B and values lower than 1 indicate a decrease in A relative to B.
 odd-ratios are useful when looking at differences that are in log-odds because while the relative magnitude of the difference in proportion between two groups is the same independent of the baseline proportion, the *absolute* magnitude of the difference depends on the baseline value.
@@
 #| label: voi-prop-bm-diff-cri
+#| message: false
 
@@
 
-voi_prop_bm_diff |> 
+voi_bm_quant <- voi_prop_bm_diff |> 
   mutate(c2p_tk_diff_ratio = exp(c2p_tk_diff)) |> 
@@
 
-So, based on the model and data, there is a 90% probability that the voicing proportion in /p/ is 1.07-1.38 times longer (or 7-38% increase) than in /t, k/.
+```{r}
+#| label: tbl-voi-bm-quant
+#| tbl-cap: "Upper and lower limits of 90, 80 and 60% Credible Intervals of the difference ratio of voicing proportion in /t, k/ vs /p/."
+#| echo: false
+
+voi_bm_quant |> knitr::kable()
+```
+
+Based on the model and data, there is a 90% probability that the voicing proportion in /p/ is 1.07-1.38 times longer (or 7-38% increase) than in /t, k/.
 At 80% confidence, the change ratio is 1.10-1.34 (or 10-34% increase) while at 60% confidence is 1.14-1.30 (14-30% increase).
 In other words we can be quite confident that the voicing proportion in /p/ is longer than in /t, k/ and that the increase is less than 35%.
-
-brms comes with a convenient function, `conditional_effects()`, to plot posterior means and CrIs based on predictors in the model.
-In the following example, we plot the predicted proportion of voicing by consonant and vowel.
+The brms package comes with a convenient function, `conditional_effects()`, to plot posterior means and CrIs based on predictors in the model.
+In @fig-voi-prop-bm-cond, we plot the predicted proportion of voicing by consonant and vowel.
 
 ```{r}
-#| label: voi-prop-bm-cond
+#| label: fig-voi-prop-bm-cond
+#| fig-cap: "Expected voicing proportion by consonant and vowel with 95% Credible Intervals."
 conditional_effects(voi_prop_bm, "c2:vowel")
@@
 ```{r}
-#| label: marginal
+#| label: marginal-code
+#| eval: false
 
@@
 avg_predictions(voi_prop_bm, variables = "vowel", conf_level = 0.8)
+```
+
+```{r}
+#| label: tbl-marginal-1
+#| tbl-cap: "Difference in expected voicing proportion by consonant, averaged across vowels, with 80% CrIs."
+#| echo: false
+
+avg_comparisons(voi_prop_bm, variables = list(c2 = "pairwise"), conf_level = 0.8, type = "link") |>
+  mutate(across(where(is.numeric), ~round(.x, 3))) |>
+  select(-(predicted_lo:tmp_idx)) |> 
+  knitr::kable()
+```
+
+```{r}
+#| label: tbl-marginal-2
+#| tbl-cap: "Expected voicing proportion by vowel, averaged across vowels, with 80% CrIs."
+#| echo: false
+
+avg_predictions(voi_prop_bm, variables = "vowel", conf_level = 0.8) |>
+  mutate(across(where(is.numeric), ~round(.x, 3))) |>
+  knitr::kable()
 ```
@@
 
-First, let's read and plot the data.
-
-```{r}
-#| label: nasal
-#| message: false
-#| warning: false
-
-nasal <- read_csv("data/carignan2021/nasal.csv")
-nasal
-```
+@tbl-nasal shows the first ten rows of the data. The data contains the following columns:
 
@@
 -   `nas_prop` is the proportion of coarticulatory nasalisation of the vowel.
+
+```{r}
+#| label: nasal
+#| message: false
+#| warning: false
+
+nasal <- read_csv("data/carignan2021/nasal.csv")
+```
+
+```{r}
+#| label: tbl-nasal
+#| tbl-cap: "First ten rows of the nasal proportion data."
+#| echo: false
+
+nasal |> head(n = 10) |> knitr::kable()
+```
 
@@
 
-Let's inspect the output of `fixef()`.
+Let's inspect the output of `fixef()`, reported in @tbl-nas-prop-fixef.
 
@@
 #| label: nas-prop-bm-fixef
+#| eval: false
+
+fixef(nas_prop_bm, prob = c(0.1, 0.9))
+```
+
+```{r}
+#| label: tbl-nas-prop-fixef
+#| tbl-cap: "Regression coefficients of a beta regression of nasalisation proportion (`nas_prop_bm`)."
+#| echo: false
+
 round(
@@
 However there is vowel-specific variation, and there doesn't seem to be much of a difference in nasalisation proportion in /e/ and /u/.
-Plotting the expected predictions with `conditional_effects()` should make this clearer.
+@fig-nas-prop shows the expected predictions with `conditional_effects()`, which should make the interpretation clearer.
 
 ```{r}
-#| label: nas-prop
+#| label: fig-nas-prop
+#| fig-cap: "Expected nasalisation proportion by vowel and voicing, with 95% CrIs."
 
@@
 
-To answer question 1, we can calculate the average difference in nasalisation proportion by first calculating the average nasalisation across all vowels for voiced and voiceless sequences and then take the difference of those, similarly to what we have done in the Case Study 1 above.
+To answer question 1, we can calculate the average difference in nasalisation proportion by first calculating the average nasalisation across all vowels for voiced and voiceless sequences (see @tbl-nas-draws for the output of this step) and then take the difference of those, similarly to what we have done in the Case Study 1 above.
 
@@
 #| label: nas-prop-bm-draws
+#| warning: false
 
@@
   separate(coeff, into = c("type", "voicing", "vowel"))
+```
 
-nas_prop_bm_draws_long
+```{r}
+#| label: tbl-nas-draws
+#| tbl-cap: "First ten rows of posterior draws from `nas_prop_bm` in long format."
+#| echo: false
+
+nas_prop_bm_draws_long |> head(n = 10) |> knitr::kable()
 ```
@@
 Note that, as discussed for Case study 1, when working wit log-odds it is important to first do all necessary calculations in log-odds, here calculate the mean log-odds across vowels, and *then* transform the calculated estimands to proportions/probabilities.
+The output of the following code is show in @tbl-nas-long-voi and the density plot of the calculated draws is in @fig-nas-prop-bm-draws-long-voicing-plot.
 
@@
   )
-
-nas_prop_bm_draws_long_voicing
 ```
@@
 ```{r}
-#| label: nas-prop-bm-draws-long-voicing-plot
+#| label: tbl-nas-long-voi
+#| tbl-cap: "Mean nasalisation proportion by draw and voicing, averaged across vowels (first 10 rows)."
+#| echo: false
+
+nas_prop_bm_draws_long_voicing |> head(n = 10) |> knitr::kable()
+```
+
+```{r}
+#| label: fig-nas-prop-bm-draws-long-voicing-plot
+#| fig-cap: "Expected nasalisation proportion by voicing, averaged across vowels."
 
@@
 we will convert log-odds to odd-ratios using the `exp()` (exponential) function (and as before we first calculate the difference and then exponentiate the resulting values, after which we can take summary measures, like means and quantile-based measures such as CrIs).
+tbl-nas-prop-bm-diff-quant show the 90, 80 and 60% CrIs of the difference ratio of nasalisation proportion in voiceless vs voiced NC sequences.
 
@@
   )
-nas_prop_bm_diff
-```
 
-```{r}
-#| label: nas-prop-bm-diff-cris
-
-nas_prop_bm_diff |> 
+nas_prop_bm_diff_quant <- nas_prop_bm_diff |> 
   reframe(
@@
 
-The CrIs of the ratio difference in nasalisation proportion in voiceless vs voiced NC sequences suggest a robust increase of nasalisation in the voiceless NC sequences, with a 90% probability that the increase is between 23% and 69% of the proportion in voiced NC sequences.
+```{r}
+#| label: tbl-nas-prop-bm-diff-quant
+#| tbl-cap: "Upper and lower limits of 90, 80 and 60% Credible Intervals of the difference ratio of nasalisation proportion in vowels followed by voiced vs voiceless consonants."
+#| echo: false
+
+nas_prop_bm_diff_quant |> knitr::kable()
+
+```
+
+The CrIs of the ratio difference in nasalisation proportion in voiceless vs voiced NC sequences suggest an increase of nasalisation in the voiceless NC sequences, with a 90% probability that the increase is between 23% and 69% of the proportion in voiced NC sequences.
 
@@
   mutate(voicing_diff = voicingvoiceless - voicingvoiced)
-
-nas_prop_r
 ```
@@
 #| label: fig-nas-prop-r
+#| fig-cap: "Expected log-odd difference of coarticulatory nasalisation in voiceless vs voiced NC sequences for each speaker."
 #| warning: false

### 2025-02-05 manuscript/manuscript.tex
use tables in manuscript

@@
 that the values of the difference are in log-odds! We can convert these
-into odd-ratios with \texttt{exp()}. odd-ratios indicate the ratio of
+into odd-ratios with \texttt{exp()}. The output of the code below is
+shown in Table~\ref{tbl-voi-bm-quant}. Odd-ratios indicate the ratio of
 the difference between A and B, so that 1 means no difference, values
@@
 \FunctionTok{library}\NormalTok{(posterior)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-This is posterior version 1.6.0
-\end{verbatim}
-
-\begin{verbatim}
-
-Attaching package: 'posterior'
-\end{verbatim}
-
-\begin{verbatim}
-The following objects are masked from 'package:stats':
-
-    mad, sd, var
-\end{verbatim}
-
-\begin{verbatim}
-The following objects are masked from 'package:base':
-
-    %in%, match
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{voi\_prop\_bm\_diff }\SpecialCharTok{|\textgreater{}} 
+
+\NormalTok{voi\_bm\_quant }\OtherTok{\textless{}{-}}\NormalTok{ voi\_prop\_bm\_diff }\SpecialCharTok{|\textgreater{}} 
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{c2p\_tk\_diff\_ratio =} \FunctionTok{exp}\NormalTok{(c2p\_tk\_diff)) }\SpecialCharTok{|\textgreater{}} 
@@
 
-\begin{verbatim}
-# A tibble: 2 x 3
-    q90   q80   q60
-  <dbl> <dbl> <dbl>
-1  1.07  1.1   1.14
-2  1.38  1.34  1.3 
-\end{verbatim}
-
-So, based on the model and data, there is a 90\% probability that the
+\begin{longtable}[]{@{}rrr@{}}
+
+\caption{\label{tbl-voi-bm-quant}Upper and lower limits of 90, 80 and
+60\% Credible Intervals of the difference ratio of voicing proportion in
+/t, k/ vs /p/.}
+
+\tabularnewline
+
+\toprule\noalign{}
+q90 & q80 & q60 \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+1.07 & 1.10 & 1.14 \\
+1.38 & 1.34 & 1.30 \\
+
+\end{longtable}
+
+Based on the model and data, there is a 90\% probability that the
 voicing proportion in /p/ is 1.07-1.38 times longer (or 7-38\% increase)
@@
 proportion in /p/ is longer than in /t, k/ and that the increase is less
-than 35\%.
-
-brms comes with a convenient function, \texttt{conditional\_effects()},
-to plot posterior means and CrIs based on predictors in the model. In
-the following example, we plot the predicted proportion of voicing by
-consonant and vowel.
+than 35\%. The brms package comes with a convenient function,
+\texttt{conditional\_effects()}, to plot posterior means and CrIs based
+on predictors in the model. In Figure~\ref{fig-voi-prop-bm-cond}, we
+plot the predicted proportion of voicing by consonant and vowel.
 
@@
 
-\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/voi-prop-bm-cond-1.pdf}}
+\begin{figure}[H]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-voi-prop-bm-cond-1.pdf}}
+
+}
+
+\caption{\label{fig-voi-prop-bm-cond}Expected voicing proportion by
+consonant and vowel with 95\% Credible Intervals.}
+
+\end{figure}%
 
@@
 \FunctionTok{avg\_comparisons}\NormalTok{(voi\_prop\_bm, }\AttributeTok{variables =} \FunctionTok{list}\NormalTok{(}\AttributeTok{c2 =} \StringTok{"pairwise"}\NormalTok{), }\AttributeTok{conf\_level =} \FloatTok{0.8}\NormalTok{, }\AttributeTok{type =} \StringTok{"link"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-
-          Contrast Estimate   10.0 % 90.0 %
- mean(t) - mean(k)   0.0348 -0.00418 0.0738
- mean(p) - mean(k)   0.2176  0.17837 0.2558
- mean(p) - mean(t)   0.1829  0.14474 0.2192
-
-Term: c2
-Type:  link 
-Columns: term, contrast, estimate, conf.low, conf.high, predicted_lo, predicted_hi, predicted, tmp_idx 
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
 \FunctionTok{avg\_predictions}\NormalTok{(voi\_prop\_bm, }\AttributeTok{variables =} \StringTok{"vowel"}\NormalTok{, }\AttributeTok{conf\_level =} \FloatTok{0.8}\NormalTok{)}
@@
 
-\begin{verbatim}
-
- vowel Estimate 10.0 % 90.0 %
-     i    0.331  0.325  0.338
-     e    0.300  0.293  0.307
-     a    0.245  0.238  0.252
-     o    0.305  0.298  0.312
-     u    0.348  0.341  0.354
-
-Type:  response 
-Columns: vowel, estimate, conf.low, conf.high 
-\end{verbatim}
+\begin{longtable}[]{@{}llrrr@{}}
+
+\caption{\label{tbl-marginal-1}Difference in expected voicing proportion
+by consonant, averaged across vowels, with 80\% CrIs.}
+
+\tabularnewline
+
+\toprule\noalign{}
+term & contrast & estimate & conf.low & conf.high \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+c2 & mean(t) - mean(k) & 0.035 & -0.004 & 0.074 \\
+c2 & mean(p) - mean(k) & 0.218 & 0.178 & 0.256 \\
+c2 & mean(p) - mean(t) & 0.183 & 0.145 & 0.219 \\
+
+\end{longtable}
+
+\begin{longtable}[]{@{}lrrr@{}}
+
+\caption{\label{tbl-marginal-2}Expected voicing proportion by vowel,
+averaged across vowels, with 80\% CrIs.}
+
+\tabularnewline
+
+\toprule\noalign{}
+vowel & estimate & conf.low & conf.high \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+i & 0.331 & 0.325 & 0.338 \\
+e & 0.300 & 0.293 & 0.307 \\
+a & 0.245 & 0.238 & 0.252 \\
+o & 0.305 & 0.298 & 0.312 \\
+u & 0.348 & 0.341 & 0.354 \\
+
+\end{longtable}
 
@@
 
-First, let's read and plot the data.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{nasal }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"data/carignan2021/nasal.csv"}\NormalTok{)}
-\NormalTok{nasal}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-# A tibble: 705 x 6
-   speaker label               vowel NC    voicing   nas_prop
-   <chr>   <chr>               <chr> <chr> <chr>        <dbl>
- 1 S03     b__U_nt_@___N_B17/s u     nt    voiceless  0.367  
- 2 S03     b__a_nd_@___N_B19/s a     nd    voiced     0.195  
- 3 S03     b__a_nt_@___N_B15/s a     nt    voiceless  0.279  
- 4 S03     f__I_nt_@___N_B05/s i     nt    voiceless  0.764  
- 5 S03     l__I_nd_@___N_B06/s i     nd    voiced     0.00529
- 6 S03     p__E_nt_____N_B09/s e     nt    voiceless  0.335  
- 7 S03     r__a_nt_@___N_B06/s a     nt    voiceless  0.243  
- 8 S03     v__I_nd_@___N_B07/s i     nd    voiced     0.0248 
- 9 S03     v__I_nt_6___N_B15/s i     nt    voiceless  0.135  
-10 S03     z__E_nd_@___N_B17/s e     nd    voiced     0.538  
-# i 695 more rows
-\end{verbatim}
+Table~\ref{tbl-nasal} shows the first ten rows of the data. The data
+contains the following columns:
 
@@
 
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{nasal }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"data/carignan2021/nasal.csv"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{longtable}[]{@{}lllllr@{}}
+
+\caption{\label{tbl-nasal}First ten rows of the nasal proportion data.}
+
+\tabularnewline
+
+\toprule\noalign{}
+speaker & label & vowel & NC & voicing & nas\_prop \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+S03 & b\_\emph{U\_nt}@\_\_\_N\_B17/s & u & nt & voiceless & 0.3668820 \\
+S03 & b\_\emph{a\_nd}@\_\_\_N\_B19/s & a & nd & voiced & 0.1954858 \\
+S03 & b\_\emph{a\_nt}@\_\_\_N\_B15/s & a & nt & voiceless & 0.2786485 \\
+S03 & f\_\emph{I\_nt}@\_\_\_N\_B05/s & i & nt & voiceless & 0.7642259 \\
+S03 & l\_\emph{I\_nd}@\_\_\_N\_B06/s & i & nd & voiced & 0.0052949 \\
+S03 & p\_\emph{E\_nt}\_\_\_\_N\_B09/s & e & nt & voiceless &
+0.3347331 \\
+S03 & r\_\emph{a\_nt}@\_\_\_N\_B06/s & a & nt & voiceless & 0.2431760 \\
+S03 & v\_\emph{I\_nd}@\_\_\_N\_B07/s & i & nd & voiced & 0.0247572 \\
+S03 & v\_\_I\_nt\_6\_\_\_N\_B15/s & i & nt & voiceless & 0.1350081 \\
+S03 & z\_\emph{E\_nd}@\_\_\_N\_B17/s & e & nd & voiced & 0.5378522 \\
+
+\end{longtable}
+
 \begin{figure}
@@
 
-Let's inspect the output of \texttt{fixef()}.
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{round}\NormalTok{(}
-  \FunctionTok{fixef}\NormalTok{(nas\_prop\_bm, }\AttributeTok{prob =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{)),}
-  \AttributeTok{digits =} \DecValTok{2}
-\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-                        Estimate Est.Error   Q10   Q90
-voicingvoiced:vowela       -0.61      0.12 -0.77 -0.46
-voicingvoiceless:vowela    -0.08      0.09 -0.19  0.04
-voicingvoiced:vowele       -0.19      0.16 -0.38  0.01
-voicingvoiceless:vowele    -0.25      0.10 -0.38 -0.12
-voicingvoiced:voweli       -0.74      0.19 -0.98 -0.51
-voicingvoiceless:voweli    -0.12      0.18 -0.35  0.10
-voicingvoiced:vowelo       -0.34      0.16 -0.54 -0.14
-voicingvoiceless:vowelo     0.25      0.15  0.05  0.44
-voicingvoiced:vowelu       -0.71      0.18 -0.94 -0.49
-voicingvoiceless:vowelu    -0.58      0.16 -0.78 -0.38
-\end{verbatim}
+Let's inspect the output of \texttt{fixef()}, reported in
+Table~\ref{tbl-nas-prop-fixef}.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{fixef}\NormalTok{(nas\_prop\_bm, }\AttributeTok{prob =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.1}\NormalTok{, }\FloatTok{0.9}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{longtable}[]{@{}lrrrr@{}}
+
+\caption{\label{tbl-nas-prop-fixef}Regression coefficients of a beta
+regression of nasalisation proportion (\texttt{nas\_prop\_bm}).}
+
+\tabularnewline
+
+\toprule\noalign{}
+& Estimate & Est.Error & Q10 & Q90 \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+voicingvoiced:vowela & -0.61 & 0.12 & -0.77 & -0.46 \\
+voicingvoiceless:vowela & -0.08 & 0.09 & -0.19 & 0.04 \\
+voicingvoiced:vowele & -0.19 & 0.16 & -0.38 & 0.01 \\
+voicingvoiceless:vowele & -0.25 & 0.10 & -0.38 & -0.12 \\
+voicingvoiced:voweli & -0.74 & 0.19 & -0.98 & -0.51 \\
+voicingvoiceless:voweli & -0.12 & 0.18 & -0.35 & 0.10 \\
+voicingvoiced:vowelo & -0.34 & 0.16 & -0.54 & -0.14 \\
+voicingvoiceless:vowelo & 0.25 & 0.15 & 0.05 & 0.44 \\
+voicingvoiced:vowelu & -0.71 & 0.18 & -0.94 & -0.49 \\
+voicingvoiceless:vowelu & -0.58 & 0.16 & -0.78 & -0.38 \\
+
+\end{longtable}
 
@@
 variation, and there doesn't seem to be much of a difference in
-nasalisation proportion in /e/ and /u/. Plotting the expected
-predictions with \texttt{conditional\_effects()} should make this
-clearer.
+nasalisation proportion in /e/ and /u/. Figure~\ref{fig-nas-prop} shows
+the expected predictions with \texttt{conditional\_effects()}, which
+should make the interpretation clearer.
 
@@
 
-\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/nas-prop-1.pdf}}
+\begin{figure}[H]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-nas-prop-1.pdf}}
+
+}
+
+\caption{\label{fig-nas-prop}Expected nasalisation proportion by vowel
+and voicing, with 95\% CrIs.}
+
+\end{figure}%
 
@@
 nasalisation proportion by first calculating the average nasalisation
-across all vowels for voiced and voiceless sequences and then take the
+across all vowels for voiced and voiceless sequences (see
+Table~\ref{tbl-nas-draws} for the output of this step) and then take the
 difference of those, similarly to what we have done in the Case Study 1
@@
 
-\begin{verbatim}
-Warning: Dropping 'draws_df' class as required metadata was removed.
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{nas\_prop\_bm\_draws\_long}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-# A tibble: 40,000 x 7
-   .chain .iteration .draw type  voicing          vowel    value
-    <int>      <int> <int> <chr> <chr>            <chr>    <dbl>
- 1      1          1     1 b     voicingvoiced    vowela -0.463 
- 2      1          1     1 b     voicingvoiceless vowela -0.0944
- 3      1          1     1 b     voicingvoiced    vowele -0.155 
- 4      1          1     1 b     voicingvoiceless vowele -0.208 
- 5      1          1     1 b     voicingvoiced    voweli -0.481 
- 6      1          1     1 b     voicingvoiceless voweli -0.0529
- 7      1          1     1 b     voicingvoiced    vowelo -0.310 
- 8      1          1     1 b     voicingvoiceless vowelo  0.321 
- 9      1          1     1 b     voicingvoiced    vowelu -0.620 
-10      1          1     1 b     voicingvoiceless vowelu -0.368 
-# i 39,990 more rows
-\end{verbatim}
+\begin{longtable}[]{@{}rrrlllr@{}}
+
+\caption{\label{tbl-nas-draws}First ten rows of posterior draws from
+\texttt{nas\_prop\_bm} in long format.}
+
+\tabularnewline
+
+\toprule\noalign{}
+.chain & .iteration & .draw & type & voicing & vowel & value \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+1 & 1 & 1 & b & voicingvoiced & vowela & -0.4632870 \\
+1 & 1 & 1 & b & voicingvoiceless & vowela & -0.0943699 \\
+1 & 1 & 1 & b & voicingvoiced & vowele & -0.1546258 \\
+1 & 1 & 1 & b & voicingvoiceless & vowele & -0.2080810 \\
+1 & 1 & 1 & b & voicingvoiced & voweli & -0.4810383 \\
+1 & 1 & 1 & b & voicingvoiceless & voweli & -0.0529377 \\
+1 & 1 & 1 & b & voicingvoiced & vowelo & -0.3096219 \\
+1 & 1 & 1 & b & voicingvoiceless & vowelo & 0.3212456 \\
+1 & 1 & 1 & b & voicingvoiced & vowelu & -0.6198312 \\
+1 & 1 & 1 & b & voicingvoiceless & vowelu & -0.3680900 \\
+
+\end{longtable}
 
@@
 log-odds across vowels, and \emph{then} transform the calculated
-estimands to proportions/probabilities.
+estimands to proportions/probabilities. The output of the following code
+is show in Table~\ref{tbl-nas-long-voi} and the density plot of the
+calculated draws is in
+Figure~\ref{fig-nas-prop-bm-draws-long-voicing-plot}.
 
@@
 \NormalTok{  )}
-
-\NormalTok{nas\_prop\_bm\_draws\_long\_voicing}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-# A tibble: 8,000 x 4
-   .draw voicing          value_mean value_mean_prop
-   <int> <chr>                 <dbl>           <dbl>
- 1     1 voicingvoiced       -0.406            0.400
- 2     1 voicingvoiceless    -0.0804           0.480
- 3     2 voicingvoiced       -0.391            0.404
- 4     2 voicingvoiceless    -0.123            0.469
- 5     3 voicingvoiced       -0.444            0.391
- 6     3 voicingvoiceless    -0.132            0.467
- 7     4 voicingvoiced       -0.439            0.392
- 8     4 voicingvoiceless    -0.0982           0.475
- 9     5 voicingvoiced       -0.383            0.405
-10     5 voicingvoiceless    -0.266            0.434
-# i 7,990 more rows
-\end{verbatim}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{longtable}[]{@{}rlrr@{}}
+
+\caption{\label{tbl-nas-long-voi}Mean nasalisation proportion by draw
+and voicing, averaged across vowels (first 10 rows).}
+
+\tabularnewline
+
+\toprule\noalign{}
+.draw & voicing & value\_mean & value\_mean\_prop \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+1 & voicingvoiced & -0.4056808 & 0.3999482 \\
+1 & voicingvoiceless & -0.0804466 & 0.4798992 \\
+2 & voicingvoiced & -0.3905744 & 0.4035790 \\
+2 & voicingvoiceless & -0.1227124 & 0.4693603 \\
+3 & voicingvoiced & -0.4436082 & 0.3908816 \\
+3 & voicingvoiceless & -0.1320190 & 0.4670431 \\
+4 & voicingvoiced & -0.4385607 & 0.3920840 \\
+4 & voicingvoiceless & -0.0981768 & 0.4754755 \\
+5 & voicingvoiced & -0.3827134 & 0.4054726 \\
+5 & voicingvoiceless & -0.2657250 & 0.4339569 \\
+
+\end{longtable}
 
@@
 
-\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/nas-prop-bm-draws-long-voicing-plot-1.pdf}}
+\begin{figure}[H]
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-nas-prop-bm-draws-long-voicing-plot-1.pdf}}
+
+}
+
+\caption{\label{fig-nas-prop-bm-draws-long-voicing-plot}Expected
+nasalisation proportion by voicing, averaged across vowels.}
+
+\end{figure}%
 
@@
 measures, like means and quantile-based measures such as CrIs).
+tbl-nas-prop-bm-diff-quant show the 90, 80 and 60\% CrIs of the
+difference ratio of nasalisation proportion in voiceless vs voiced NC
+sequences.
 
@@
 \NormalTok{  )}
-\NormalTok{nas\_prop\_bm\_diff}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-# A tibble: 4,000 x 5
-   .draw voicingvoiced voicingvoiceless voicing_diff voicing_diff_ratio
-   <int>         <dbl>            <dbl>        <dbl>              <dbl>
- 1     1        -0.406          -0.0804        0.325               1.38
- 2     2        -0.391          -0.123         0.268               1.31
- 3     3        -0.444          -0.132         0.312               1.37
- 4     4        -0.439          -0.0982        0.340               1.41
- 5     5        -0.383          -0.266         0.117               1.12
- 6     6        -0.541          -0.0903        0.450               1.57
- 7     7        -0.512          -0.111         0.401               1.49
- 8     8        -0.493          -0.152         0.341               1.41
- 9     9        -0.527          -0.175         0.352               1.42
-10    10        -0.491          -0.257         0.234               1.26
-# i 3,990 more rows
-\end{verbatim}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{nas\_prop\_bm\_diff }\SpecialCharTok{|\textgreater{}} 
+
+\NormalTok{nas\_prop\_bm\_diff\_quant }\OtherTok{\textless{}{-}}\NormalTok{ nas\_prop\_bm\_diff }\SpecialCharTok{|\textgreater{}} 
   \FunctionTok{reframe}\NormalTok{(}
@@
 
-\begin{verbatim}
-# A tibble: 2 x 3
-    q90   q80   q60
-  <dbl> <dbl> <dbl>
-1  1.23  1.27  1.33
-2  1.69  1.63  1.56
-\end{verbatim}
+\begin{longtable}[]{@{}rrr@{}}
+
+\caption{\label{tbl-nas-prop-bm-diff-quant}Upper and lower limits of 90,
+80 and 60\% Credible Intervals of the difference ratio of nasalisation
+proportion in vowels followed by voiced vs voiceless consonants.}
+
+\tabularnewline
+
+\toprule\noalign{}
+q90 & q80 & q60 \\
+\midrule\noalign{}
+\endhead
+\bottomrule\noalign{}
+\endlastfoot
+1.23 & 1.27 & 1.33 \\
+1.69 & 1.63 & 1.56 \\
+
+\end{longtable}
 
 The CrIs of the ratio difference in nasalisation proportion in voiceless
-vs voiced NC sequences suggest a robust increase of nasalisation in the
+vs voiced NC sequences suggest an increase of nasalisation in the
 voiceless NC sequences, with a 90\% probability that the increase is
@@
   \FunctionTok{mutate}\NormalTok{(}\AttributeTok{voicing\_diff =}\NormalTok{ voicingvoiceless }\SpecialCharTok{{-}}\NormalTok{ voicingvoiced)}
-
-\NormalTok{nas\_prop\_r}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{verbatim}
-# A tibble: 140,000 x 5
-# Groups:   .draw, speaker [140,000]
-   .draw speaker voicingvoiced voicingvoiceless voicing_diff
-   <int> <chr>           <dbl>            <dbl>        <dbl>
- 1     1 S03          -1.42              0.0676      1.49   
- 2     1 S04           0.209             0.185      -0.0238 
- 3     1 S05          -0.284            -0.0546      0.229  
- 4     1 S06           0.00669          -0.324      -0.331  
- 5     1 S07          -0.446            -0.147       0.299  
- 6     1 S08          -0.610            -0.127       0.483  
- 7     1 S09          -0.473             0.0497      0.523  
- 8     1 S10          -0.556            -0.311       0.245  
- 9     1 S11           0.201            -0.426      -0.628  
-10     1 S12          -0.0812           -0.0856     -0.00439
-# i 139,990 more rows
-\end{verbatim}
+\end{Highlighting}
+\end{Shaded}
 
@@
 
-\caption{\label{fig-nas-prop-r}}
+\caption{\label{fig-nas-prop-r}Expected log-odd difference of
+coarticulatory nasalisation in voiceless vs voiced NC sequences for each
+speaker.}
 

### 2025-02-06 manuscript/manuscript.qmd
edit manuscript

@@
 
-Phonetic research often involves numeric continuous outcome variables, like durations, frequencies, loudness and ratios.
-Another commonly employed type of outcome variable are proportions: for example, proportion of voicing during closure, vocal folds contact quotient, gesture amplitude, nasalance.
+Phonetic research often involves numeric continuous outcome variables, like durations, frequencies, loudness and ratios [@kluender1988; @johnson2003; @gick2013; @ladefoged2014; @abramson2017; @roettger2018; @coretta2023].
+Another commonly employed type of outcome variable are proportions: for example, proportion of voicing during closure [@davidson2016], vocal folds contact quotient [@herbst2017], gesture amplitude [@carignan2021], nasalance [@carignan2021].
 Moreover, virtually any measure can be MIN-MAX normalised, a procedure which transforms values so that they are in the range 0--1.
 
-Regression models are very common, but there is a tendency of using Gaussian distribution families (i.e.
-probability distributions for the outcome variable) for anything that is numeric.
-A possible reason is that the base R function for fitting regression models, `lm()`, and the lme4 function used to fit regression models with varying terms, `lmer()`, both fit Gaussian regressions by default and the user does not have to specify the distribution family.
-This tacit defaulting to Gaussian models is also reflected in teaching practices, where test and models using the Gaussian distribution are the first to be taught, due to their relative simplicity and the fact that other models are generalisations of Gaussian models.
+Regression models (and extensions like generalised additive models) have become a *de facto* standard for the statistical analysis of a variety of measures in phonetic research [@kirby2018; @politzerahles2018; @tavakoli2024].
+However, there is a tendency for researchers to use Gaussian distribution families (i.e.
+probability distributions for the outcome variable) for any measure that is numeric, irrespective of whether the measure is unbounded, like in truly Gaussian variables, or bounded, like in proportions.
+A possible reason is that the base R function for fitting regression models, `lm()`, and the lme4 function used to fit regression models with varying terms, `lmer()` from lme4 [@bates2015], both fit Gaussian regressions by default and the user does not have to specify the distribution family.
+This tacit default of using Gaussian models is also reflected in teaching practices, where significance test and models using the Gaussian distribution are the first to be taught [@baayen2008; @winter2020], due to their relative simplicity and the fact that regression models with other families are conceptual generalisations of Gaussian regression models.
 
-However, proportion are usually not Gaussian, since they are limited between 0 and 1.
+While most researchers approach proportions with Gaussian regression models, proportion are not Gaussian by nature, since they are continuous variables bounded between 0 and 1.
 Thus, regression models with proportions as outcome variable should be fitted using a likelihood family that assumes such unit interval data.
@@
 Familiarity with regression modelling in R with a package like lme4 is assumed, but no prior knowledge of Bayesian statistics is necessary.
+The rest of the paper is structured as follows: @sec-beta introduces the mathematical underpinnings of the beta distribution, @sec-bayes provides the readers with a brief explanation and justification of Bayesian regression models, while @sec-case-1 and @sec-case-2 illustrate how to fit Bayesian beta regression models with two case studies.
 
-# The beta distribution
+# The beta distribution  {#sec-beta}
 
-The beta distribution is often considered as *the* distribution for the modeling of unit interval data [@ferrari_beta_2004; @cribari-neto_beta_2010].
+The beta distribution is often considered as *the* distribution for the modelling of unit interval data [@ferrari_beta_2004; @cribari-neto_beta_2010].
 It has been extensively studied theoretically [@krysicki_new_1999; @gupta_handbook_2014; @espinheira_beta_2008] and is used a baseline to compare other unit interval distributions against [@kieschnick_regression_2003; @bonat_regression_2013; @lopez_bayesian_2013].
@@
 
-We use a common mean parameterization for the beta distribution with mean parameter $\mu$, bounded between 0 and 1, and positive precision parameter $\phi$ that is roughly proportional to the inverse of the variance $\text{Var}(y) = \frac{\mu \, (1 - \mu)}{\phi + 1}$.
+We use a common mean parametrisation for the beta distribution with mean parameter $\mu$, bounded between 0 and 1, and positive precision parameter $\phi$ that is roughly proportional to the inverse of the variance $\text{Var}(y) = \frac{\mu \, (1 - \mu)}{\phi + 1}$.
 That is, the larger $\phi$ the smaller the variance of the corresponding beta distribution.
@@
 
-where $B$ is the Beta function, a complex integral for whose numerical approximation efficient algorithms
-exist in every programming language.
+where $B$ is the Beta function, a complex integral for whose numerical approximation efficient algorithms exist in every programming language.
 
-# Case study 1: voicing within consonant closure
+# Bayesian regression models  {#sec-bayes}
+
+Bayesian regression models are being increasingly adopted within phonetics and language research more broadly [@vasishth2018; @nalborczyk2019; @verissimo2021].
+Bayesian inference involves updating of prior probability distributions in light of evidence from data, to produce posterior probability distributions.
+In Bayesian regression models, model parameters are modelled as full probability distributions, rather than point estimates as in Null Hypothesis Significance Testing.
+Given the difficulty of analytical solutions of model equations, Bayesian regressions rely on sampling algorithms to reconstruct the posterior distributions.
+The statistical language Stan [@standevelopmentteam2017] employs efficient and robust Markov Chain Monte Carlo algorithms for fitting a variety of models, and the R package brms allows R users to interface with Stan from within R to fit Bayesian regression models [@burkner2017; @burkner2018; @burkner2021].
+
+The main practical advantage of Bayesian regression models over maximum-likelihood-based frequentist regression models, like those fitted with the lme4 package [@bates2015], is that Bayesian regression models don't suffer from the convergence issues that models fitted in lme4 [@bates2015] and other packages [@cribarineto2010] that fit frequentist models do, independent of sample size.
+A second, long-term advantage is that Bayesian regression models allow researchers to statistically re-use information from previous studies by specifying informative priors.
+While prior specification is one of the main features of Bayesian inference, in this tutorial we will use the default priors as set by brms.
+These are sensible priors estimated from the data that facilitate convergence but bear virtually no influence on the estimated posteriors.
+Specifying priors requires a great deal of precise quantitative knowledge, which in most areas of phonetics we still do not possess, so that using default uninformative priors is, for the time being, theoretically sound.
+
+The output of Bayesian regression models is (posterior) probability distributions for the model parameters, through which researchers can quantify (un)certainty.
+Bayesian Credible Intervals (CrIs) can be calculated from the posterior distributions at several probability levels (e.g.
+95, 90, 80, 60%) for a more complete view on estimated parameters.
+Bayesian CrIs indicate that at a certain probability levels the parameter's estimate lies within that interval: so, for example, a 90% CrI \[A, B\] indicates that there is a 90% probability that the estimate is between A and B.
+Different probability levels correspond to different levels of confidence: the higher the probability the higher the confidence (always conditional on data and model).
+Readers interested in a full and accessible exposition of Bayesian statistics are referred to @mcelreath2019.
+Shorter introductions can be found in @etz2018, @vasishth2018 and @nalborczyk2019.
+
+# Case study 1: voicing within consonant closure  {#sec-case-1}
 
 For the first case study, we will model the proportion of voicing within consonant closure.
-The measurements come from a data set of audio and electroglottographic (EGG) recordings of 19 speakers of Northwestern Italian.
+The measurements come from a data set of audio and electroglottographic (EGG) recordings of 19 speakers of Northwestern Italian [@coretta2019k; @coretta2020b].
 The participants read frame sentences which included target words of the form /CVCo/, where /C/ was either /k, t, p/ in all permutations and /V/ was either /i, e, a, ɔ, u/ (two resulting words, /peto/ and /kako/ were excluded because they are profanities), for a total of 43 target words.
-There were 4 different frame sentence, with a total of 172 trials per participant (3,268 grand total).
+There were 4 different frame sentence: *Scrivete X sul foglio* 'Write X on the sheet', *Ha detto X sei volte* 'She said X six times', *Sentivo X di nuovo* 'I heard X again', *Ripete X da sempre* 'She's been repeating X since ever'.
+There is a total of 172 trials per participant (3,268 grand total).
 The actual observation count is 2,419, after removing speech errors, EGG measurement errors, and cases in which voicing ceased before the closure onset/after the closure offset of the second /C/.
@@
 The proportion of voicing during the closure of the second /C/ was calculated as the proportion of contiguous voicing duration after closure onset to total duration of closure.
-The following code chunk attaches the tidyverse packages (for reading and wrangling data) and loads the `ita_egg` tibble (data frame).
+The following code chunk attaches the tidyverse packages [for reading and wrangling data, @wickham2019] and loads the `ita_egg` tibble (data frame).
 The tibble is filtered so as to remove voicing proportions (`voi_clo_prop`) that are smaller than 0 and greater than 1.
@@
 
-We will use brms to fit Bayesian beta regressions.
-\[XXX why Bayesian\].
+We will use brms to fit Bayesian beta regressions [@burkner2017].
 The model has voicing proportion as the outcome variable and the following terms: an interaction between vowel (/i, e, a, ɔ, u/) and second consonant C2 (/k, t, p/), centred speech rate (number of syllables per second); as varying (aka random) terms, by-speaker varying coefficients for the vowel/consonant interaction and for centred speech rate.[^1] The categorical predictors vowel and C2 are coded using indexing rather than traditional R contrasts: in R, this corresponds to suppressing the model's intercept with the `0 +` syntax; using indexing instead of contrasts makes it easier to specify priors.
@@
 There is nothing special about 95% CrI within Bayesian inference and instead experts recommend to check and report a variety of CrIs.
-Bayesian CrIs indicate that at a certain probability levels the "true" estimate lies within that interval: so, for example, a 90% CrI \[A, B\] indicates that there is a 90% probability that the "true" estimate is between A and B.
-Different probability levels correspond to different levels of confidence: the higher the probability the higher the confidence (always conditional on data and model).
 Obtaining CrIs at different probability levels allows researchers to make more fine-grained inferential statements than the frequentist significance dichotomy affords.
@@
 
-Finally, the package marginaleffects [XXX] has two other convenience functions that return CrIs of comparisons across predictor levels (`avg_comparisons()`) and CrIs of posterior predictions across predictor levels (`avg_predictions`).
+Finally, the package marginaleffects \[XXX\] has two other convenience functions that return CrIs of comparisons across predictor levels (`avg_comparisons()`) and CrIs of posterior predictions across predictor levels (`avg_predictions`).
 \[XXX\]
@@
 
-# Case study 2: coarticulatory vowel nasalisation
+# Case study 2: coarticulatory vowel nasalisation  {#sec-case-2}
 
@@
 Moving onto question 2: is there individual speaker variation?
-[TAMMINGA XXX] For this, we will use the `spread_draws()` function from tidybayes [XXX] to extract the draws of the varying terms (in brms these are the coefficients that start with `r_`).
+\[TAMMINGA XXX\] For this, we will use the `spread_draws()` function from tidybayes \[XXX\] to extract the draws of the varying terms (in brms these are the coefficients that start with `r_`).
 There is quite a few steps of processing to get from the raw draws to the estimand we need: while we have commented the following code, we encourage readers to test each line sequentialy and inspect the intermediate output to fully understand the process.

### 2025-02-06 manuscript/manuscript.tex
edit manuscript

@@
 Phonetic research often involves numeric continuous outcome variables,
-like durations, frequencies, loudness and ratios. Another commonly
-employed type of outcome variable are proportions: for example,
-proportion of voicing during closure, vocal folds contact quotient,
-gesture amplitude, nasalance. Moreover, virtually any measure can be
-MIN-MAX normalised, a procedure which transforms values so that they are
-in the range 0--1.
-
-Regression models are very common, but there is a tendency of using
-Gaussian distribution families (i.e.~probability distributions for the
-outcome variable) for anything that is numeric. A possible reason is
-that the base R function for fitting regression models, \texttt{lm()},
-and the lme4 function used to fit regression models with varying terms,
-\texttt{lmer()}, both fit Gaussian regressions by default and the user
-does not have to specify the distribution family. This tacit defaulting
-to Gaussian models is also reflected in teaching practices, where test
-and models using the Gaussian distribution are the first to be taught,
-due to their relative simplicity and the fact that other models are
-generalisations of Gaussian models.
-
-However, proportion are usually not Gaussian, since they are limited
-between 0 and 1. Thus, regression models with proportions as outcome
-variable should be fitted using a likelihood family that assumes such
-unit interval data. A common and flexible distribution for this purpose
-is the beta distribution. This tutorial introduces researchers to beta
-regression models in R using the package brms. Familiarity with
-regression modelling in R with a package like lme4 is assumed, but no
-prior knowledge of Bayesian statistics is necessary.
-
-\section{The beta distribution}\label{the-beta-distribution}
+like durations, frequencies, loudness and ratios
+\citep{kluender1988, johnson2003, gick2013, ladefoged2014, abramson2017, roettger2018, coretta2023}.
+Another commonly employed type of outcome variable are proportions: for
+example, proportion of voicing during closure \citep{davidson2016},
+vocal folds contact quotient \citep{herbst2017}, gesture amplitude
+\citep{carignan2021}, nasalance \citep{carignan2021}. Moreover,
+virtually any measure can be MIN-MAX normalised, a procedure which
+transforms values so that they are in the range 0--1.
+
+Regression models (and extensions like generalised additive models) have
+become a \emph{de facto} standard for the statistical analysis of a
+variety of measures in phonetic research
+\citep{kirby2018, politzerahles2018, tavakoli2024}. However, there is a
+tendency for researchers to use Gaussian distribution families
+(i.e.~probability distributions for the outcome variable) for any
+measure that is numeric, irrespective of whether the measure is
+unbounded, like in truly Gaussian variables, or bounded, like in
+proportions. A possible reason is that the base R function for fitting
+regression models, \texttt{lm()}, and the lme4 function used to fit
+regression models with varying terms, \texttt{lmer()} from lme4
+\citep{bates2015}, both fit Gaussian regressions by default and the user
+does not have to specify the distribution family. This tacit default of
+using Gaussian models is also reflected in teaching practices, where
+significance test and models using the Gaussian distribution are the
+first to be taught \citep{baayen2008, winter2020}, due to their relative
+simplicity and the fact that regression models with other families are
+conceptual generalisations of Gaussian regression models.
+
+While most researchers approach proportions with Gaussian regression
+models, proportion are not Gaussian by nature, since they are continuous
+variables bounded between 0 and 1. Thus, regression models with
+proportions as outcome variable should be fitted using a likelihood
+family that assumes such unit interval data. A common and flexible
+distribution for this purpose is the beta distribution. This tutorial
+introduces researchers to beta regression models in R using the package
+brms. Familiarity with regression modelling in R with a package like
+lme4 is assumed, but no prior knowledge of Bayesian statistics is
+necessary. The rest of the paper is structured as follows:
+Section~\ref{sec-beta} introduces the mathematical underpinnings of the
+beta distribution, Section~\ref{sec-bayes} provides the readers with a
+brief explanation and justification of Bayesian regression models, while
+Section~\ref{sec-case-1} and Section~\ref{sec-case-2} illustrate how to
+fit Bayesian beta regression models with two case studies.
+
+\section{The beta distribution}\label{sec-beta}
 
 The beta distribution is often considered as \emph{the} distribution for
-the modeling of unit interval data
+the modelling of unit interval data
 \citep{ferrari_beta_2004, cribari-neto_beta_2010}. It has been
@@
 
-We use a common mean parameterization for the beta distribution with
-mean parameter \(\mu\), bounded between 0 and 1, and positive precision
+We use a common mean parametrisation for the beta distribution with mean
+parameter \(\mu\), bounded between 0 and 1, and positive precision
 parameter \(\phi\) that is roughly proportional to the inverse of the
@@
 
+\section{Bayesian regression models}\label{sec-bayes}
+
+Bayesian regression models are being increasingly adopted within
+phonetics and language research more broadly
+\citep{vasishth2018, nalborczyk2019, verissimo2021}. Bayesian inference
+involves updating of prior probability distributions in light of
+evidence from data, to produce posterior probability distributions. In
+Bayesian regression models, model parameters are modelled as full
+probability distributions, rather than point estimates as in Null
+Hypothesis Significance Testing. Given the difficulty of analytical
+solutions of model equations, Bayesian regressions rely on sampling
+algorithms to reconstruct the posterior distributions. The statistical
+language Stan \citep{standevelopmentteam2017} employs efficient and
+robust Markov Chain Monte Carlo algorithms for fitting a variety of
+models, and the R package brms allows R users to interface with Stan
+from within R to fit Bayesian regression models
+\citep{burkner2017, burkner2018, burkner2021}.
+
+The main practical advantage of Bayesian regression models over
+maximum-likelihood-based frequentist regression models, like those
+fitted with the lme4 package \citep{bates2015}, is that Bayesian
+regression models don't suffer from the convergence issues that models
+fitted in lme4 \citep{bates2015} and other packages
+\citep{cribarineto2010} that fit frequentist models do, independent of
+sample size. A second, long-term advantage is that Bayesian regression
+models allow researchers to statistically re-use information from
+previous studies by specifying informative priors. While prior
+specification is one of the main features of Bayesian inference, in this
+tutorial we will use the default priors as set by brms. These are
+sensible priors estimated from the data that facilitate convergence but
+bear virtually no influence on the estimated posteriors. Specifying
+priors requires a great deal of precise quantitative knowledge, which in
+most areas of phonetics we still do not possess, so that using default
+uninformative priors is, for the time being, theoretically sound.
+
+The output of Bayesian regression models is (posterior) probability
+distributions for the model parameters, through which researchers can
+quantify (un)certainty. Bayesian Credible Intervals (CrIs) can be
+calculated from the posterior distributions at several probability
+levels (e.g.~95, 90, 80, 60\%) for a more complete view on estimated
+parameters. Bayesian CrIs indicate that at a certain probability levels
+the parameter's estimate lies within that interval: so, for example, a
+90\% CrI {[}A, B{]} indicates that there is a 90\% probability that the
+estimate is between A and B. Different probability levels correspond to
+different levels of confidence: the higher the probability the higher
+the confidence (always conditional on data and model). Readers
+interested in a full and accessible exposition of Bayesian statistics
+are referred to \citet{mcelreath2019}. Shorter introductions can be
+found in \citet{etz2018}, \citet{vasishth2018} and
+\citet{nalborczyk2019}.
+
 \section{Case study 1: voicing within consonant
-closure}\label{case-study-1-voicing-within-consonant-closure}
+closure}\label{sec-case-1}
 
@@
 electroglottographic (EGG) recordings of 19 speakers of Northwestern
-Italian. The participants read frame sentences which included target
-words of the form /CVCo/, where /C/ was either /k, t, p/ in all
-permutations and /V/ was either /i, e, a, ɔ, u/ (two resulting words,
-/peto/ and /kako/ were excluded because they are profanities), for a
-total of 43 target words. There were 4 different frame sentence, with a
-total of 172 trials per participant (3,268 grand total). The actual
-observation count is 2,419, after removing speech errors, EGG
-measurement errors, and cases in which voicing ceased before the closure
-onset/after the closure offset of the second /C/.
+Italian \citep{coretta2019k, coretta2020b}. The participants read frame
+sentences which included target words of the form /CVCo/, where /C/ was
+either /k, t, p/ in all permutations and /V/ was either /i, e, a, ɔ, u/
+(two resulting words, /peto/ and /kako/ were excluded because they are
+profanities), for a total of 43 target words. There were 4 different
+frame sentence: \emph{Scrivete X sul foglio} `Write X on the sheet',
+\emph{Ha detto X sei volte} `She said X six times', \emph{Sentivo X di
+nuovo} `I heard X again', \emph{Ripete X da sempre} `She's been
+repeating X since ever'. There is a total of 172 trials per participant
+(3,268 grand total). The actual observation count is 2,419, after
+removing speech errors, EGG measurement errors, and cases in which
+voicing ceased before the closure onset/after the closure offset of the
+second /C/.
 
@@
 closure onset to total duration of closure. The following code chunk
-attaches the tidyverse packages (for reading and wrangling data) and
-loads the \texttt{ita\_egg} tibble (data frame). The tibble is filtered
-so as to remove voicing proportions (\texttt{voi\_clo\_prop}) that are
-smaller than 0 and greater than 1. The variables \texttt{vowel} and
-\texttt{c2} are converted to factors to specify the order of the levels.
+attaches the tidyverse packages \citep[for reading and wrangling
+data,][]{wickham2019} and loads the \texttt{ita\_egg} tibble (data
+frame). The tibble is filtered so as to remove voicing proportions
+(\texttt{voi\_clo\_prop}) that are smaller than 0 and greater than 1.
+The variables \texttt{vowel} and \texttt{c2} are converted to factors to
+specify the order of the levels.
 
@@
 
-We will use brms to fit Bayesian beta regressions. {[}XXX why
-Bayesian{]}. The model has voicing proportion as the outcome variable
-and the following terms: an interaction between vowel (/i, e, a, ɔ, u/)
-and second consonant C2 (/k, t, p/), centred speech rate (number of
+We will use brms to fit Bayesian beta regressions \citep{burkner2017}.
+The model has voicing proportion as the outcome variable and the
+following terms: an interaction between vowel (/i, e, a, ɔ, u/) and
+second consonant C2 (/k, t, p/), centred speech rate (number of
 syllables per second); as varying (aka random) terms, by-speaker varying
@@
 inference and instead experts recommend to check and report a variety of
-CrIs. Bayesian CrIs indicate that at a certain probability levels the
-``true'' estimate lies within that interval: so, for example, a 90\% CrI
-{[}A, B{]} indicates that there is a 90\% probability that the ``true''
-estimate is between A and B. Different probability levels correspond to
-different levels of confidence: the higher the probability the higher
-the confidence (always conditional on data and model). Obtaining CrIs at
-different probability levels allows researchers to make more
-fine-grained inferential statements than the frequentist significance
-dichotomy affords. For simplicity of exposition, we will use 80\% CrIs
-in this case study but we strongly recommend researchers to always
-obtain CrIs at different levels of probability and base their inferences
-on all and not one in particular. To reiterate, in Bayesian inference,
-an 80\% CrI indicates the range of values within which the true estimate
-falls at 80\% probability or confidence.
+CrIs. Obtaining CrIs at different probability levels allows researchers
+to make more fine-grained inferential statements than the frequentist
+significance dichotomy affords. For simplicity of exposition, we will
+use 80\% CrIs in this case study but we strongly recommend researchers
+to always obtain CrIs at different levels of probability and base their
+inferences on all and not one in particular. To reiterate, in Bayesian
+inference, an 80\% CrI indicates the range of values within which the
+true estimate falls at 80\% probability or confidence.
 
@@
 \section{Case study 2: coarticulatory vowel
-nasalisation}\label{case-study-2-coarticulatory-vowel-nasalisation}
+nasalisation}\label{sec-case-2}
 

### 2025-02-06 manuscript/manuscript.qmd
edit manuscript

@@
 Specifying priors requires a great deal of precise quantitative knowledge, which in most areas of phonetics we still do not possess, so that using default uninformative priors is, for the time being, theoretically sound.
+For practical guidance on prior specification, see <https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations>.
 
@@
 
-# Case study 1: voicing within consonant closure  {#sec-case-1}
+# Case study 1: coarticulatory voicing within consonant closure {#sec-case-1}
 
-For the first case study, we will model the proportion of voicing within consonant closure.
+For the first case study, we will model the proportion of coarticulatory progressive voicing within the consonant closure of voiceless stops.
 The measurements come from a data set of audio and electroglottographic (EGG) recordings of 19 speakers of Northwestern Italian [@coretta2019k; @coretta2020b].
-The participants read frame sentences which included target words of the form /CVCo/, where /C/ was either /k, t, p/ in all permutations and /V/ was either /i, e, a, ɔ, u/ (two resulting words, /peto/ and /kako/ were excluded because they are profanities), for a total of 43 target words.
+The participants read frame sentences which included target words of the form /CVCo/, where C was either /k, t, p/ in all permutations and V was either /i, e, a, o, u/ (two resulting words, /peto/ and /kako/ were excluded because they are profanities), for a total of 43 target words.
 There were 4 different frame sentence: *Scrivete X sul foglio* 'Write X on the sheet', *Ha detto X sei volte* 'She said X six times', *Sentivo X di nuovo* 'I heard X again', *Ripete X da sempre* 'She's been repeating X since ever'.
 There is a total of 172 trials per participant (3,268 grand total).
-The actual observation count is 2,419, after removing speech errors, EGG measurement errors, and cases in which voicing ceased before the closure onset/after the closure offset of the second /C/.
+The actual observation count is 2,419, after removing speech errors, EGG measurement errors, and cases in which voicing ceased before the closure onset/after the closure offset of the second consonant in the target word.
 
-The proportion of voicing during the closure of the second /C/ was calculated as the proportion of contiguous voicing duration after closure onset to total duration of closure.
+The proportion of voicing during the closure of the second consonant was calculated as the proportion of contiguous voicing duration after closure onset to total duration of closure.
 The following code chunk attaches the tidyverse packages [for reading and wrangling data, @wickham2019] and loads the `ita_egg` tibble (data frame).
 The tibble is filtered so as to remove voicing proportions (`voi_clo_prop`) that are smaller than 0 and greater than 1.
-The variables `vowel` and `c2` are converted to factors to specify the order of the levels.
+The variables `vowel` (first vowel in the target word) and `c2` (second consonant in the target word) are converted to factors to specify the order of the levels.
+@tbl-ita-egg shows the first ten rows of the tibble (only relevant columns are included).
 
@@
 
-@tbl-ita-egg shows the first ten rows of the tibble (only relevant columns are included).
-
 ```{r}
@@
 
-@fig-ita-egg shows the raw voicing duration proportion values split by vowel /i, e, a, ɔ, u/ and second consonant /k, t, p/ in the /CVCo/ target words.
+@fig-ita-egg shows the raw voicing duration proportion values split by vowel /i, e, a, o, u/ and second consonant /k, t, p/ in the /CVCo/ target words.
 The plot suggests that, on average, the voicing proportion is slightly lower with /k/ than with /t, p/.
 Moreover, there is greater variability between vowels in /t, p/ than in /k/.
-We will use a beta regression model to assess these patterns.
-\["expectations" XXX\]
+We will use a beta regression model to assess whether the place of articulation of the consonant affects the proportion of coarticulatory progressive voicing, while averaging across vowels (of course, another question that could be answered with the same model is if different vowels lead to difference voicing proportions, but for sake of space we restrict this tutorial to the former question).
+A general expectation is that the proportion of progressive voicing should increase with fronter places of articulation (so from /k/ to /t/ to /p/) given the aerodynamic constraints which obtain with vocal fold vibration during a fully occluded oral tract [@vandenberg1958; @rothenberg1967; @ohala2011].
 
@@
 
-We will use brms to fit Bayesian beta regressions [@burkner2017].
-The model has voicing proportion as the outcome variable and the following terms: an interaction between vowel (/i, e, a, ɔ, u/) and second consonant C2 (/k, t, p/), centred speech rate (number of syllables per second); as varying (aka random) terms, by-speaker varying coefficients for the vowel/consonant interaction and for centred speech rate.[^1] The categorical predictors vowel and C2 are coded using indexing rather than traditional R contrasts: in R, this corresponds to suppressing the model's intercept with the `0 +` syntax; using indexing instead of contrasts makes it easier to specify priors.
-For pedagogical simplicity, the model will use the default priors, but note that in real data analyses contexts, priors should be specified by the user.
-I refer the readers to XXX.
+We will use brms to fit a Bayesian beta regression [@burkner2017].
+The model has voicing proportion as the outcome variable and the following terms: an interaction between vowel (/i, e, a, o, u/) and second consonant C2 (/k, t, p/), centred speech rate (number of syllables per second); as varying (aka random) terms, we included by-speaker varying coefficients for the vowel/consonant interaction and for centred speech rate.[^1] The categorical predictors vowel and C2 are coded using indexing rather than the traditional R contrasts: in R, this corresponds to suppressing the model's intercept with the `0 +` syntax; using indexing instead of contrasts makes it easier to specify priors and uncertainty is distributed equally across all levels of categorical predictors.
+For pedagogical simplicity, the model will use the default priors as explained in @sec-bayes.
 
-[^1]: Footnote about Gelman's terminology for random effects.
+[^1]: While the terms "random effects, intercept and slopes" are commonly employed in applied statistical work in linguistics and phonetics, we opt to use the conceptually more appropriate terminology proposed by @gelman2005: varying terms, intercepts, slopes, coefficients.
+Other common terms are multilevel and hierarchical terms.
 
@@
 For conciseness, we will use the `fixef()` function which prints the regression coefficients.
-\[EXPLAIN EXPECTED PREDICTIONS XXX\].
 The full summary with an explanation of each part can be found in XXX.
@@
 For each coefficient in the model, `fixef()` prints the name of the coefficient, the mean estimate, the estimate error and the lower and upper limits of a Bayesian Credible interval (CrI).
+Note that all values are in log-odds, since the link function in beta regressions is the logistic function, like in binomial/Bernoulli regressions.
 Here, we print an 80% CrI.
@@
 
-The coefficients of a beta regression are estimated on the log-odds scale, as in Bernoulli/binomial (aka logistic) regressions.
-From the summary, we see that speech rate (number of syllables per second) has a positive effect on voicing proportion: the 80% CrI is between 0.01 and 0.15 log-odds \[$\beta$ = 0.08, SD = 0.06\].
+From the summary, we see that speech rate (number of syllables per second) has a positive effect on voicing proportion: the 80% CrI is between 0.01 and 0.15 log-odds [$\beta$ = 0.08, SD = 0.06].
 Log-odds can be converted to odd-ratios by exponentiating the value: 0.01-0.15 log odds correspond to an odd-ratio of 1.01 to 1.16, or as percentages, to an increase of voicing of 1 to 16% for every increase of one syllable per second.
 Since this is an 80% CrI, we can be 80% confident that the true effect of speech rate is between 1-16% increase of voicing proportion, conditional on the data and model.
-Note that transforming measures this way is appropriate *only* with quantile-based measures (like CrIs) but not with moments like the mean and standard deviation: to correctly get mean and SDs in the transformed scale, you must first extract the posterior draws (see below), convert them and then take moments such as mean and SD (for a more detailed explanation, see XXX).
+Note that transforming measures this way is appropriate *only* with quantile-based measures (like CrIs) but not with moments like the mean and standard deviation: to correctly get mean and SDs in the transformed scale, you must first extract the posterior draws (see below), convert them and then take moments such as mean and SD [for a more detailed explanation, see @burkner2025, p.
+101].
 In the avoidance of doubt, we will always transform the drawn values first and then take summary measures.

### 2025-02-06 manuscript/manuscript.tex
edit manuscript

@@
 most areas of phonetics we still do not possess, so that using default
-uninformative priors is, for the time being, theoretically sound.
+uninformative priors is, for the time being, theoretically sound. For
+practical guidance on prior specification, see
+\url{https://github.com/stan-dev/stan/wiki/Prior-Choice-Recommendations}.
 
@@
 
-\section{Case study 1: voicing within consonant
+\section{Case study 1: coarticulatory voicing within consonant
 closure}\label{sec-case-1}
 
-For the first case study, we will model the proportion of voicing within
-consonant closure. The measurements come from a data set of audio and
-electroglottographic (EGG) recordings of 19 speakers of Northwestern
-Italian \citep{coretta2019k, coretta2020b}. The participants read frame
-sentences which included target words of the form /CVCo/, where /C/ was
-either /k, t, p/ in all permutations and /V/ was either /i, e, a, ɔ, u/
+For the first case study, we will model the proportion of coarticulatory
+progressive voicing within the consonant closure of voiceless stops. The
+measurements come from a data set of audio and electroglottographic
+(EGG) recordings of 19 speakers of Northwestern Italian
+\citep{coretta2019k, coretta2020b}. The participants read frame
+sentences which included target words of the form /CVCo/, where C was
+either /k, t, p/ in all permutations and V was either /i, e, a, o, u/
 (two resulting words, /peto/ and /kako/ were excluded because they are
@@
 voicing ceased before the closure onset/after the closure offset of the
-second /C/.
-
-The proportion of voicing during the closure of the second /C/ was
+second consonant in the target word.
+
+The proportion of voicing during the closure of the second consonant was
 calculated as the proportion of contiguous voicing duration after
@@
 (\texttt{voi\_clo\_prop}) that are smaller than 0 and greater than 1.
-The variables \texttt{vowel} and \texttt{c2} are converted to factors to
-specify the order of the levels.
+The variables \texttt{vowel} (first vowel in the target word) and
+\texttt{c2} (second consonant in the target word) are converted to
+factors to specify the order of the levels. Table~\ref{tbl-ita-egg}
+shows the first ten rows of the tibble (only relevant columns are
+included).
 
@@
 \end{Shaded}
-
-Table~\ref{tbl-ita-egg} shows the first ten rows of the tibble (only
-relevant columns are included).
 
@@
 Figure~\ref{fig-ita-egg} shows the raw voicing duration proportion
-values split by vowel /i, e, a, ɔ, u/ and second consonant /k, t, p/ in
+values split by vowel /i, e, a, o, u/ and second consonant /k, t, p/ in
 the /CVCo/ target words. The plot suggests that, on average, the voicing
@@
 is greater variability between vowels in /t, p/ than in /k/. We will use
-a beta regression model to assess these patterns. {[}``expectations''
-XXX{]}
+a beta regression model to assess whether the place of articulation of
+the consonant affects the proportion of coarticulatory progressive
+voicing, while averaging across vowels (of course, another question that
+could be answered with the same model is if different vowels lead to
+difference voicing proportions, but for sake of space we restrict this
+tutorial to the former question). A general expectation is that the
+proportion of progressive voicing should increase with fronter places of
+articulation (so from /k/ to /t/ to /p/) given the aerodynamic
+constraints which obtain with vocal fold vibration during a fully
+occluded oral tract \citep{vandenberg1958, rothenberg1967, ohala2011}.
 
@@
 
-We will use brms to fit Bayesian beta regressions \citep{burkner2017}.
+We will use brms to fit a Bayesian beta regression \citep{burkner2017}.
 The model has voicing proportion as the outcome variable and the
-following terms: an interaction between vowel (/i, e, a, ɔ, u/) and
+following terms: an interaction between vowel (/i, e, a, o, u/) and
 second consonant C2 (/k, t, p/), centred speech rate (number of
-syllables per second); as varying (aka random) terms, by-speaker varying
-coefficients for the vowel/consonant interaction and for centred speech
-rate.\footnote{Footnote about Gelman's terminology for random effects.}
-The categorical predictors vowel and C2 are coded using indexing rather
-than traditional R contrasts: in R, this corresponds to suppressing the
-model's intercept with the \texttt{0\ +} syntax; using indexing instead
-of contrasts makes it easier to specify priors. For pedagogical
-simplicity, the model will use the default priors, but note that in real
-data analyses contexts, priors should be specified by the user. I refer
-the readers to XXX.
+syllables per second); as varying (aka random) terms, we included
+by-speaker varying coefficients for the vowel/consonant interaction and
+for centred speech rate.\footnote{While the terms ``random effects,
+  intercept and slopes'' are commonly employed in applied statistical
+  work in linguistics and phonetics, we opt to use the conceptually more
+  appropriate terminology proposed by \citet{gelman2005}: varying terms,
+  intercepts, slopes, coefficients. Other common terms are multilevel
+  and hierarchical terms.} The categorical predictors vowel and C2 are
+coded using indexing rather than the traditional R contrasts: in R, this
+corresponds to suppressing the model's intercept with the \texttt{0\ +}
+syntax; using indexing instead of contrasts makes it easier to specify
+priors and uncertainty is distributed equally across all levels of
+categorical predictors. For pedagogical simplicity, the model will use
+the default priors as explained in Section~\ref{sec-bayes}.
 
@@
 conciseness, we will use the \texttt{fixef()} function which prints the
-regression coefficients. {[}EXPLAIN EXPECTED PREDICTIONS XXX{]}. The
-full summary with an explanation of each part can be found in XXX.
-Table~\ref{tbl-voi-prop-fixef} reports the output of \texttt{fixef()} as
-a table (we round all values to the nearest 2 digits for clarity.). For
-each coefficient in the model, \texttt{fixef()} prints the name of the
-coefficient, the mean estimate, the estimate error and the lower and
-upper limits of a Bayesian Credible interval (CrI). Here, we print an
-80\% CrI. There is nothing special about 95\% CrI within Bayesian
-inference and instead experts recommend to check and report a variety of
-CrIs. Obtaining CrIs at different probability levels allows researchers
-to make more fine-grained inferential statements than the frequentist
-significance dichotomy affords. For simplicity of exposition, we will
-use 80\% CrIs in this case study but we strongly recommend researchers
-to always obtain CrIs at different levels of probability and base their
-inferences on all and not one in particular. To reiterate, in Bayesian
-inference, an 80\% CrI indicates the range of values within which the
-true estimate falls at 80\% probability or confidence.
+regression coefficients. The full summary with an explanation of each
+part can be found in XXX. Table~\ref{tbl-voi-prop-fixef} reports the
+output of \texttt{fixef()} as a table (we round all values to the
+nearest 2 digits for clarity.). For each coefficient in the model,
+\texttt{fixef()} prints the name of the coefficient, the mean estimate,
+the estimate error and the lower and upper limits of a Bayesian Credible
+interval (CrI). Note that all values are in log-odds, since the link
+function in beta regressions is the logistic function, like in
+binomial/Bernoulli regressions. Here, we print an 80\% CrI. There is
+nothing special about 95\% CrI within Bayesian inference and instead
+experts recommend to check and report a variety of CrIs. Obtaining CrIs
+at different probability levels allows researchers to make more
+fine-grained inferential statements than the frequentist significance
+dichotomy affords. For simplicity of exposition, we will use 80\% CrIs
+in this case study but we strongly recommend researchers to always
+obtain CrIs at different levels of probability and base their inferences
+on all and not one in particular. To reiterate, in Bayesian inference,
+an 80\% CrI indicates the range of values within which the true estimate
+falls at 80\% probability or confidence.
 
@@
 
-The coefficients of a beta regression are estimated on the log-odds
-scale, as in Bernoulli/binomial (aka logistic) regressions. From the
-summary, we see that speech rate (number of syllables per second) has a
-positive effect on voicing proportion: the 80\% CrI is between 0.01 and
-0.15 log-odds {[}\(\beta\) = 0.08, SD = 0.06{]}. Log-odds can be
-converted to odd-ratios by exponentiating the value: 0.01-0.15 log odds
-correspond to an odd-ratio of 1.01 to 1.16, or as percentages, to an
-increase of voicing of 1 to 16\% for every increase of one syllable per
-second. Since this is an 80\% CrI, we can be 80\% confident that the
-true effect of speech rate is between 1-16\% increase of voicing
-proportion, conditional on the data and model. Note that transforming
-measures this way is appropriate \emph{only} with quantile-based
-measures (like CrIs) but not with moments like the mean and standard
-deviation: to correctly get mean and SDs in the transformed scale, you
-must first extract the posterior draws (see below), convert them and
-then take moments such as mean and SD (for a more detailed explanation,
-see XXX). In the avoidance of doubt, we will always transform the drawn
-values first and then take summary measures.
+From the summary, we see that speech rate (number of syllables per
+second) has a positive effect on voicing proportion: the 80\% CrI is
+between 0.01 and 0.15 log-odds {[}\(\beta\) = 0.08, SD = 0.06{]}.
+Log-odds can be converted to odd-ratios by exponentiating the value:
+0.01-0.15 log odds correspond to an odd-ratio of 1.01 to 1.16, or as
+percentages, to an increase of voicing of 1 to 16\% for every increase
+of one syllable per second. Since this is an 80\% CrI, we can be 80\%
+confident that the true effect of speech rate is between 1-16\% increase
+of voicing proportion, conditional on the data and model. Note that
+transforming measures this way is appropriate \emph{only} with
+quantile-based measures (like CrIs) but not with moments like the mean
+and standard deviation: to correctly get mean and SDs in the transformed
+scale, you must first extract the posterior draws (see below), convert
+them and then take moments such as mean and SD \citep[for a more
+detailed explanation, see][p.~101]{burkner2025}. In the avoidance of
+doubt, we will always transform the drawn values first and then take
+summary measures.
 

### 2025-02-07 manuscript/manuscript.qmd
edit manuscript

@@
 
-Phonetic research often involves numeric continuous outcome variables, like durations, frequencies, loudness and ratios [@kluender1988; @johnson2003; @gick2013; @ladefoged2014; @abramson2017; @roettger2018; @coretta2023].
+Phonetic research often involves numeric continuous outcome variables, like durations, frequencies, loudness and ratios [@kluender1988; @johnson2003; @gick2013; @ladefoged2014; @abramson2017; @roettger2018; @coretta2023; @rosen2005].
 Another commonly employed type of outcome variable are proportions: for example, proportion of voicing during closure [@davidson2016], vocal folds contact quotient [@herbst2017], gesture amplitude [@carignan2021], nasalance [@carignan2021].
@@
 We will use brms to fit a Bayesian beta regression [@burkner2017].
-The model has voicing proportion as the outcome variable and the following terms: an interaction between vowel (/i, e, a, o, u/) and second consonant C2 (/k, t, p/), centred speech rate (number of syllables per second); as varying (aka random) terms, we included by-speaker varying coefficients for the vowel/consonant interaction and for centred speech rate.[^1] The categorical predictors vowel and C2 are coded using indexing rather than the traditional R contrasts: in R, this corresponds to suppressing the model's intercept with the `0 +` syntax; using indexing instead of contrasts makes it easier to specify priors and uncertainty is distributed equally across all levels of categorical predictors.
+The model has voicing proportion as the outcome variable and the following terms: an interaction between vowel (/i, e, a, o, u/) and second consonant C2 (/k, t, p/), centred speech rate (number of syllables per second); as varying (aka random) terms, we included by-speaker varying coefficients for the vowel/consonant interaction and for centred speech rate.[^1] The categorical predictors vowel and C2 are coded using indexing rather than the traditional R contrasts, as advocated in @mcelreath2019: in R, this corresponds to suppressing the model's intercept with the `0 +` syntax; using indexing instead of contrasts makes it easier to specify priors and uncertainty is distributed equally across all levels of categorical predictors.
 For pedagogical simplicity, the model will use the default priors as explained in @sec-bayes.
@@
 This matches the general trends in the raw data, which we plotted in @fig-ita-egg.
-
 Next, we will plot the predicted proportions of each vowel/C2 combination at mean speech rate (i.e.
 centred speech rate = 0) and then calculate the average pair-wise difference in voicing proportion between /k, t, p/.
-Finally, we will assess whether there is greater between-vowel variability in /t, p/ relative to /k/.
 
 Before being able to plot the predictions, it's important to get familiar with the so-called posterior draws.
-\[Bayesian MCMC XXX\].
-Posterior draws can be conveniently obtained with the `as_draws_df()`.
+Bayesian regression models are fitted using Markov Chain Monte Carlo (MCMC) methods, as briefly explained in @sec-bayes.
+The `brm()` function runs three MCMC chains with 2000 iterations each by default.
+These are sufficient in most cases.
+The first 1000 iterations of each chain are used for "warmup", where the MCMC algorithm is optimised to find the posterior probability.
+The second half of the iterations, 1000 per chain, are used for sampling from the posterior probability.
+Each iteration returns a drawn value for each parameter in the model.
+The list of values drawn during the Monte Carlo sampling are called the "posterior draws".
+The posterior draws are used for inference: we can plot them, summarise them, transform them and combine them in any way to answer specific questions.
+
+Posterior draws can be conveniently obtained with the `as_draws_df()` from brms.
 For the moment, we will extract only the draws of the constant regression coefficients (model variables starting with `b_`).
-To check which coefficients are available in a model, use `get_variables()` from the tidybayes package.
+To check which coefficients are available in a model, use `get_variables()` from the tidybayes package [@kay2019].
 `as_draws_df()` returns a tibble where each column contains the drawn values of a coefficient.
@@
 
-For plotting, we can use ggplot2 statistics layers from the ggdist package.
+For plotting, we can use ggplot2 statistics layers from the ggdist package [@kay2023; kay2024].
 `stat_halfeye()` plots the density of the posterior probability (in grey), its median (point) and CrIs (lines).
@@
   stat_halfeye(.width = c(0.6, 0.8)) +
-  facet_grid(rows = vars(c2))
+  facet_grid(rows = vars(c2)) +
+  labs(
+    caption = "The points are the medians and the two lines are 60 and 80% CrIs."
+  )
 ```
@@
   ggplot(aes(plogis(value_mean), c2)) +
-  stat_halfeye(.width = c(0.6, 0.8))
+  stat_halfeye(.width = c(0.6, 0.8)) +
+  labs(
+    caption = "The points are the medians and the two lines are 60 and 80% CrIs."
+  )
 ```
@@
 
-Once we have the posterior difference, we can obtain CrIs of the difference using `quantile2()` from the posterior package.
+Once we have the posterior difference, we can obtain CrIs of the difference using `quantile2()` from the posterior package [@burkner2024].
 Beware that the values of the difference are in log-odds!
@@
 Odd-ratios indicate the ratio of the difference between A and B, so that 1 means no difference, values greater than 1 indicate an increase in A relative to B and values lower than 1 indicate a decrease in A relative to B.
-odd-ratios are useful when looking at differences that are in log-odds because while the relative magnitude of the difference in proportion between two groups is the same independent of the baseline proportion, the *absolute* magnitude of the difference depends on the baseline value.
-For example, an odd-ratio difference of 1.25 would correspond to a proportion increase of 13 percentage points if the baseline proportion is 0.62 but it would correspond to a proportion increase of 17 percentage points if the baseline proportion is 0.73.
-Of course, in real research contexts it is still usefull to think about absolute magnitudes and their relevance from a conceptual and methodological perspective.
+Odd-ratios are useful when looking at differences that are in log-odds because while the relative magnitude of the difference in proportion between two groups is the same independent of the baseline proportion, the *absolute* magnitude of the difference depends on the baseline value.
+For example, an odd-ratio difference of 1.25 would correspond to a proportion increase of about 3 percentage points if the baseline proportion is 0.62 but it would correspond to a proportion increase of 5 percentage points if the baseline proportion is 0.73 and 1 percentage point if the baseline is 0.99.
+Of course, in real research contexts it is still useful to think about absolute magnitudes and their relevance from a conceptual and methodological perspective.
 In this tutorial we just focus on odd-ratios for simplicity.
+
+```{r}
+#| label: odds
+#| include: false
+#| eval: false
+
+logs <- qlogis(0.62)
+p <- plogis(logs * 1.25)
+round(p - 0.62, 2)
+
+logs <- qlogis(0.73)
+p <- plogis(logs * 1.25)
+round(p - 0.73, 2)
+
+logs <- qlogis(0.99)
+p <- plogis(logs * 1.25)
+round(p - 0.99, 2)
+```
 
@@
 
-Finally, the package marginaleffects \[XXX\] has two other convenience functions that return CrIs of comparisons across predictor levels (`avg_comparisons()`) and CrIs of posterior predictions across predictor levels (`avg_predictions`).
-\[XXX\]
+Finally, the package marginaleffects [@arelbundock2025] has two other convenience functions that return CrIs of comparisons across predictor levels (`avg_comparisons()`) and CrIs of posterior predictions across predictor levels (`avg_predictions`).
+Both functions take the model and a list of variables as arguments.
+Optionally, you can set the confidence level with the `conf_level` argument (here, we set it to 0.8, or 80%).
+`avg_comparisons()` allows the user to specify the type of comparison for each variable: here we set `"pairwise"` to calculate pairwise comparisons of each level of `c2`.
+The output of the two functions is shows in @tbl-marginal-1 and @tbl-marginal-2.
 
@@
 Here, we will focus on the effect of C voicing (voiceless /t/ vs voiced /d/) on the proportion of nasalisation within the vowel in the VNC sequence.
-Previous work on coarticulatory nasalisation in English has suggested that vowels followed by an NC sequence where C is voiceless (NT) should show earlier coarticulatory nasalisation than vowels followed by an NC sequence where C is voiced [ND, see review in @carignan2021].
+Previous work on coarticulatory nasalisation in English has suggested that vowels followed by an NC sequence where C is voiceless (NT) should show earlier coarticulatory nasalisation than vowels followed by an NC sequence where C is voiced (ND).
 This pattern has been suggested to be driven by the articulatory and acoustic incompatibility of voicelessness and nasalisation, by which the velum opening gesture of the nasal consonant is pushed away (i.e.
-earlier) when the consonant following the nasal is voiceless.
+earlier) when the consonant following the nasal is voiceless [@fowler2005; @beddor2009; @cho2017; @carignan2015].
 Everything else being equal, a greater proportion of vowel nasalisation (from the perspective of time) should be found in vowels followed by NT than in vowels followed by ND.
@@
 
--   `NC` is the NC sequence.
+-   `NC` is the NC sequence (voiceless /nt/ or voiced /nd/).
 
@@
 Negative log-odds indicate a proportion that is smaller than 50%, while positive log-odds a proportion that is greater than 50%.
-Generally, the expected log-odd predictions are negative, indicating an overall tendency for the nasalisation to take less than 50% of the duration of the vowel.
+Generally, the expected log-odd predictions in @tbl-nas-prop-fixef are negative, indicating an overall tendency for the nasalisation to take less than 50% of the duration of the vowel.
 Moreover, the predictions are higher for voiceless NC sequences than for voiced NC sequences, indicating a greater proportion of nasalisation in the former.
 However there is vowel-specific variation, and there doesn't seem to be much of a difference in nasalisation proportion in /e/ and /u/.
-@fig-nas-prop shows the expected predictions with `conditional_effects()`, which should make the interpretation clearer.
+@fig-nas-prop shows the expected predictions with `conditional_effects()`, which should make the interpretation of the predictions from @tbl-nas-prop-fixef clearer.
 
@@
 
-Now that we fitted the model we can use the draws to answer the two research questions (repeated from above):
+Now that we fitted the model we can use the posterior draws to answer the two research questions (repeated from above):
 
@@
 
-To answer question 1, we can calculate the average difference in nasalisation proportion by first calculating the average nasalisation across all vowels for voiced and voiceless sequences (see @tbl-nas-draws for the output of this step) and then take the difference of those, similarly to what we have done in the Case Study 1 above.
+To answer question 1, we can calculate the average difference in nasalisation proportion by first calculating the average nasalisation across all vowels for voiced and voiceless sequences (see @tbl-nas-draws for the output of this step) and then take the difference of those, similarly to what we have done in the case study in @sec-case-1.
 
@@
 Now let's calculate the mean nasalisation proportion within each draw by voicing, and plot the resulting posterior distributions.
-Note that, as discussed for Case study 1, when working wit log-odds it is important to first do all necessary calculations in log-odds, here calculate the mean log-odds across vowels, and *then* transform the calculated estimands to proportions/probabilities.
+Note that, as discussed in @sec-case-1, when working with log-odds it is important to first do all necessary calculations in log-odds (here calculate the mean log-odds across vowels) and *then* transform the calculated estimands to proportions/probabilities.
 The output of the following code is show in @tbl-nas-long-voi and the density plot of the calculated draws is in @fig-nas-prop-bm-draws-long-voicing-plot.
@@
 The plot suggests an overall greater nasalisation proportion in voiceless NC sequences.
-Let's quantify how greater as we did in Case Study 1.
+Let's quantify how greater as we did in @sec-case-1.
 We will use odd-ratios in this context as well, i.e.
 we will convert log-odds to odd-ratios using the `exp()` (exponential) function (and as before we first calculate the difference and then exponentiate the resulting values, after which we can take summary measures, like means and quantile-based measures such as CrIs).
-tbl-nas-prop-bm-diff-quant show the 90, 80 and 60% CrIs of the difference ratio of nasalisation proportion in voiceless vs voiced NC sequences.
+@tbl-nas-prop-bm-diff-quant show the 90, 80 and 60% CrIs of the difference ratio of nasalisation proportion in voiceless vs voiced NC sequences.
 
@@
 Moving onto question 2: is there individual speaker variation?
-\[TAMMINGA XXX\] For this, we will use the `spread_draws()` function from tidybayes \[XXX\] to extract the draws of the varying terms (in brms these are the coefficients that start with `r_`).
-There is quite a few steps of processing to get from the raw draws to the estimand we need: while we have commented the following code, we encourage readers to test each line sequentialy and inspect the intermediate output to fully understand the process.
+While in the recent tradition of linguistic research to use varying (aka random) effects to *control* for differences across participants, it has been proposed to use the estimates of the varying coefficients to investigate individual differences [@tamminga2016].
+We will use the `spread_draws()` function from tidybayes [@kay2019] to extract the draws of the varying terms (in brms these are the coefficients that start with `r_`).
+There is quite a few steps of processing to get from the raw draws to the estimand we need: while we have commented the following code, we encourage readers to test each line sequentially and inspect the intermediate output to fully understand the process.
 We assume that readers are familiar enough with models with varying terms (aka random effects, mixed-effects models).
@@
 Even among those speakers that do have a more robust positive difference, there is a lot of uncertainty as to the magnitude of the difference.
+
+# Conclusion
+
+In this tutorial we have presented two case studies where the outcome variable is a beta variable, i.e.
+a numeric continuous variable bounded between 0 and 1.
+We have shown how to fit Bayesian beta regressions to coarticulatory progressive voicing in voiceless stops in Italian in @sec-case-1 and coarticulatory regressive nasalisation in vowels in German in @sec-case-2.
+While this tutorial alone will not be sufficient to be able to independently run full Bayesian analyses, it will serve as a first introduction for readers to jump start their learning.

### 2025-02-07 manuscript/manuscript.tex
edit manuscript

@@
 like durations, frequencies, loudness and ratios
-\citep{kluender1988, johnson2003, gick2013, ladefoged2014, abramson2017, roettger2018, coretta2023}.
+\citep{kluender1988, johnson2003, gick2013, ladefoged2014, abramson2017, roettger2018, coretta2023, rosen2005}.
 Another commonly employed type of outcome variable are proportions: for
@@
   and hierarchical terms.} The categorical predictors vowel and C2 are
-coded using indexing rather than the traditional R contrasts: in R, this
-corresponds to suppressing the model's intercept with the \texttt{0\ +}
-syntax; using indexing instead of contrasts makes it easier to specify
-priors and uncertainty is distributed equally across all levels of
-categorical predictors. For pedagogical simplicity, the model will use
-the default priors as explained in Section~\ref{sec-bayes}.
+coded using indexing rather than the traditional R contrasts, as
+advocated in \citet{mcelreath2019}: in R, this corresponds to
+suppressing the model's intercept with the \texttt{0\ +} syntax; using
+indexing instead of contrasts makes it easier to specify priors and
+uncertainty is distributed equally across all levels of categorical
+predictors. For pedagogical simplicity, the model will use the default
+priors as explained in Section~\ref{sec-bayes}.
 
@@
 log-odds). This matches the general trends in the raw data, which we
-plotted in Figure~\ref{fig-ita-egg}.
-
-Next, we will plot the predicted proportions of each vowel/C2
-combination at mean speech rate (i.e.~centred speech rate = 0) and then
-calculate the average pair-wise difference in voicing proportion between
-/k, t, p/. Finally, we will assess whether there is greater
-between-vowel variability in /t, p/ relative to /k/.
+plotted in Figure~\ref{fig-ita-egg}. Next, we will plot the predicted
+proportions of each vowel/C2 combination at mean speech rate
+(i.e.~centred speech rate = 0) and then calculate the average pair-wise
+difference in voicing proportion between /k, t, p/.
 
 Before being able to plot the predictions, it's important to get
-familiar with the so-called posterior draws. {[}Bayesian MCMC XXX{]}.
+familiar with the so-called posterior draws. Bayesian regression models
+are fitted using Markov Chain Monte Carlo (MCMC) methods, as briefly
+explained in Section~\ref{sec-bayes}. The \texttt{brm()} function runs
+three MCMC chains with 2000 iterations each by default. These are
+sufficient in most cases. The first 1000 iterations of each chain are
+used for ``warmup'', where the MCMC algorithm is optimised to find the
+posterior probability. The second half of the iterations, 1000 per
+chain, are used for sampling from the posterior probability. Each
+iteration returns a drawn value for each parameter in the model. The
+list of values drawn during the Monte Carlo sampling are called the
+``posterior draws''. The posterior draws are used for inference: we can
+plot them, summarise them, transform them and combine them in any way to
+answer specific questions.
+
 Posterior draws can be conveniently obtained with the
-\texttt{as\_draws\_df()}. For the moment, we will extract only the draws
-of the constant regression coefficients (model variables starting with
-\texttt{b\_}). To check which coefficients are available in a model, use
-\texttt{get\_variables()} from the tidybayes package.
-\texttt{as\_draws\_df()} returns a tibble where each column contains the
-drawn values of a coefficient. Table~\ref{tbl-voi-prop-draws} shows the
-first ten rows and first five columns of the output of
-\texttt{as\_draws\_df()}. The probability distribution of the drawn
-values of each coefficient is the posterior probability distribution of
-that coefficient. Note that, due to the indexing coding of vowel and C2,
-all coefficient except \texttt{b\_speech\_rate\_c} are \emph{predicted
-log-odds} for each vowel/C2 combination (the drawn values for
-\texttt{b\_speech\_rate\_c} are drawn \emph{differences} in log-odds for
-each unit increase of speech rate). The drawn values are in log-odds,
-but we can convert them to proportions with \texttt{plogis()} (we will
-do this when plotting below).
+\texttt{as\_draws\_df()} from brms. For the moment, we will extract only
+the draws of the constant regression coefficients (model variables
+starting with \texttt{b\_}). To check which coefficients are available
+in a model, use \texttt{get\_variables()} from the tidybayes package
+\citep{kay2019}. \texttt{as\_draws\_df()} returns a tibble where each
+column contains the drawn values of a coefficient.
+Table~\ref{tbl-voi-prop-draws} shows the first ten rows and first five
+columns of the output of \texttt{as\_draws\_df()}. The probability
+distribution of the drawn values of each coefficient is the posterior
+probability distribution of that coefficient. Note that, due to the
+indexing coding of vowel and C2, all coefficient except
+\texttt{b\_speech\_rate\_c} are \emph{predicted log-odds} for each
+vowel/C2 combination (the drawn values for \texttt{b\_speech\_rate\_c}
+are drawn \emph{differences} in log-odds for each unit increase of
+speech rate). The drawn values are in log-odds, but we can convert them
+to proportions with \texttt{plogis()} (we will do this when plotting
+below).
 
@@
 For plotting, we can use ggplot2 statistics layers from the ggdist
-package. \texttt{stat\_halfeye()} plots the density of the posterior
-probability (in grey), its median (point) and CrIs (lines). Let's use 60
-and 80\% CrIs and transform the log-odds values to proportions with
-\texttt{plogis()}. See Figure~\ref{fig-voi-prop-bm-1}.
+package {[}\citet{kay2023}; kay2024{]}. \texttt{stat\_halfeye()} plots
+the density of the posterior probability (in grey), its median (point)
+and CrIs (lines). Let's use 60 and 80\% CrIs and transform the log-odds
+values to proportions with \texttt{plogis()}. See
+Figure~\ref{fig-voi-prop-bm-1}.
 
@@
   \FunctionTok{stat\_halfeye}\NormalTok{(}\AttributeTok{.width =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.6}\NormalTok{, }\FloatTok{0.8}\NormalTok{)) }\SpecialCharTok{+}
-  \FunctionTok{facet\_grid}\NormalTok{(}\AttributeTok{rows =} \FunctionTok{vars}\NormalTok{(c2))}
+  \FunctionTok{facet\_grid}\NormalTok{(}\AttributeTok{rows =} \FunctionTok{vars}\NormalTok{(c2)) }\SpecialCharTok{+}
+  \FunctionTok{labs}\NormalTok{(}
+    \AttributeTok{caption =} \StringTok{"The points are the medians and the two lines are 60 and 80\% CrIs."}
+\NormalTok{  )}
 \end{Highlighting}
@@
   \FunctionTok{ggplot}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\FunctionTok{plogis}\NormalTok{(value\_mean), c2)) }\SpecialCharTok{+}
-  \FunctionTok{stat\_halfeye}\NormalTok{(}\AttributeTok{.width =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.6}\NormalTok{, }\FloatTok{0.8}\NormalTok{))}
+  \FunctionTok{stat\_halfeye}\NormalTok{(}\AttributeTok{.width =} \FunctionTok{c}\NormalTok{(}\FloatTok{0.6}\NormalTok{, }\FloatTok{0.8}\NormalTok{)) }\SpecialCharTok{+}
+  \FunctionTok{labs}\NormalTok{(}
+    \AttributeTok{caption =} \StringTok{"The points are the medians and the two lines are 60 and 80\% CrIs."}
+\NormalTok{  )}
 \end{Highlighting}
@@
 Once we have the posterior difference, we can obtain CrIs of the
-difference using \texttt{quantile2()} from the posterior package. Beware
-that the values of the difference are in log-odds! We can convert these
-into odd-ratios with \texttt{exp()}. The output of the code below is
-shown in Table~\ref{tbl-voi-bm-quant}. Odd-ratios indicate the ratio of
-the difference between A and B, so that 1 means no difference, values
-greater than 1 indicate an increase in A relative to B and values lower
-than 1 indicate a decrease in A relative to B. odd-ratios are useful
-when looking at differences that are in log-odds because while the
-relative magnitude of the difference in proportion between two groups is
-the same independent of the baseline proportion, the \emph{absolute}
-magnitude of the difference depends on the baseline value. For example,
-an odd-ratio difference of 1.25 would correspond to a proportion
-increase of 13 percentage points if the baseline proportion is 0.62 but
-it would correspond to a proportion increase of 17 percentage points if
-the baseline proportion is 0.73. Of course, in real research contexts it
-is still usefull to think about absolute magnitudes and their relevance
-from a conceptual and methodological perspective. In this tutorial we
-just focus on odd-ratios for simplicity.
+difference using \texttt{quantile2()} from the posterior package
+\citep{burkner2024}. Beware that the values of the difference are in
+log-odds! We can convert these into odd-ratios with \texttt{exp()}. The
+output of the code below is shown in Table~\ref{tbl-voi-bm-quant}.
+Odd-ratios indicate the ratio of the difference between A and B, so that
+1 means no difference, values greater than 1 indicate an increase in A
+relative to B and values lower than 1 indicate a decrease in A relative
+to B. Odd-ratios are useful when looking at differences that are in
+log-odds because while the relative magnitude of the difference in
+proportion between two groups is the same independent of the baseline
+proportion, the \emph{absolute} magnitude of the difference depends on
+the baseline value. For example, an odd-ratio difference of 1.25 would
+correspond to a proportion increase of about 3 percentage points if the
+baseline proportion is 0.62 but it would correspond to a proportion
+increase of 5 percentage points if the baseline proportion is 0.73 and 1
+percentage point if the baseline is 0.99. Of course, in real research
+contexts it is still useful to think about absolute magnitudes and their
+relevance from a conceptual and methodological perspective. In this
+tutorial we just focus on odd-ratios for simplicity.
 
@@
 
-Finally, the package marginaleffects {[}XXX{]} has two other convenience
-functions that return CrIs of comparisons across predictor levels
-(\texttt{avg\_comparisons()}) and CrIs of posterior predictions across
-predictor levels (\texttt{avg\_predictions}). {[}XXX{]}
+Finally, the package marginaleffects \citep{arelbundock2025} has two
+other convenience functions that return CrIs of comparisons across
+predictor levels (\texttt{avg\_comparisons()}) and CrIs of posterior
+predictions across predictor levels (\texttt{avg\_predictions}). Both
+functions take the model and a list of variables as arguments.
+Optionally, you can set the confidence level with the
+\texttt{conf\_level} argument (here, we set it to 0.8, or 80\%).
+\texttt{avg\_comparisons()} allows the user to specify the type of
+comparison for each variable: here we set \texttt{"pairwise"} to
+calculate pairwise comparisons of each level of \texttt{c2}. The output
+of the two functions is shows in Table~\ref{tbl-marginal-1} and
+Table~\ref{tbl-marginal-2}.
 
@@
 (NT) should show earlier coarticulatory nasalisation than vowels
-followed by an NC sequence where C is voiced \citep[ND, see review
-in][]{carignan2021}. This pattern has been suggested to be driven by the
-articulatory and acoustic incompatibility of voicelessness and
-nasalisation, by which the velum opening gesture of the nasal consonant
-is pushed away (i.e.~earlier) when the consonant following the nasal is
-voiceless. Everything else being equal, a greater proportion of vowel
-nasalisation (from the perspective of time) should be found in vowels
-followed by NT than in vowels followed by ND.
+followed by an NC sequence where C is voiced (ND). This pattern has been
+suggested to be driven by the articulatory and acoustic incompatibility
+of voicelessness and nasalisation, by which the velum opening gesture of
+the nasal consonant is pushed away (i.e.~earlier) when the consonant
+following the nasal is voiceless
+\citep{fowler2005, beddor2009, cho2017, carignan2015}. Everything else
+being equal, a greater proportion of vowel nasalisation (from the
+perspective of time) should be found in vowels followed by NT than in
+vowels followed by ND.
 
@@
 \item
-  \texttt{NC} is the NC sequence.
+  \texttt{NC} is the NC sequence (voiceless /nt/ or voiced /nd/).
 \item
@@
 positive log-odds a proportion that is greater than 50\%. Generally, the
-expected log-odd predictions are negative, indicating an overall
-tendency for the nasalisation to take less than 50\% of the duration of
-the vowel. Moreover, the predictions are higher for voiceless NC
-sequences than for voiced NC sequences, indicating a greater proportion
-of nasalisation in the former. However there is vowel-specific
-variation, and there doesn't seem to be much of a difference in
-nasalisation proportion in /e/ and /u/. Figure~\ref{fig-nas-prop} shows
-the expected predictions with \texttt{conditional\_effects()}, which
-should make the interpretation clearer.
+expected log-odd predictions in Table~\ref{tbl-nas-prop-fixef} are
+negative, indicating an overall tendency for the nasalisation to take
+less than 50\% of the duration of the vowel. Moreover, the predictions
+are higher for voiceless NC sequences than for voiced NC sequences,
+indicating a greater proportion of nasalisation in the former. However
+there is vowel-specific variation, and there doesn't seem to be much of
+a difference in nasalisation proportion in /e/ and /u/.
+Figure~\ref{fig-nas-prop} shows the expected predictions with
+\texttt{conditional\_effects()}, which should make the interpretation of
+the predictions from Table~\ref{tbl-nas-prop-fixef} clearer.
 
@@
 
-Now that we fitted the model we can use the draws to answer the two
-research questions (repeated from above):
+Now that we fitted the model we can use the posterior draws to answer
+the two research questions (repeated from above):
 
@@
 Table~\ref{tbl-nas-draws} for the output of this step) and then take the
-difference of those, similarly to what we have done in the Case Study 1
-above.
+difference of those, similarly to what we have done in the case study in
+Section~\ref{sec-case-1}.
 
@@
 voicing, and plot the resulting posterior distributions. Note that, as
-discussed for Case study 1, when working wit log-odds it is important to
-first do all necessary calculations in log-odds, here calculate the mean
-log-odds across vowels, and \emph{then} transform the calculated
-estimands to proportions/probabilities. The output of the following code
-is show in Table~\ref{tbl-nas-long-voi} and the density plot of the
-calculated draws is in
+discussed in Section~\ref{sec-case-1}, when working with log-odds it is
+important to first do all necessary calculations in log-odds (here
+calculate the mean log-odds across vowels) and \emph{then} transform the
+calculated estimands to proportions/probabilities. The output of the
+following code is show in Table~\ref{tbl-nas-long-voi} and the density
+plot of the calculated draws is in
 Figure~\ref{fig-nas-prop-bm-draws-long-voicing-plot}.
@@
 The plot suggests an overall greater nasalisation proportion in
-voiceless NC sequences. Let's quantify how greater as we did in Case
-Study 1. We will use odd-ratios in this context as well, i.e.~we will
-convert log-odds to odd-ratios using the \texttt{exp()} (exponential)
-function (and as before we first calculate the difference and then
-exponentiate the resulting values, after which we can take summary
-measures, like means and quantile-based measures such as CrIs).
-tbl-nas-prop-bm-diff-quant show the 90, 80 and 60\% CrIs of the
-difference ratio of nasalisation proportion in voiceless vs voiced NC
-sequences.
+voiceless NC sequences. Let's quantify how greater as we did in
+Section~\ref{sec-case-1}. We will use odd-ratios in this context as
+well, i.e.~we will convert log-odds to odd-ratios using the
+\texttt{exp()} (exponential) function (and as before we first calculate
+the difference and then exponentiate the resulting values, after which
+we can take summary measures, like means and quantile-based measures
+such as CrIs). Table~\ref{tbl-nas-prop-bm-diff-quant} show the 90, 80
+and 60\% CrIs of the difference ratio of nasalisation proportion in
+voiceless vs voiced NC sequences.
 
@@
 
-Moving onto question 2: is there individual speaker variation?
-{[}TAMMINGA XXX{]} For this, we will use the \texttt{spread\_draws()}
-function from tidybayes {[}XXX{]} to extract the draws of the varying
-terms (in brms these are the coefficients that start with \texttt{r\_}).
-There is quite a few steps of processing to get from the raw draws to
-the estimand we need: while we have commented the following code, we
-encourage readers to test each line sequentialy and inspect the
-intermediate output to fully understand the process. We assume that
-readers are familiar enough with models with varying terms (aka random
-effects, mixed-effects models). What readers should note is that to
-obtain the expected predictions of nasalisation proportion for each
-speaker, the constant terms and the varying terms should be added (since
-the varying terms indicate the deviation of each speaker from the
-overall estimate).
+Moving onto question 2: is there individual speaker variation? While in
+the recent tradition of linguistic research to use varying (aka random)
+effects to \emph{control} for differences across participants, it has
+been proposed to use the estimates of the varying coefficients to
+investigate individual differences \citep{tamminga2016}. We will use the
+\texttt{spread\_draws()} function from tidybayes \citep{kay2019} to
+extract the draws of the varying terms (in brms these are the
+coefficients that start with \texttt{r\_}). There is quite a few steps
+of processing to get from the raw draws to the estimand we need: while
+we have commented the following code, we encourage readers to test each
+line sequentially and inspect the intermediate output to fully
+understand the process. We assume that readers are familiar enough with
+models with varying terms (aka random effects, mixed-effects models).
+What readers should note is that to obtain the expected predictions of
+nasalisation proportion for each speaker, the constant terms and the
+varying terms should be added (since the varying terms indicate the
+deviation of each speaker from the overall estimate).
 
@@
 
+\section{Conclusion}\label{conclusion}
+
+In this tutorial we have presented two case studies where the outcome
+variable is a beta variable, i.e.~a numeric continuous variable bounded
+between 0 and 1. We have shown how to fit Bayesian beta regressions to
+coarticulatory progressive voicing in voiceless stops in Italian in
+Section~\ref{sec-case-1} and coarticulatory regressive nasalisation in
+vowels in German in Section~\ref{sec-case-2}. While this tutorial alone
+will not be sufficient to be able to independently run full Bayesian
+analyses, it will serve as a first introduction for readers to jump
+start their learning.
+
   \bibliography{linguistics.bib,statistics.bib}

### 2025-02-10 manuscript/manuscript.qmd
add example of bad gaussian

@@
 This tacit default of using Gaussian models is also reflected in teaching practices, where significance test and models using the Gaussian distribution are the first to be taught [@baayen2008; @winter2020], due to their relative simplicity and the fact that regression models with other families are conceptual generalisations of Gaussian regression models.
+@fig-bg shows how poorly a Gaussian regression model fits beta-distributed data: the dark blue line is the density of data simulated to be distributed according to a beta distribution; the light blue lines are predicted distributions as generated by a Gaussian model of the data.
+The predicted Gaussian distributions over-predict values between 0.5 and 0.75 and under-predict values between 0.75 and 1.
+Moreover, the density peak (median) of the Gaussian distributions is lower than the peak in the original data.
+
+```{r}
+#| label: fig-bg
+#| fig-cap: "Gaussian model fitted to beta-distributed data.
+The dark blue line is the density of the beta data, while the light blue lines are simulated densities of the same data based on a Gaussian model."
+#| echo: false
+
+bg <- brm(
+  x ~ 1,
+  data = data.frame(x),
+  file = "data/cache/bg"
+)
+
+pp_check(bg, ndraws = 10)
+```
 

### 2025-02-10 manuscript/manuscript.tex
add example of bad gaussian

@@
 conceptual generalisations of Gaussian regression models.
+Figure~\ref{fig-bg} shows how poorly a Gaussian regression model fits
+beta-distributed data: the dark blue line is the density of data
+simulated to be distributed according to a beta distribution; the light
+blue lines are predicted distributions as generated by a Gaussian model
+of the data. The predicted Gaussian distributions over-predict values
+between 0.5 and 0.75 and under-predict values between 0.75 and 1.
+Moreover, the density peak (median) of the Gaussian distributions is
+lower than the peak in the original data.
+
+\begin{figure}
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-bg-1.pdf}}
+
+}
+
+\caption{\label{fig-bg}Gaussian model fitted to beta-distributed data.
+The dark blue line is the density of the beta data, while the light blue
+lines are simulated densities of the same data based on a Gaussian
+model.}
+
+\end{figure}%
 

### 2025-02-10 README.qmd
render pdf and html

---
title: "Bayesian beta regressions with brms in R: A tutorial for phoneticians"
format: md
---

This is the GitHub repository of the paper *Bayesian beta regressions with brms in R: A tutorial for phoneticians* by Stefano Coretta and Paul Bürkner.

The HTML version of the paper can be found [here](manuscript/manuscript.html).

### 2025-02-10 manuscript/manuscript.qmd
render pdf and html

@@
 ---
-title: A tutorial of Bayesian beta regressions with brms in R
+title: "Bayesian beta regressions with brms in R: A tutorial for phoneticians"
 author:
@@
       cite-style: authoryear
+  elsevier-html: default
 execute:

### 2025-02-10 manuscript/manuscript.tex
render pdf and html

@@
 \hypersetup{
-  pdftitle={A tutorial of Bayesian beta regressions with brms in R},
+  pdftitle={Bayesian beta regressions with brms in R: A tutorial for phoneticians},
   pdfauthor={Stefano Coretta; Paul Bürkner},
@@
 \begin{frontmatter}
-\title{A tutorial of Bayesian beta regressions with brms in R}
+\title{Bayesian beta regressions with brms in R: A tutorial for
+phoneticians}
 \author[1]{Stefano Coretta%

### 2025-02-10 README.qmd
render all

@@
 
-The HTML version of the paper can be found [here](manuscript/manuscript.html).
+The code of the manuscript is in `manuscript/manuscript.qmd`. The `code` folder contains code with data checks and plotting.
+
+The rendered HTML version of the paper and the data checks can be found on the [repository website](https://stefanocoretta.github.io/beta-phon/):
+
+- [manuscript](manuscript/manuscript.html).
+- [data checks](code/01_data_checks.html).

### 2025-02-10 manuscript/manuscript.qmd
render all

--- 2025-02-10 manuscript/manuscript.qmd
@@
           address: 3 George Sq
-          city: Edinburgh
-          state: United Kingdom
+          city: Edinburgh, United Kingdom
           postal-code: EH8 9AD
@@
     email: paul.buerkner@gmail.com
+    affiliations:
+        - id: dort
+          name: TU Dortmund University
+          department: Department of Statistics
+          adress: Technische Universität Dortmund
+          city: Dortmund, Germany
+          postal-code: 44221
 abstract: |
-  This is the abstract.
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Vestibulum augue turpis, dictum non malesuada a, volutpat eget velit.
-Nam placerat turpis purus, eu tristique ex tincidunt et.
-Mauris sed augue eget turpis ultrices tincidunt.
-Sed et mi in leo porta egestas.
-Aliquam non laoreet velit.
-Nunc quis ex vitae eros aliquet auctor nec ac libero.
-Duis laoreet sapien eu mi luctus, in bibendum leo molestie.
-Sed hendrerit diam diam, ac dapibus nisl volutpat vitae.
-Aliquam bibendum varius libero, eu efficitur justo rutrum at.
-Sed at tempus elit.
+  Phonetic research frequently involves analyzing numeric continuous outcome variables, such as durations, frequencies, loudness, and ratios.
+Another commonly used outcome type is proportions, including measures like the proportion of voicing during closure, gesture amplitude, and nasalance.
+Despite their bounded nature, proportions are often modeled using Gaussian regression, largely due to the default settings of commonly used statistical functions in R (e.g., lm() and lmer() from lme4).
+This practice persists in teaching and research, despite the fact that Gaussian models assume unbounded continuous data and may poorly fit proportion data.
+To address this issue, this tutorial introduces beta regression models, a more appropriate statistical approach for analyzing proportions.
+The beta distribution provides a flexible framework for modelling continuous data constrained between 0 and 1.
+The tutorial employs the brms package in R and assumes familiarity with regression modeling but no prior knowledge of Bayesian statistics.
+The tutorial includes two case studies illustrating the practical implementation of Bayesian beta regression models.
+Data and code are available at <https://github.com/stefanocoretta/beta-phon>.
 keywords: 
-  - keyword1
-  - keyword2
+  - acoustics
+  - articulation
+  - proportions
+  - Bayesian regressions
+  - beta distribution
 date: last-modified
@@
 The rest of the paper is structured as follows: @sec-beta introduces the mathematical underpinnings of the beta distribution, @sec-bayes provides the readers with a brief explanation and justification of Bayesian regression models, while @sec-case-1 and @sec-case-2 illustrate how to fit Bayesian beta regression models with two case studies.
+The research compendium with data and code can be found at <https://github.com/stefanocoretta/beta-phon>.
 

### 2025-02-10 manuscript/manuscript.tex
render all

@@
   pdfauthor={Stefano Coretta; Paul Bürkner},
-  pdfkeywords={keyword1, keyword2},
+  pdfkeywords={acoustics, articulation, proportions, Bayesian
+regressions, beta distribution},
   colorlinks=true,
@@
  \ead{s.coretta@ed.ac.uk} 
-\author[]{Paul Bürkner%
+\author[2]{Paul Bürkner%
 %
@@
 \affiliation[1]{organization={University of Edinburgh, Linguistics and
-English Language},addressline={3 George
-Sq},city={Edinburgh},postcode={EH8 9AD},postcodesep={}}
+English Language},addressline={3 George Sq},city={Edinburgh, United
+Kingdom},postcode={EH8 9AD},postcodesep={}}
+\affiliation[2]{organization={TU Dortmund University, Department of
+Statistics},city={Dortmund, Germany},postcode={44221},postcodesep={}}
 
@@
 \begin{abstract}
-This is the abstract. Lorem ipsum dolor sit amet, consectetur adipiscing
-elit. Vestibulum augue turpis, dictum non malesuada a, volutpat eget
-velit. Nam placerat turpis purus, eu tristique ex tincidunt et. Mauris
-sed augue eget turpis ultrices tincidunt. Sed et mi in leo porta
-egestas. Aliquam non laoreet velit. Nunc quis ex vitae eros aliquet
-auctor nec ac libero. Duis laoreet sapien eu mi luctus, in bibendum leo
-molestie. Sed hendrerit diam diam, ac dapibus nisl volutpat vitae.
-Aliquam bibendum varius libero, eu efficitur justo rutrum at. Sed at
-tempus elit.
+Phonetic research frequently involves analyzing numeric continuous
+outcome variables, such as durations, frequencies, loudness, and ratios.
+Another commonly used outcome type is proportions, including measures
+like the proportion of voicing during closure, gesture amplitude, and
+nasalance. Despite their bounded nature, proportions are often modeled
+using Gaussian regression, largely due to the default settings of
+commonly used statistical functions in R (e.g., lm() and lmer() from
+lme4). This practice persists in teaching and research, despite the fact
+that Gaussian models assume unbounded continuous data and may poorly fit
+proportion data. To address this issue, this tutorial introduces beta
+regression models, a more appropriate statistical approach for analyzing
+proportions. The beta distribution provides a flexible framework for
+modelling continuous data constrained between 0 and 1. The tutorial
+employs the brms package in R and assumes familiarity with regression
+modeling but no prior knowledge of Bayesian statistics. The tutorial
+includes two case studies illustrating the practical implementation of
+Bayesian beta regression models. Data and code are available at
+\url{https://github.com/stefanocoretta/beta-phon}.
 \end{abstract}
@@
 \begin{keyword}
-    keyword1 \sep 
-    keyword2
+    acoustics \sep articulation \sep proportions \sep Bayesian
+regressions \sep 
+    beta distribution
 \end{keyword}
@@
 Section~\ref{sec-case-1} and Section~\ref{sec-case-2} illustrate how to
-fit Bayesian beta regression models with two case studies.
+fit Bayesian beta regression models with two case studies. The research
+compendium with data and code can be found at
+\url{https://github.com/stefanocoretta/beta-phon}.
 

### 2025-02-11 manuscript/manuscript.qmd
clear cache and rerender
fix typos

@@
 library(tidyverse)
+theme_set(theme_light())
 library(brms)
@@
 library(marginaleffects)
+library(posterior)
 ```
@@
 95, 90, 80, 60%) for a more complete view on estimated parameters.
-Bayesian CrIs indicate that at a certain probability levels the parameter's estimate lies within that interval: so, for example, a 90% CrI \[A, B\] indicates that there is a 90% probability that the estimate is between A and B.
+Bayesian CrIs indicate that, at a certain probability level, the parameter's estimate lies within that interval: so, for example, a 90% CrI \[A, B\] indicates that there is a 90% probability that the estimate is between A and B.
 Different probability levels correspond to different levels of confidence: the higher the probability the higher the confidence (always conditional on data and model).
@@
 Moreover, there is greater variability between vowels in /t, p/ than in /k/.
-We will use a beta regression model to assess whether the place of articulation of the consonant affects the proportion of coarticulatory progressive voicing, while averaging across vowels (of course, another question that could be answered with the same model is if different vowels lead to difference voicing proportions, but for sake of space we restrict this tutorial to the former question).
+We will use a beta regression model to assess whether the place of articulation of the consonant affects the proportion of coarticulatory progressive voicing, while averaging across vowels (of course, another question that could be answered with the same model is if different vowels lead to different voicing proportions, but for sake of space we restrict this tutorial to the former question).
 A general expectation is that the proportion of progressive voicing should increase with fronter places of articulation (so from /k/ to /t/ to /p/) given the aerodynamic constraints which obtain with vocal fold vibration during a fully occluded oral tract [@vandenberg1958; @rothenberg1967; @ohala2011].
@@
 For conciseness, we will use the `fixef()` function which prints the regression coefficients.
-The full summary with an explanation of each part can be found in XXX.
 @tbl-voi-prop-fixef reports the output of `fixef()` as a table (we round all values to the nearest 2 digits for clarity.).
@@
 Bayesian regression models are fitted using Markov Chain Monte Carlo (MCMC) methods, as briefly explained in @sec-bayes.
-The `brm()` function runs three MCMC chains with 2000 iterations each by default.
+The `brm()` function runs four MCMC chains with 2000 iterations each by default.
 These are sufficient in most cases.
@@
 The list of values drawn during the Monte Carlo sampling are called the "posterior draws".
+Since four chains are run, a total of 4000 posterior values are generated.
 The posterior draws are used for inference: we can plot them, summarise them, transform them and combine them in any way to answer specific questions.
@@
 Posterior draws can be conveniently obtained with the `as_draws_df()` from brms.
-For the moment, we will extract only the draws of the constant regression coefficients (model variables starting with `b_`).
+For the moment, we will extract only the draws of the constant (aka fixed) regression coefficients (model variables starting with `b_`).
 To check which coefficients are available in a model, use `get_variables()` from the tidybayes package [@kay2019].
@@
 The probability distribution of the drawn values of each coefficient is the posterior probability distribution of that coefficient.
-Note that, due to the indexing coding of vowel and C2, all coefficient except `b_speech_rate_c` are *predicted log-odds* for each vowel/C2 combination (the drawn values for `b_speech_rate_c` are drawn *differences* in log-odds for each unit increase of speech rate).
+Note that, due to the indexing coding of vowel and C2, all coefficient except `b_speech_rate_c` are drawn *log-odds values* for each vowel/C2 combination (the drawn values for `b_speech_rate_c` are drawn *differences* in log-odds values for each unit increase of speech rate).
 The drawn values are in log-odds, but we can convert them to proportions with `plogis()` (we will do this when plotting below).
@@
 #| label: voi-prop-bm-draws
+#| warning: false
 
@@
 One approach is to take the mean across vowels within each consonant for each posterior draw, and the posterior distribution of the resulting list of values is the predicted posterior distribution of voicing proportion for each consonant, assuming an "average" vowel.
+See @fig-voi-pbdl-c2.
 
@@
 At 80% confidence, the change ratio is 1.10-1.34 (or 10-34% increase) while at 60% confidence is 1.14-1.30 (14-30% increase).
-In other words we can be quite confident that the voicing proportion in /p/ is longer than in /t, k/ and that the increase is less than 35%.
+In other words we can be quite confident that the voicing proportion in /p/ is greater than in /t, k/ and that the increase is less than 35%.
 The brms package comes with a convenient function, `conditional_effects()`, to plot posterior means and CrIs based on predictors in the model.
@@
 `avg_comparisons()` allows the user to specify the type of comparison for each variable: here we set `"pairwise"` to calculate pairwise comparisons of each level of `c2`.
-The output of the two functions is shows in @tbl-marginal-1 and @tbl-marginal-2.
+The output of the two functions is shown in @tbl-marginal-1 and @tbl-marginal-2.
 
@@
 Moving onto question 2: is there individual speaker variation?
-While in the recent tradition of linguistic research to use varying (aka random) effects to *control* for differences across participants, it has been proposed to use the estimates of the varying coefficients to investigate individual differences [@tamminga2016].
+While in the recent tradition of linguistic research, varying (aka random) effects are used to *control* for differences across participants, it has been proposed to use the estimates of the varying coefficients to investigate individual differences [@tamminga2016].
 We will use the `spread_draws()` function from tidybayes [@kay2019] to extract the draws of the varying terms (in brms these are the coefficients that start with `r_`).
 There is quite a few steps of processing to get from the raw draws to the estimand we need: while we have commented the following code, we encourage readers to test each line sequentially and inspect the intermediate output to fully understand the process.
-We assume that readers are familiar enough with models with varying terms (aka random effects, mixed-effects models).
+We assume that readers are familiar enough with models with varying terms (aka random effects, mixed-effects, hierarchical, multilevel, nested models).
 What readers should note is that to obtain the expected predictions of nasalisation proportion for each speaker, the constant terms and the varying terms should be added (since the varying terms indicate the deviation of each speaker from the overall estimate).

### 2025-02-11 manuscript/manuscript.tex
fix typos

@@
 levels (e.g.~95, 90, 80, 60\%) for a more complete view on estimated
-parameters. Bayesian CrIs indicate that at a certain probability levels
+parameters. Bayesian CrIs indicate that, at a certain probability level,
 the parameter's estimate lies within that interval: so, for example, a
@@
 could be answered with the same model is if different vowels lead to
-difference voicing proportions, but for sake of space we restrict this
+different voicing proportions, but for sake of space we restrict this
 tutorial to the former question). A general expectation is that the
@@
 conciseness, we will use the \texttt{fixef()} function which prints the
-regression coefficients. The full summary with an explanation of each
-part can be found in XXX. Table~\ref{tbl-voi-prop-fixef} reports the
+regression coefficients. Table~\ref{tbl-voi-prop-fixef} reports the
 output of \texttt{fixef()} as a table (we round all values to the
@@
 explained in Section~\ref{sec-bayes}. The \texttt{brm()} function runs
-three MCMC chains with 2000 iterations each by default. These are
+four MCMC chains with 2000 iterations each by default. These are
 sufficient in most cases. The first 1000 iterations of each chain are
@@
 list of values drawn during the Monte Carlo sampling are called the
-``posterior draws''. The posterior draws are used for inference: we can
-plot them, summarise them, transform them and combine them in any way to
-answer specific questions.
+``posterior draws''. Since four chains are run, a total of 4000
+posterior values are generated. The posterior draws are used for
+inference: we can plot them, summarise them, transform them and combine
+them in any way to answer specific questions.
 
@@
 \texttt{as\_draws\_df()} from brms. For the moment, we will extract only
-the draws of the constant regression coefficients (model variables
-starting with \texttt{b\_}). To check which coefficients are available
-in a model, use \texttt{get\_variables()} from the tidybayes package
-\citep{kay2019}. \texttt{as\_draws\_df()} returns a tibble where each
-column contains the drawn values of a coefficient.
+the draws of the constant (aka fixed) regression coefficients (model
+variables starting with \texttt{b\_}). To check which coefficients are
+available in a model, use \texttt{get\_variables()} from the tidybayes
+package \citep{kay2019}. \texttt{as\_draws\_df()} returns a tibble where
+each column contains the drawn values of a coefficient.
 Table~\ref{tbl-voi-prop-draws} shows the first ten rows and first five
@@
 indexing coding of vowel and C2, all coefficient except
-\texttt{b\_speech\_rate\_c} are \emph{predicted log-odds} for each
+\texttt{b\_speech\_rate\_c} are drawn \emph{log-odds values} for each
 vowel/C2 combination (the drawn values for \texttt{b\_speech\_rate\_c}
-are drawn \emph{differences} in log-odds for each unit increase of
-speech rate). The drawn values are in log-odds, but we can convert them
-to proportions with \texttt{plogis()} (we will do this when plotting
-below).
+are drawn \emph{differences} in log-odds values for each unit increase
+of speech rate). The drawn values are in log-odds, but we can convert
+them to proportions with \texttt{plogis()} (we will do this when
+plotting below).
 
@@
 distribution of voicing proportion for each consonant, assuming an
-``average'' vowel.
+``average'' vowel. See Figure~\ref{fig-voi-pbdl-c2}.
 
@@
 increase). In other words we can be quite confident that the voicing
-proportion in /p/ is longer than in /t, k/ and that the increase is less
-than 35\%. The brms package comes with a convenient function,
+proportion in /p/ is greater than in /t, k/ and that the increase is
+less than 35\%. The brms package comes with a convenient function,
 \texttt{conditional\_effects()}, to plot posterior means and CrIs based
@@
 calculate pairwise comparisons of each level of \texttt{c2}. The output
-of the two functions is shows in Table~\ref{tbl-marginal-1} and
+of the two functions is shown in Table~\ref{tbl-marginal-1} and
 Table~\ref{tbl-marginal-2}.
@@
 Moving onto question 2: is there individual speaker variation? While in
-the recent tradition of linguistic research to use varying (aka random)
-effects to \emph{control} for differences across participants, it has
-been proposed to use the estimates of the varying coefficients to
+the recent tradition of linguistic research, varying (aka random)
+effects are used to \emph{control} for differences across participants,
+it has been proposed to use the estimates of the varying coefficients to
 investigate individual differences \citep{tamminga2016}. We will use the
@@
 understand the process. We assume that readers are familiar enough with
-models with varying terms (aka random effects, mixed-effects models).
-What readers should note is that to obtain the expected predictions of
-nasalisation proportion for each speaker, the constant terms and the
-varying terms should be added (since the varying terms indicate the
-deviation of each speaker from the overall estimate).
+models with varying terms (aka random effects, mixed-effects,
+hierarchical, multilevel, nested models). What readers should note is
+that to obtain the expected predictions of nasalisation proportion for
+each speaker, the constant terms and the varying terms should be added
+(since the varying terms indicate the deviation of each speaker from the
+overall estimate).
 

### 2025-02-14 manuscript/manuscript.qmd
add preprint doi

@@
 title: "Bayesian beta regressions with brms in R: A tutorial for phoneticians"
+doi: "10.31219/osf.io/f9rqg_v1"
 author:
